---
title: "Metagenome_analyses"
author: "Julia Gendre"
date: "2024-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Redaction/Phage_dynamic_fermented_food")
```

## Packages loading

```{r packages, message=FALSE, warning=FALSE}
library(phyloseq)
library(readxl)  
library(phyloseq.extended)
library(vegan)
library(gridExtra)
library(rstatix)
library(gplots)
library(tidyverse)
library(DT)
library(pheatmap)
library(DESeq2)
library(ggplotify)
library(cowplot)
library(ggpubr)
library(ggforce)
library(ggplot2)
library(ggrepel)
library(ggbreak)
library(VennDiagram)
library(grid)
library(factoextra)
library(Hmisc)
library(dplyr)
sessionInfo()

```

# Kaiju data analyses

## Function creation to treat Kaiju data

```{r function, message=FALSE, warning=FALSE}
# Function to treat kaiju data 
process_kaiju_data <- function(data, Samples_name) {
  # Add colonne 'Samples' to raw data vith Samples_name variable and select column : percentage of reads belong to taxa, number of reads belongt to taxa, taxon_id et taxon_name.
  data <- data %>%
    mutate(Samples = Samples_name) %>%
    select(Samples, percent, reads, taxon_id, taxon_name)
  
  # split taxon_name column according to different taxonomic rank.
  data <- data %>%
    separate(taxon_name, into = c("superkingdom", "phylum", "class", "order", "family", "genus", "species"), sep = ";")
  
  # Apply conditional modifications using `mutate` and `case_when`.
  data <- data %>%
    mutate(
      superkingdom = case_when(
        superkingdom == "cannot be assigned to a (non-viral) genus" ~ "unclassified",
        TRUE ~ superkingdom
      ),
      genus = case_when(
        superkingdom == "unclassified" ~ "unclassified",
        TRUE ~ genus
      ),
      family = case_when(
        superkingdom == "unclassified" ~ "unclassified",
        TRUE ~ family
      ),
      order = case_when(
        superkingdom == "unclassified" ~ "unclassified",
        TRUE ~ order
      )
    )
  
  # Filtrate Bacteria and Eukaryota data
  bacteria_data <- filter(data, superkingdom == "Bacteria")
  euka_data <- filter(data, superkingdom == "Eukaryota")
  viruses_data <- filter(data, superkingdom == "Viruses")
  
  # The function returns processed data and the filtered subsets.
  return(list(
    processed_data = data,
    bacteria = bacteria_data,
    eukaryota = euka_data,
    viruses = viruses_data
  ))
}

#Function to summary of domain taxon for all Samples
abundance_summary_ki <- function(data, group_column, value_column, new_column_name) {
  # Summary of abundance
  summary_data <- data %>%
    select({{group_column}}, {{value_column}}) %>%
    group_by({{group_column}}) %>%
    summarise(total_count = sum({{value_column}}, na.rm = TRUE))
  
  # Rename columns
  names(summary_data) <- c(as.character(substitute(group_column)), new_column_name)
  
  return(summary_data)
}

# Function to summary genus taxon for all Samples  
summarize_genus_abundance <- function(data, new_column_name, genus_col = "genus", reads_col = "reads") {
  
  # Selecte column
  summary <- data %>%
    select(!!sym(genus_col), !!sym(reads_col)) %>%
    
    # Group by 'genus' and sum 'reads'
    group_by_at(genus_col) %>%
    summarise(total_count = sum(!!sym(reads_col), na.rm = TRUE)) %>%
    ungroup()
  
  # Sum total 'reads'
  column_sum <- sum(summary$total_count, na.rm = TRUE)
  
  # renames columns
  names(summary) <- c("genus", "reads")
  
  # Calcul percentage of reads per genus
  summary <- summary %>%
    mutate(percent = (reads / column_sum) * 100)
  
  # group per genu and sum percentage
  summary_per <- summary %>%
    group_by(genus) %>%
    summarise(total_count = sum(percent, na.rm = TRUE)) %>%
    ungroup()
  
  # rename columns
  names(summary_per) <- c("genus", new_column_name)

  
  # final dataframe
  return(summary_per)
}  
  
# Function to summary species taxon for all Samples  
summarize_species_abundance <- function(data, new_column_name, species_col = "species", reads_col = "reads") {
  
  # Selecte column
  summary <- data %>%
    select(!!sym(species_col), !!sym(reads_col)) %>%
    
    # Group by 'species' and sum 'reads'
    group_by_at(species_col) %>%
    summarise(total_count = sum(!!sym(reads_col), na.rm = TRUE)) %>%
    ungroup()
  
  # Sum total 'reads'
  column_sum <- sum(summary$total_count, na.rm = TRUE)
  
  # renames columns
  names(summary) <- c("species", "reads")
  
  # Calcul percentage of reads per species
  summary <- summary %>%
    mutate(percent = (reads / column_sum) * 100)
  
  # group per genu and sum percentage
  summary_per <- summary %>%
    group_by(species) %>%
    summarise(total_count = sum(percent, na.rm = TRUE)) %>%
    ungroup()
  
  # rename columns
  names(summary_per) <- c("species", new_column_name)

  
  # final dataframe
  return(summary_per)
}
```

## Analyses of the cabbage fermentation

## Import fermentation metadata

```{r Import data sauerkraut, message=FALSE, warning=FALSE}

Fermentation_metadata <- read.csv("results/pH_acidity_microbial_results.txt", sep = "\t",dec= ",",header = TRUE,
                                    fileEncoding = "latin1")

Fermentation_metadata$Day_sampling [Fermentation_metadata$Day_sampling=="22"]<- "21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T0"]<- "Day 0"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T2"]<- "Day 2"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T7"]<- "Day 7"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T14"]<- "Day 14"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T21"]<- "Day 21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T22"]<- "Day 21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T28"]<- "Day 28"


Fermentation_metadata$Day_sampling <- as.numeric(Fermentation_metadata$Day_sampling)

Fermentation_metadata$Day_sampling_text <- factor(Fermentation_metadata$Day_sampling_text, levels = c("Day 0", "Day 2", "Day 7","Day 14","Day 21","Day 28"))

```

### Import kaiju results from cabbage fermentation

```{r Import data sauerkraut, message=FALSE, warning=FALSE}
setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

SA_0A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_0A_libLAH8279_summary_viruses.tsv",sep="\t",header = TRUE)
SA_0B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_0B_libLAH8280_summary_viruses.tsv",sep="\t",header = TRUE)
SA_0C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_0C_libLAH8281_summary_viruses.tsv",sep="\t",header = TRUE)

SA_2A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_2A_libLAH8282_summary_viruses.tsv", sep="\t",header = TRUE)
SA_2B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_2B_libLAH8283_summary_viruses.tsv", sep="\t",header = TRUE)
SA_2C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_2C_libLAH8284_summary_viruses.tsv", sep="\t",header = TRUE)

SA_7A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_7A_libLAH8285_summary_viruses.tsv", sep="\t",header = TRUE)
SA_7B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_7B_libLAH8286_summary_viruses.tsv", sep="\t",header = TRUE)
SA_7C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_7C_libLAH8287_summary_viruses.tsv", sep="\t",header = TRUE)

SA_14A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_14A_libLAH8288_summary_viruses.tsv", sep="\t",header = TRUE)
SA_14B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_14B_libLAH8289_summary_viruses.tsv", sep="\t",header = TRUE)
SA_14C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_14C_libLAH8290_summary_viruses.tsv", sep="\t",header = TRUE)

SA_21A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_21A_libLAH8291_summary_viruses.tsv", sep="\t",header = TRUE)
SA_21B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_21B_libLAH8292_summary_viruses.tsv", sep="\t",header = TRUE)
SA_21C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_21C_libLAH8293_summary_viruses.tsv", sep="\t",header = TRUE)

SA_28A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_28A_libLAH8294_summary_viruses.tsv", sep="\t",header = TRUE)
SA_28B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_28B_libLAH8295_summary_viruses.tsv", sep="\t",header = TRUE)
SA_28C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_28C_libLAH8296_summary_viruses.tsv", sep="\t",header = TRUE)

res_SA_0A <- process_kaiju_data(SA_0A_Kaiju, "SA_0A")
res_SA_0B <- process_kaiju_data(SA_0B_Kaiju, "SA_0B")
res_SA_0C <- process_kaiju_data(SA_0C_Kaiju, "SA_0C")
res_SA_2A <- process_kaiju_data(SA_2A_Kaiju, "SA_2A")
res_SA_2B <- process_kaiju_data(SA_2B_Kaiju, "SA_2B")
res_SA_2C <- process_kaiju_data(SA_2C_Kaiju, "SA_2C")
res_SA_7A <- process_kaiju_data(SA_7A_Kaiju, "SA_7A")
res_SA_7B <- process_kaiju_data(SA_7B_Kaiju, "SA_7B")
res_SA_7C <- process_kaiju_data(SA_7C_Kaiju, "SA_7C")
res_SA_14A <- process_kaiju_data(SA_14A_Kaiju, "SA_14A")
res_SA_14B <- process_kaiju_data(SA_14B_Kaiju, "SA_14B")
res_SA_14C <- process_kaiju_data(SA_14C_Kaiju, "SA_14C")
res_SA_21A <- process_kaiju_data(SA_21A_Kaiju, "SA_21A")
res_SA_21B <- process_kaiju_data(SA_21B_Kaiju, "SA_21B")
res_SA_21C <- process_kaiju_data(SA_21C_Kaiju, "SA_21C")
res_SA_28A <- process_kaiju_data(SA_28A_Kaiju, "SA_28A")
res_SA_28B <- process_kaiju_data(SA_28B_Kaiju, "SA_28B")
res_SA_28C <- process_kaiju_data(SA_28C_Kaiju, "SA_28C")


SA_0A_Kaiju_process <- res_SA_0A$processed_data
SA_0B_Kaiju_process <- res_SA_0B$processed_data
SA_0C_Kaiju_process <- res_SA_0C$processed_data
SA_2A_Kaiju_process <- res_SA_2A$processed_data
SA_2B_Kaiju_process <- res_SA_2B$processed_data
SA_2C_Kaiju_process <- res_SA_2C$processed_data
SA_7A_Kaiju_process <- res_SA_7A$processed_data
SA_7B_Kaiju_process <- res_SA_7B$processed_data
SA_7C_Kaiju_process <- res_SA_7C$processed_data
SA_14A_Kaiju_process <- res_SA_14A$processed_data
SA_14B_Kaiju_process <- res_SA_14B$processed_data
SA_14C_Kaiju_process <- res_SA_14C$processed_data
SA_21A_Kaiju_process <- res_SA_21A$processed_data
SA_21B_Kaiju_process <- res_SA_21B$processed_data
SA_21C_Kaiju_process <- res_SA_21C$processed_data
SA_28A_Kaiju_process <- res_SA_28A$processed_data
SA_28B_Kaiju_process <- res_SA_28B$processed_data
SA_28C_Kaiju_process <- res_SA_28C$processed_data

abundance_kaiju_cabbage <- full_join (SA_0A_Kaiju_process,SA_0B_Kaiju_process)%>%
  full_join(.,SA_0C_Kaiju_process)%>%
  full_join(.,SA_2A_Kaiju_process)%>%
  full_join(.,SA_2B_Kaiju_process)%>%
  full_join(.,SA_2C_Kaiju_process)%>%
  full_join(.,SA_7A_Kaiju_process)%>%
  full_join(.,SA_7B_Kaiju_process)%>%
  full_join(.,SA_7C_Kaiju_process)%>%
  full_join(.,SA_14A_Kaiju_process)%>%
  full_join(.,SA_14B_Kaiju_process)%>%
  full_join(.,SA_14C_Kaiju_process)%>%
  full_join(.,SA_21A_Kaiju_process)%>%
  full_join(.,SA_21B_Kaiju_process)%>%
  full_join(.,SA_21C_Kaiju_process)%>%
  full_join(.,SA_28A_Kaiju_process)%>%
  full_join(.,SA_28B_Kaiju_process)%>%
  full_join(.,SA_28C_Kaiju_process)

```

### Microbial community at the domain level

Relative abundance for reads for each kingdom were calculated and represented by bar plot.

```{r sauerkraut kingdom plot, message=FALSE, warning=FALSE}

abundance_SA_0A_kingdom <- abundance_summary_ki(SA_0A_Kaiju_process, superkingdom, percent, "SA_0A")
abundance_SA_0B_kingdom <- abundance_summary_ki(SA_0B_Kaiju_process, superkingdom, percent, "SA_0B")
abundance_SA_0C_kingdom <- abundance_summary_ki(SA_0C_Kaiju_process, superkingdom, percent, "SA_0C")
abundance_SA_2A_kingdom <- abundance_summary_ki(SA_2A_Kaiju_process, superkingdom, percent, "SA_2A")
abundance_SA_2B_kingdom <- abundance_summary_ki(SA_2B_Kaiju_process, superkingdom, percent, "SA_2B")
abundance_SA_2C_kingdom <- abundance_summary_ki(SA_2C_Kaiju_process, superkingdom, percent, "SA_2C")
abundance_SA_7A_kingdom <- abundance_summary_ki(SA_7A_Kaiju_process, superkingdom, percent, "SA_7A")
abundance_SA_7B_kingdom <- abundance_summary_ki(SA_7B_Kaiju_process, superkingdom, percent, "SA_7B")
abundance_SA_7C_kingdom <- abundance_summary_ki(SA_7C_Kaiju_process, superkingdom, percent, "SA_7C")
abundance_SA_14A_kingdom <- abundance_summary_ki(SA_14A_Kaiju_process, superkingdom, percent, "SA_14A")
abundance_SA_14B_kingdom <- abundance_summary_ki(SA_14B_Kaiju_process, superkingdom, percent, "SA_14B")
abundance_SA_14C_kingdom <- abundance_summary_ki(SA_14C_Kaiju_process, superkingdom, percent, "SA_14C")
abundance_SA_21A_kingdom <- abundance_summary_ki(SA_21A_Kaiju_process, superkingdom, percent, "SA_21A")
abundance_SA_21B_kingdom <- abundance_summary_ki(SA_21B_Kaiju_process, superkingdom, percent, "SA_21B")
abundance_SA_21C_kingdom <- abundance_summary_ki(SA_21C_Kaiju_process, superkingdom, percent, "SA_21C")
abundance_SA_28A_kingdom <- abundance_summary_ki(SA_28A_Kaiju_process, superkingdom, percent, "SA_28A")
abundance_SA_28B_kingdom <- abundance_summary_ki(SA_28B_Kaiju_process, superkingdom, percent, "SA_28B")
abundance_SA_28C_kingdom <- abundance_summary_ki(SA_28C_Kaiju_process, superkingdom, percent, "SA_28C")

abundance_kaiju_kingdom_cabb <- full_join(abundance_SA_0A_kingdom,abundance_SA_0B_kingdom)%>%
  full_join(.,abundance_SA_0C_kingdom)%>%
  full_join(.,abundance_SA_2A_kingdom)%>%
  full_join(.,abundance_SA_2B_kingdom)%>%
  full_join(.,abundance_SA_2C_kingdom)%>%
  full_join(.,abundance_SA_7A_kingdom)%>%
  full_join(.,abundance_SA_7B_kingdom)%>%
  full_join(.,abundance_SA_7C_kingdom)%>%
  full_join(.,abundance_SA_14A_kingdom)%>%
  full_join(.,abundance_SA_14B_kingdom)%>%
  full_join(.,abundance_SA_14C_kingdom)%>%
  full_join(.,abundance_SA_21A_kingdom)%>%
  full_join(.,abundance_SA_21B_kingdom)%>%
  full_join(.,abundance_SA_21C_kingdom)%>%
  full_join(.,abundance_SA_28A_kingdom)%>%
  full_join(.,abundance_SA_28B_kingdom)%>%
  full_join(.,abundance_SA_28C_kingdom)

# Convertir en format long pour ggplot2
abundance_long_cabb <- abundance_kaiju_kingdom_cabb %>%
  pivot_longer(cols = -superkingdom, names_to = "Samples", values_to = "abundance")

abundance_long_cabb$Samples <- factor(abundance_long_cabb$Samples, levels = c("SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C"))

abundance_long_cabb$superkingdom <- factor(abundance_long_cabb$superkingdom , levels = c("Archaea","Bacteria","Eukaryota","Viruses","unclassified"))
kingdom_col <- c("Archaea"='#d7191c',"Bacteria"='#fdae61',"Eukaryota"='#ffffbf',"unclassified"='#abd9e9', "Viruses"='#2c7bb6' )

abundance_long_superkingdom_cabbage2 <- left_join(abundance_long_cabb,Fermentation_metadata, by = c("Samples" = "Samples_id"))

per_read_kingdom_cabb <- ggplot(abundance_long_superkingdom_cabbage2, aes(fill=superkingdom, y=abundance, x=Samples)) +
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Kingdom")+
  scale_fill_manual(values = kingdom_col)+
  theme(legend.text = element_text(size = 11,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 11))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_kingdom_cabb


```

### Bacterial community at the genus level

Relative abundance for reads for each kingdom were calculated and represented by bar plot.

```{r sauerkraut bacteria plot, message=FALSE, warning=FALSE}

SA_0A_Kaiju_bacteria <- res_SA_0A$bacteria
SA_0B_Kaiju_bacteria <- res_SA_0B$bacteria
SA_0C_Kaiju_bacteria <- res_SA_0C$bacteria
SA_2A_Kaiju_bacteria <- res_SA_2A$bacteria
SA_2B_Kaiju_bacteria <- res_SA_2B$bacteria
SA_2C_Kaiju_bacteria <- res_SA_2C$bacteria
SA_7A_Kaiju_bacteria <- res_SA_7A$bacteria
SA_7B_Kaiju_bacteria <- res_SA_7B$bacteria
SA_7C_Kaiju_bacteria <- res_SA_7C$bacteria
SA_14A_Kaiju_bacteria <- res_SA_14A$bacteria
SA_14B_Kaiju_bacteria <- res_SA_14B$bacteria
SA_14C_Kaiju_bacteria <- res_SA_14C$bacteria
SA_21A_Kaiju_bacteria <- res_SA_21A$bacteria
SA_21B_Kaiju_bacteria <- res_SA_21B$bacteria
SA_21C_Kaiju_bacteria <- res_SA_21C$bacteria
SA_28A_Kaiju_bacteria <- res_SA_28A$bacteria
SA_28B_Kaiju_bacteria <- res_SA_28B$bacteria
SA_28C_Kaiju_bacteria <- res_SA_28C$bacteria


abundance_SA_0A_b_genus_per <- summarize_genus_abundance(SA_0A_Kaiju_bacteria,"SA_0A")
abundance_SA_0B_b_genus_per <- summarize_genus_abundance(SA_0B_Kaiju_bacteria,"SA_0B")
abundance_SA_0C_b_genus_per <- summarize_genus_abundance(SA_0C_Kaiju_bacteria,"SA_0C")
abundance_SA_2A_b_genus_per <- summarize_genus_abundance(SA_2A_Kaiju_bacteria,"SA_2A")
abundance_SA_2B_b_genus_per <- summarize_genus_abundance(SA_2B_Kaiju_bacteria,"SA_2B")
abundance_SA_2C_b_genus_per <- summarize_genus_abundance(SA_2C_Kaiju_bacteria,"SA_2C")
abundance_SA_7A_b_genus_per <- summarize_genus_abundance(SA_7A_Kaiju_bacteria,"SA_7A")
abundance_SA_7B_b_genus_per <- summarize_genus_abundance(SA_7B_Kaiju_bacteria,"SA_7B")
abundance_SA_7C_b_genus_per <- summarize_genus_abundance(SA_7C_Kaiju_bacteria,"SA_7C")
abundance_SA_14A_b_genus_per <- summarize_genus_abundance(SA_14A_Kaiju_bacteria,"SA_14A")
abundance_SA_14B_b_genus_per <- summarize_genus_abundance(SA_14B_Kaiju_bacteria,"SA_14B")
abundance_SA_14C_b_genus_per <- summarize_genus_abundance(SA_14C_Kaiju_bacteria,"SA_14C")
abundance_SA_21A_b_genus_per <- summarize_genus_abundance(SA_21A_Kaiju_bacteria,"SA_21A")
abundance_SA_21B_b_genus_per <- summarize_genus_abundance(SA_21B_Kaiju_bacteria,"SA_21B")
abundance_SA_21C_b_genus_per <- summarize_genus_abundance(SA_21C_Kaiju_bacteria,"SA_21C")
abundance_SA_28A_b_genus_per <- summarize_genus_abundance(SA_28A_Kaiju_bacteria,"SA_28A")
abundance_SA_28B_b_genus_per <- summarize_genus_abundance(SA_28B_Kaiju_bacteria,"SA_28B")
abundance_SA_28C_b_genus_per <- summarize_genus_abundance(SA_28C_Kaiju_bacteria,"SA_28C")

abundance_kaiju_bacteria_cab <- full_join(abundance_SA_0A_b_genus_per,abundance_SA_0B_b_genus_per)%>%
  full_join(.,abundance_SA_0C_b_genus_per)%>%
  full_join(.,abundance_SA_2A_b_genus_per)%>%
  full_join(.,abundance_SA_2B_b_genus_per)%>%
  full_join(.,abundance_SA_2C_b_genus_per)%>%
  full_join(.,abundance_SA_7A_b_genus_per)%>%
  full_join(.,abundance_SA_7B_b_genus_per)%>%
  full_join(.,abundance_SA_7C_b_genus_per)%>%
  full_join(.,abundance_SA_14A_b_genus_per)%>%
  full_join(.,abundance_SA_14B_b_genus_per)%>%
  full_join(.,abundance_SA_14C_b_genus_per)%>%
  full_join(.,abundance_SA_21A_b_genus_per)%>%
  full_join(.,abundance_SA_21B_b_genus_per)%>%
  full_join(.,abundance_SA_21C_b_genus_per)%>%
  full_join(.,abundance_SA_28A_b_genus_per)%>%
  full_join(.,abundance_SA_28B_b_genus_per)%>%
  full_join(.,abundance_SA_28C_b_genus_per)


abundance_kaiju_bacteria_cab [is.na(abundance_kaiju_bacteria_cab)] <- 0

abundance_kaiju_bacteria_cab_2 <- column_to_rownames(abundance_kaiju_bacteria_cab,"genus")%>%
  mutate(sum=rowSums(.)) %>%
  arrange(-sum) %>%
  rownames_to_column(var ="genus")

abundance_kaiju_bacteria_cab_3 <- mutate(abundance_kaiju_bacteria_cab_2,genus = ifelse(sum < 5, "Other", genus)) %>%
  group_by(genus)

abundance_kaiju_bacteria_cab_4 <- abundance_kaiju_bacteria_cab_3[c("genus","SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C")]

abundance_long_bacteria_cabbage <- abundance_kaiju_bacteria_cab_4 %>%
  pivot_longer(cols = -genus, names_to = "Samples", values_to = "abundance")

#print(abundance_kaiju_bacteria_4$genus)

abundance_long_bacteria_cabbage$genus <- factor(abundance_long_bacteria_cabbage$genus, levels = c("Leuconostoc","Secundilactobacillus","Lactiplantibacillus","Levilactobacillus","Lactobacillus","Pantoea","Pediococcus","Rahnella","Weissella","Enterobacter","Other"))

abundance_long_bacteria_cabbage$Samples <- factor(abundance_long_bacteria_cabbage$Samples, levels = c("SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C"))


abundance_long_bacteria_cabbage2 <- left_join(abundance_long_bacteria_cabbage,Fermentation_metadata, by = c("Samples" = "Samples_id"))



genus_col <- c("Leuconostoc" = '#a6cee3' ,"Secundilactobacillus"= "#1f78b4" ,"Lactiplantibacillus" = '#b2df8a',"Levilactobacillus" = '#fb9a99',"Lactobacillus"= '#e31a1c',"Pantoea"= '#cab2d6',"Pediococcus"=  '#5e4fa2',"Rahnella"='#ff7f00' ,"Weissella" ='#de77ae',"Enterobacter" ='#8e0152',"Other" ='#d9d9d9')

per_read_genus_bact_cabb <- ggplot(abundance_long_bacteria_cabbage2, aes(fill=genus, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Genus")+
  scale_fill_manual(values = genus_col)+
  theme(legend.text = element_text(size = 11,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 11))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_genus_bact_cabb

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/",Sys.Date(),"_Figure_4A.svg")
ggsave(file_name,plot =per_read_genus_bact_cabb,device = "svg",width = 10, height = 8, dpi = 300)

```

#### Statistic analyses of the bacterial community variation during cabbage fermentation

```{r Cabbage Bray-curtis/permanova, message=FALSE, warning=FALSE}
abundance_kaiju_bacteria_cabbage_obj <-abundance_kaiju_bacteria_cab%>%
  column_to_rownames(.,"genus")

Fermentation_metadata_obj <- Fermentation_metadata
Fermentation_metadata_obj$Figure_label <- Fermentation_metadata$Samples_id
Fermentation_metadata_obj<-column_to_rownames(Fermentation_metadata_obj,"Samples_id")

obj_Cabbage_metag_bact <- phyloseq(otu_table(abundance_kaiju_bacteria_cabbage_obj, taxa_are_rows = TRUE), sample_data(Fermentation_metadata_obj))
obj_Cabbage_metag_bact

dist_bc_metag_bact_cab <- phyloseq::distance(obj_Cabbage_metag_bact, method = "bray") 
#dist_jcc_metag_bact <- phyloseq::distance(obj_Carrot_metag_bact, method = "jaccard")

# on BC 
set.seed(030623)
adonis_bc_sample_cab <- adonis2(formula = dist_bc_metag_bact_cab~Day_sampling_text, data = data.frame(sample_data(obj_Cabbage_metag_bact)))
adonis_bc_sample_cab
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample_cab$R2[1],3), ", p =", adonis_bc_sample_cab$`Pr(>F)`[1],"**") 



pal_lib <- c("Day 0"="#3D5941FF","Day 2"="#778868FF","Day 7"="#B5B991FF", "Day 14"="#DE8A5AFF","Day 21"="#BE7245FF","Day 28"="#8C311CFF")

# Si vous voulez ordonner sampling_day selon un ordre spécifique
#obj_Carrot_metag_bact@sam_data$sampling_day <- factor(obj_Cabbage_metag_bact@sam_data$sampling_day, 
                                                 #levels = c("0", "2", "7","14","21","28"))  # Spécifiez votre ordre ici

#PCoA with BC
ord_PCoA_bc <- ordinate(obj_Cabbage_metag_bact, method = "PCoA", distance = dist_bc_metag_bact_cab)

p1_metag_bact_cabb_BC <- plot_ordination(obj_Cabbage_metag_bact, ord_PCoA_bc,color = "Day_sampling_text")+
  labs(title = "BRAY-CURTIS",
    subtitle = res.adonis.bc_sample,
    color="Sampling day")+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 10),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 12),  # Taille du texte de la légende
    legend.title = element_text(size = 14))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib)
  
p1_metag_bact_cabb_BC


#file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_cabbage_PCoA_BC_.tiff")
#ggsave(file_name,plot =p1_metag_bact_BC,device = "tiff",width = 10, height = 6, dpi = 300)

```

#### Statistics of results

```{r Cabbage stat, message=FALSE, warning=FALSE}

abundance_kaiju_cabbage <- abundance_kaiju_cabbage %>%
  mutate(Day = case_when(
    grepl("0[A-C]$", Samples) ~ 0,
    grepl("2[A-C]$", Samples) ~ 2,
    grepl("7[A-C]$", Samples) ~ 7,
    grepl("14[A-C]$", Samples) ~ 14,
    grepl("21[A-C]$", Samples) ~ 21,
    grepl("28[A-C]$", Samples) ~ 28,
    TRUE ~ NA_real_  # Valeur par défaut si aucun motif ne correspond
  ))


# Calcul du total des reads par jour et par superkingdom
reads_summary_cabb <- abundance_kaiju_cabbage %>%
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads))  # Calcul des proportions


reads_summary_cabb_without_unclassifies <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "Unclassified") %>%  # Suppression des "Unclassified"
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads)) 




## Bacteria stat
reads_summary_cabb_bacteria <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    bacteria_reads = sum(reads[superkingdom == "Bacteria"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_bacteria = bacteria_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_cabb_bacteria)


## Eukaryota stat
reads_summary_cabb_euka <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    euka_reads = sum(reads[superkingdom == "Eukaryota"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_euka = euka_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_cabb_euka)

## Viruses stat
reads_summary_cabb_Viruses <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    Viruses_reads = sum(reads[superkingdom == "Viruses"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_Viruses = Viruses_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_cabb_Viruses)



## Archaea stat
reads_summary_cabb_Archaea <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    Archaea_reads = sum(reads[superkingdom == "Archaea"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_Archaea = Archaea_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_cabb_Archaea)

```

## Analyses of the carrot fermentation

## Import kaiju results from carrot fermentation

```{r Import data carrot, message=FALSE, warning=FALSE}

CA_0A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_0A_libLAH8297_summary_viruses.tsv",sep="\t",header = TRUE)
CA_0B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_0B_libLAH8298_summary_viruses.tsv",sep="\t",header = TRUE)
CA_0C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_0C_libLAH8299_summary_viruses.tsv",sep="\t",header = TRUE)

CA_2A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_2A_libLAH8300_summary_viruses.tsv", sep="\t",header = TRUE)
CA_2B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_2B_libLAH8301_summary_viruses.tsv", sep="\t",header = TRUE)
CA_2C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_2C_libLAH8302_summary_viruses.tsv", sep="\t",header = TRUE)

CA_7A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_7A_libLAH8303_summary_viruses.tsv", sep="\t",header = TRUE)
CA_7B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_7B_libLAH8304_summary_viruses.tsv", sep="\t",header = TRUE)
CA_7C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_7C_libLAH8305_summary_viruses.tsv", sep="\t",header = TRUE)

CA_14A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_14A_libLAH8306_summary_viruses.tsv", sep="\t",header = TRUE)
CA_14B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_14B_libLAH8307_summary_viruses.tsv", sep="\t",header = TRUE)
CA_14C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_14C_libLAH8308_summary_viruses.tsv", sep="\t",header = TRUE)

CA_21A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_21A_libLAH8309_summary_viruses.tsv", sep="\t",header = TRUE)
CA_21B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_21B_libLAH8310_summary_viruses.tsv", sep="\t",header = TRUE)
CA_21C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_21C_libLAH8311_summary_viruses.tsv", sep="\t",header = TRUE)

CA_28A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_28A_libLAH8312_summary_viruses.tsv", sep="\t",header = TRUE)
CA_28B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_28B_libLAH8313_summary_viruses.tsv", sep="\t",header = TRUE)
CA_28C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_28C_libLAH8314_libLAI5154_summary_viruses.tsv", sep="\t",header = TRUE)

res_CA_0A <- process_kaiju_data(CA_0A_Kaiju, "CA_0A")
res_CA_0B <- process_kaiju_data(CA_0B_Kaiju, "CA_0B")
res_CA_0C <- process_kaiju_data(CA_0C_Kaiju, "CA_0C")
res_CA_2A <- process_kaiju_data(CA_2A_Kaiju, "CA_2A")
res_CA_2B <- process_kaiju_data(CA_2B_Kaiju, "CA_2B")
res_CA_2C <- process_kaiju_data(CA_2C_Kaiju, "CA_2C")
res_CA_7A <- process_kaiju_data(CA_7A_Kaiju, "CA_7A")
res_CA_7B <- process_kaiju_data(CA_7B_Kaiju, "CA_7B")
res_CA_7C <- process_kaiju_data(CA_7C_Kaiju, "CA_7C")
res_CA_14A <- process_kaiju_data(CA_14A_Kaiju, "CA_14A")
res_CA_14B <- process_kaiju_data(CA_14B_Kaiju, "CA_14B")
res_CA_14C <- process_kaiju_data(CA_14C_Kaiju, "CA_14C")
res_CA_21A <- process_kaiju_data(CA_21A_Kaiju, "CA_21A")
res_CA_21B <- process_kaiju_data(CA_21B_Kaiju, "CA_21B")
res_CA_21C <- process_kaiju_data(CA_21C_Kaiju, "CA_21C")
res_CA_28A <- process_kaiju_data(CA_28A_Kaiju, "CA_28A")
res_CA_28B <- process_kaiju_data(CA_28B_Kaiju, "CA_28B")
res_CA_28C <- process_kaiju_data(CA_28C_Kaiju, "CA_28C")

CA_0A_Kaiju_process <- res_CA_0A$processed_data
CA_0B_Kaiju_process <- res_CA_0B$processed_data
CA_0C_Kaiju_process <- res_CA_0C$processed_data
CA_2A_Kaiju_process <- res_CA_2A$processed_data
CA_2B_Kaiju_process <- res_CA_2B$processed_data
CA_2C_Kaiju_process <- res_CA_2C$processed_data
CA_7A_Kaiju_process <- res_CA_7A$processed_data
CA_7B_Kaiju_process <- res_CA_7B$processed_data
CA_7C_Kaiju_process <- res_CA_7C$processed_data
CA_14A_Kaiju_process <- res_CA_14A$processed_data
CA_14B_Kaiju_process <- res_CA_14B$processed_data
CA_14C_Kaiju_process <- res_CA_14C$processed_data
CA_21A_Kaiju_process <- res_CA_21A$processed_data
CA_21B_Kaiju_process <- res_CA_21B$processed_data
CA_21C_Kaiju_process <- res_CA_21C$processed_data
CA_28A_Kaiju_process <- res_CA_28A$processed_data
CA_28B_Kaiju_process <- res_CA_28B$processed_data
CA_28C_Kaiju_process <- res_CA_28C$processed_data

abundance_kaiju_carrot <- full_join (CA_0A_Kaiju_process,CA_0B_Kaiju_process)%>%
  full_join(.,CA_0C_Kaiju_process)%>%
  full_join(.,CA_2A_Kaiju_process)%>%
  full_join(.,CA_2B_Kaiju_process)%>%
  full_join(.,CA_2C_Kaiju_process)%>%
  full_join(.,CA_7A_Kaiju_process)%>%
  full_join(.,CA_7B_Kaiju_process)%>%
  full_join(.,CA_7C_Kaiju_process)%>%
  full_join(.,CA_14A_Kaiju_process)%>%
  full_join(.,CA_14B_Kaiju_process)%>%
  full_join(.,CA_14C_Kaiju_process)%>%
  full_join(.,CA_21A_Kaiju_process)%>%
  full_join(.,CA_21B_Kaiju_process)%>%
  full_join(.,CA_21C_Kaiju_process)%>%
  full_join(.,CA_28A_Kaiju_process)%>%
  full_join(.,CA_28B_Kaiju_process)%>%
  full_join(.,CA_28C_Kaiju_process)

write.table(as.data.frame(abundance_kaiju_carrot), file = "Carrot_read_kaiju.txt", sep = "\t", row.names = T, quote = F, col.names = NA)



options(scipen = 999)
mean_proportion_df <- abundance_kaiju_carrot %>%
  group_by(superkingdom) %>%
  summarise(
    mean_reads = mean(reads, na.rm = TRUE),
    sum_reads = sum(reads),
    .groups = 'drop'
  ) %>%
  mutate(
    proportion = (mean_reads / sum(mean_reads))*100  # Calculer la proportion des moyennes => ligne de code fausse
  )

mean_proportion_df

```

### Microbial community at the domain level

```{r carrot kingdom plot, message=FALSE, warning=FALSE}
abundance_CA_0A_kingdom <- abundance_summary_ki(CA_0A_Kaiju_process, superkingdom, percent, "CA_0A")
abundance_CA_0B_kingdom <- abundance_summary_ki(CA_0B_Kaiju_process, superkingdom, percent, "CA_0B")
abundance_CA_0C_kingdom <- abundance_summary_ki(CA_0C_Kaiju_process, superkingdom, percent, "CA_0C")
abundance_CA_2A_kingdom <- abundance_summary_ki(CA_2A_Kaiju_process, superkingdom, percent, "CA_2A")
abundance_CA_2B_kingdom <- abundance_summary_ki(CA_2B_Kaiju_process, superkingdom, percent, "CA_2B")
abundance_CA_2C_kingdom <- abundance_summary_ki(CA_2C_Kaiju_process, superkingdom, percent, "CA_2C")
abundance_CA_7A_kingdom <- abundance_summary_ki(CA_7A_Kaiju_process, superkingdom, percent, "CA_7A")
abundance_CA_7B_kingdom <- abundance_summary_ki(CA_7B_Kaiju_process, superkingdom, percent, "CA_7B")
abundance_CA_7C_kingdom <- abundance_summary_ki(CA_7C_Kaiju_process, superkingdom, percent, "CA_7C")
abundance_CA_14A_kingdom <- abundance_summary_ki(CA_14A_Kaiju_process, superkingdom, percent, "CA_14A")
abundance_CA_14B_kingdom <- abundance_summary_ki(CA_14B_Kaiju_process, superkingdom, percent, "CA_14B")
abundance_CA_14C_kingdom <- abundance_summary_ki(CA_14C_Kaiju_process, superkingdom, percent, "CA_14C")
abundance_CA_21A_kingdom <- abundance_summary_ki(CA_21A_Kaiju_process, superkingdom, percent, "CA_21A")
abundance_CA_21B_kingdom <- abundance_summary_ki(CA_21B_Kaiju_process, superkingdom, percent, "CA_21B")
abundance_CA_21C_kingdom <- abundance_summary_ki(CA_21C_Kaiju_process, superkingdom, percent, "CA_21C")
abundance_CA_28A_kingdom <- abundance_summary_ki(CA_28A_Kaiju_process, superkingdom, percent, "CA_28A")
abundance_CA_28B_kingdom <- abundance_summary_ki(CA_28B_Kaiju_process, superkingdom, percent, "CA_28B")
abundance_CA_28C_kingdom <- abundance_summary_ki(CA_28C_Kaiju_process, superkingdom, percent, "CA_28C")

abundance_kaiju_kingdom <- full_join(abundance_CA_0A_kingdom,abundance_CA_0B_kingdom)%>%
  full_join(.,abundance_CA_0C_kingdom)%>%
  full_join(.,abundance_CA_2A_kingdom)%>%
  full_join(.,abundance_CA_2B_kingdom)%>%
  full_join(.,abundance_CA_2C_kingdom)%>%
  full_join(.,abundance_CA_7A_kingdom)%>%
  full_join(.,abundance_CA_7B_kingdom)%>%
  full_join(.,abundance_CA_7C_kingdom)%>%
  full_join(.,abundance_CA_14A_kingdom)%>%
  full_join(.,abundance_CA_14B_kingdom)%>%
  full_join(.,abundance_CA_14C_kingdom)%>%
  full_join(.,abundance_CA_21A_kingdom)%>%
  full_join(.,abundance_CA_21B_kingdom)%>%
  full_join(.,abundance_CA_21C_kingdom)%>%
  full_join(.,abundance_CA_28A_kingdom)%>%
  full_join(.,abundance_CA_28B_kingdom)%>%
  full_join(.,abundance_CA_28C_kingdom)

# Convertir en format long pour ggplot2
abundance_long_carrot <- abundance_kaiju_kingdom %>%
  pivot_longer(cols = -superkingdom, names_to = "Samples", values_to = "abundance")

abundance_long_carrot$superkingdom <- factor(abundance_long_carrot$superkingdom , levels = c("Archaea","Bacteria","Eukaryota","Viruses","unclassified"))
kingdom_col <- c("Archaea"='#d7191c',"Bacteria"='#fdae61',"Eukaryota"='#ffffbf',"unclassified"='#abd9e9', "Viruses"='#2c7bb6' )

abundance_long_carrot2 <- left_join(abundance_long_carrot,Fermentation_metadata, by = c("Samples" = "Samples_id"))


per_read_kingdom_carrot <- ggplot(abundance_long_carrot2, aes(fill=superkingdom, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Kingdom")+
  scale_fill_manual(values = kingdom_col)+
  theme(legend.text = element_text(size = 12),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 12))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
  
per_read_kingdom_carrot

```

### Bacterial community at the genus level

```{r carrot bacteria plot, message=FALSE, warning=FALSE}

CA_0A_Kaiju_bacteria <- res_CA_0A$bacteria
CA_0B_Kaiju_bacteria <- res_CA_0B$bacteria
CA_0C_Kaiju_bacteria <- res_CA_0C$bacteria
CA_2A_Kaiju_bacteria <- res_CA_2A$bacteria
CA_2B_Kaiju_bacteria <- res_CA_2B$bacteria
CA_2C_Kaiju_bacteria <- res_CA_2C$bacteria
CA_7A_Kaiju_bacteria <- res_CA_7A$bacteria
CA_7B_Kaiju_bacteria <- res_CA_7B$bacteria
CA_7C_Kaiju_bacteria <- res_CA_7C$bacteria
CA_14A_Kaiju_bacteria <- res_CA_14A$bacteria
CA_14B_Kaiju_bacteria <- res_CA_14B$bacteria
CA_14C_Kaiju_bacteria <- res_CA_14C$bacteria
CA_21A_Kaiju_bacteria <- res_CA_21A$bacteria
CA_21B_Kaiju_bacteria <- res_CA_21B$bacteria
CA_21C_Kaiju_bacteria <- res_CA_21C$bacteria
CA_28A_Kaiju_bacteria <- res_CA_28A$bacteria
CA_28B_Kaiju_bacteria <- res_CA_28B$bacteria
CA_28C_Kaiju_bacteria <- res_CA_28C$bacteria


abundance_CA_0A_b_genus_per <- summarize_genus_abundance(CA_0A_Kaiju_bacteria,"CA_0A")
abundance_CA_0B_b_genus_per <- summarize_genus_abundance(CA_0B_Kaiju_bacteria,"CA_0B")
abundance_CA_0C_b_genus_per <- summarize_genus_abundance(CA_0C_Kaiju_bacteria,"CA_0C")
abundance_CA_2A_b_genus_per <- summarize_genus_abundance(CA_2A_Kaiju_bacteria,"CA_2A")
abundance_CA_2B_b_genus_per <- summarize_genus_abundance(CA_2B_Kaiju_bacteria,"CA_2B")
abundance_CA_2C_b_genus_per <- summarize_genus_abundance(CA_2C_Kaiju_bacteria,"CA_2C")
abundance_CA_7A_b_genus_per <- summarize_genus_abundance(CA_7A_Kaiju_bacteria,"CA_7A")
abundance_CA_7B_b_genus_per <- summarize_genus_abundance(CA_7B_Kaiju_bacteria,"CA_7B")
abundance_CA_7C_b_genus_per <- summarize_genus_abundance(CA_7C_Kaiju_bacteria,"CA_7C")
abundance_CA_14A_b_genus_per <- summarize_genus_abundance(CA_14A_Kaiju_bacteria,"CA_14A")
abundance_CA_14B_b_genus_per <- summarize_genus_abundance(CA_14B_Kaiju_bacteria,"CA_14B")
abundance_CA_14C_b_genus_per <- summarize_genus_abundance(CA_14C_Kaiju_bacteria,"CA_14C")
abundance_CA_21A_b_genus_per <- summarize_genus_abundance(CA_21A_Kaiju_bacteria,"CA_21A")
abundance_CA_21B_b_genus_per <- summarize_genus_abundance(CA_21B_Kaiju_bacteria,"CA_21B")
abundance_CA_21C_b_genus_per <- summarize_genus_abundance(CA_21C_Kaiju_bacteria,"CA_21C")
abundance_CA_28A_b_genus_per <- summarize_genus_abundance(CA_28A_Kaiju_bacteria,"CA_28A")
abundance_CA_28B_b_genus_per <- summarize_genus_abundance(CA_28B_Kaiju_bacteria,"CA_28B")
abundance_CA_28C_b_genus_per <- summarize_genus_abundance(CA_28C_Kaiju_bacteria,"CA_28C")

abundance_kaiju_bacteria_carrot <- full_join(abundance_CA_0A_b_genus_per,abundance_CA_0B_b_genus_per)%>%
  full_join(.,abundance_CA_0C_b_genus_per)%>%
  full_join(.,abundance_CA_2A_b_genus_per)%>%
  full_join(.,abundance_CA_2B_b_genus_per)%>%
  full_join(.,abundance_CA_2C_b_genus_per)%>%
  full_join(.,abundance_CA_7A_b_genus_per)%>%
  full_join(.,abundance_CA_7B_b_genus_per)%>%
  full_join(.,abundance_CA_7C_b_genus_per)%>%
  full_join(.,abundance_CA_14A_b_genus_per)%>%
  full_join(.,abundance_CA_14B_b_genus_per)%>%
  full_join(.,abundance_CA_14C_b_genus_per)%>%
  full_join(.,abundance_CA_21A_b_genus_per)%>%
  full_join(.,abundance_CA_21B_b_genus_per)%>%
  full_join(.,abundance_CA_21C_b_genus_per)%>%
  full_join(.,abundance_CA_28A_b_genus_per)%>%
  full_join(.,abundance_CA_28B_b_genus_per)%>%
  full_join(.,abundance_CA_28C_b_genus_per)


abundance_kaiju_bacteria_carrot [is.na(abundance_kaiju_bacteria_carrot)] <- 0

abundance_kaiju_bacteria_carrot2 <- column_to_rownames(abundance_kaiju_bacteria_carrot,"genus")%>%
  mutate(sum=rowSums(.)) %>%
  arrange(-sum) %>%
  rownames_to_column(var ="genus")

abundance_kaiju_bacteria_carrot3 <- mutate(abundance_kaiju_bacteria_carrot2,genus = ifelse(sum < 5, "Other", genus)) %>%
  group_by(genus)

abundance_kaiju_bacteria_carrot4 <- abundance_kaiju_bacteria_carrot3[c("genus","CA_0A","CA_0B","CA_0C","CA_2A","CA_2B","CA_2C","CA_7A","CA_7B","CA_7C","CA_14A","CA_14B","CA_14C","CA_21A","CA_21B","CA_21C", "CA_28A", "CA_28B", "CA_28C")]

abundance_long_bacteria_cabbage2 <- left_join(abundance_long_bacteria_cabbage,Fermentation_metadata, by = c("Samples" = "Samples_id"))


abundance_long_bacteria_carrot <- abundance_kaiju_bacteria_carrot4 %>%
  pivot_longer(cols = -genus, names_to = "Samples", values_to = "abundance")

print(abundance_kaiju_bacteria_carrot2$genus)

abundance_long_bacteria_carrot$genus <- factor(abundance_long_bacteria_carrot$genus, levels = c("Leuconostoc","Lactiplantibacillus","Levilactobacillus","Weissella","Other"))

abundance_long_bacteria_carrot$Samples <- factor(abundance_long_bacteria_carrot$Samples, levels = c("CA_0A","CA_0B","CA_0C","CA_2A","CA_2B","CA_2C","CA_7A","CA_7B","CA_7C","CA_14A","CA_14B","CA_14C","CA_21A","CA_21B","CA_21C", "CA_28A", "CA_28B", "CA_28C"))

genus_col <- c("Leuconostoc" = '#a6cee3',"Lactiplantibacillus" = '#b2df8a',"Levilactobacillus" = '#fb9a99',"Weissella" ='#de77ae',"Other" ='#d9d9d9')

abundance_long_bacteria_carrot2 <- left_join(abundance_long_bacteria_carrot,Fermentation_metadata, by = c("Samples" = "Samples_id"))

per_read_genus_bact_carr <- ggplot(abundance_long_bacteria_carrot2, aes(fill=genus, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Genus")+
  scale_fill_manual(values = genus_col)+
  theme(legend.text = element_text(size = 12,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 12))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_genus_bact_carr

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/",Sys.Date(),"_Figure_5A.svg")
ggsave(file_name,plot =per_read_genus_bact_carr,device = "svg",width = 10, height = 8, dpi = 300)

#file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_bacteria_carrot.tiff")
#ggsave(file_name,plot =per_read_genus_bact_carr,device = "tiff",width = 15, height = 6, dpi = 300)
```

#### Statistic analyses of the bacterial community variation during carrot fermentation

```{r Carrot Bray-curtis/permanova, message=FALSE, warning=FALSE}
abundance_kaiju_bacteria_carrot_obj <-abundance_kaiju_bacteria_carrot%>%
  column_to_rownames(.,"genus")

Fermentation_metadata_obj <- Fermentation_metadata
Fermentation_metadata_obj$Figure_label <- Fermentation_metadata$Samples_id
Fermentation_metadata_obj<-column_to_rownames(Fermentation_metadata_obj,"Samples_id")

obj_Carrot_metag_bact <- phyloseq(otu_table(abundance_kaiju_bacteria_carrot_obj, taxa_are_rows = TRUE), sample_data(Fermentation_metadata_obj))
obj_Carrot_metag_bact

dist_bc_metag_bact <- phyloseq::distance(obj_Carrot_metag_bact, method = "bray") 
#dist_jcc_metag_bact <- phyloseq::distance(obj_Carrot_metag_bact, method = "jaccard")

# on BC 
set.seed(030623)
adonis_bc_sample <- adonis2(formula = dist_bc_metag_bact~Day_sampling_text, data = data.frame(sample_data(obj_Carrot_metag_bact)))
adonis_bc_sample
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample$R2[1],3), ", p =", adonis_bc_sample$`Pr(>F)`[1],"**") 



pal_lib <- c("Day 0"="#3D5941FF","Day 2"="#778868FF","Day 7"="#B5B991FF", "Day 14"="#DE8A5AFF","Day 21"="#BE7245FF","Day 28"="#8C311CFF")

# Si vous voulez ordonner sampling_day selon un ordre spécifique
#obj_Carrot_metag_bact@sam_data$sampling_day <- factor(obj_Carrot_metag_bact@sam_data$sampling_day, 
                                                 #levels = c("0", "2", "7","14","21","28"))  # Spécifiez votre ordre ici

#PCoA with BC
ord_PCoA_bc <- ordinate(obj_Carrot_metag_bact, method = "PCoA", distance = dist_bc_metag_bact)

p1_metag_bact_BC <- plot_ordination(obj_Carrot_metag_bact, ord_PCoA_bc,color = "Day_sampling_text")+
  labs(title = "BRAY-CURTIS",
    subtitle = res.adonis.bc_sample,
    color="Sampling day")+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 10),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 12),  # Taille du texte de la légende
    legend.title = element_text(size = 14))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib)
  
p1_metag_bact_BC


#file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_carrot_PCoA_BC_.tiff")
#ggsave(file_name,plot =p1_metag_bact_BC,device = "tiff",width = 10, height = 6, dpi = 300)

```

#### Statistics of results

```{r Carrot stat, message=FALSE, warning=FALSE}
abundance_kaiju_carrot
abundance_kaiju_carrot <- abundance_kaiju_carrot %>%
  mutate(Day = case_when(
    grepl("0[A-C]$", Samples) ~ 0,
    grepl("2[A-C]$", Samples) ~ 2,
    grepl("7[A-C]$", Samples) ~ 7,
    grepl("14[A-C]$", Samples) ~ 14,
    grepl("21[A-C]$", Samples) ~ 21,
    grepl("28[A-C]$", Samples) ~ 28,
    TRUE ~ NA_real_  # Valeur par défaut si aucun motif ne correspond
  ))


# Calcul du total des reads par jour et par superkingdom
reads_summary_carr <- abundance_kaiju_carrot %>%
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads))  # Calcul des proportions


reads_summary_carr_without_unclassifies <- abundance_kaiju_carrot %>%
  filter(superkingdom != "Unclassified") %>%  # Suppression des "Unclassified"
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads)) 



## Bacteria stat
reads_summary_carr_bacteria <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    bacteria_reads = sum(reads[superkingdom == "Bacteria"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_bacteria = bacteria_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_carr_bacteria)


## Eukaryota stat
reads_summary_carr_euka <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    euka_reads = sum(reads[superkingdom == "Eukaryota"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_euka = euka_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_carr_euka)

## Viruses stat
reads_summary_carr_Viruses <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    Viruses_reads = sum(reads[superkingdom == "Viruses"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_Viruses = Viruses_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_carr_Viruses)



## Archaea stat
reads_summary_carr_Archaea <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>%  # Supprime les "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total des reads classifiés (sans Unclassified)
    Archaea_reads = sum(reads[superkingdom == "Archaea"], na.rm = TRUE)  # Reads des bactéries uniquement
  ) %>%
  mutate(proportion_Archaea = Archaea_reads / total_reads_classified)  # Proportion des bactéries sur les reads classifiés

# Affichage du résultat
print(reads_summary_carr_Archaea)

```

```{r figure article, message=FALSE, warning=FALSE}

# Colors
genus_col <- c(
  "Leuconostoc" = "#a6cee3", "Secundilactobacillus" = "#1f78b4",
  "Lactiplantibacillus" = "#b2df8a", "Levilactobacillus" = "#fb9a99",
  "Lactobacillus" = "#e31a1c", "Pantoea" = "#cab2d6",
  "Pediococcus" = "#5e4fa2", "Rahnella" = "#ff7f00",
  "Weissella" = "#de77ae", "Enterobacter" = "#8e0152",
  "Other" = "#d9d9d9"
)

#  Function: Average abundance by genus for a given table
mean_by_genus <- function(df, name) {
  df %>% transmute(genus, !!name := rowMeans(pick(where(is.numeric))))
}

abundance_long_bacteria_genus2 <- 
  full_join(
    mean_by_genus(abundance_kaiju_bacteria_cab,    "Cabbage"),
    mean_by_genus(abundance_kaiju_bacteria_carrot, "Carrot"),
    by = "genus"
  ) %>%
  replace_na(list(Cabbage = 0, Carrot = 0)) %>%
  mutate(genus = if_else(Cabbage + Carrot < 0.4, "Other", genus)) %>%
  pivot_longer(-genus, names_to = "Samples", values_to = "abundance") %>%
  group_by(genus, Samples) %>%
  summarise(Abondance_totale = sum(abundance, na.rm = TRUE), .groups = "drop") %>%
  mutate(genus = factor(genus, levels = names(genus_col)))

per_bacteria_genus <- ggplot(abundance_long_bacteria_genus2,
                             aes(x = Samples, y = Abondance_totale, fill = genus)) +
  geom_col() +
  scale_fill_manual(values = genus_col) +
  labs(x = "", y = "Relative abundance (%)", fill = "Genus") +
  theme(
    legend.text  = element_text(size = 12, face = "italic"),
    legend.title = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text    = element_text(size = 14),
    strip.text   = element_text(size = 14),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank()
  )
per_bacteria_genus

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/",Sys.Date(),"_Figure_3A.svg")
ggsave(file_name,plot =per_bacteria_genus,device = "svg",width = 10, height = 10, dpi = 300)

```
