---
title: "Metavirome_analyses"
author: "Julia Gendre"
date: "2024-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages loading

```{r packages, message=FALSE, warning=FALSE}
library(phyloseq)
library(readxl)  
library(phyloseq.extended)
library(vegan)
library(gridExtra)
library(rstatix)
library(gplots)
library(tidyverse)
library(DT)
library(pheatmap)
library(DESeq2)
library(ggplotify)
library(cowplot)
library(ggpubr)
library(ggforce)
library(ggplot2)
library(ggrepel)
library(ggbreak)
library(VennDiagram)
library(grid)
library(factoextra)
#library(pairwiseAdonis)
library(Hmisc)
library(dplyr)
library(multcomp)
library(RColorBrewer)
sessionInfo()

```

## Data import
Download viral metagenomics processed data (https://doi.org/10.57745/3EHB1F) from the National Research Data Platform (https://entrepot.recherche.data.gouv.fr/).


```{r Import data, message=FALSE, warning=FALSE}
setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

#Import count matrix of all samples
count <- read.table("results/presence_counts.csv") 
count <- count[-1:-2,] #delete 2 firsts line that don't corresponding to contigs.

sum_col_count<- colSums(count)

sample_metadata <- read.csv("METADATA/samples_metadata.txt", sep = "\t", dec="," ,header = TRUE,
                                    fileEncoding = "latin1")

contigs_info1 <- read.csv("METADATA/vOTU_metadata.txt", sep = "\t", dec="." ,header = TRUE,
                                    fileEncoding = "latin1")

sample_metadata<-  column_to_rownames(sample_metadata,var = "Sample_short_id")

contigs_info<-  column_to_rownames(contigs_info1,var = "vOTUs_ID")%>%
  as.matrix() # transform the count abundance table into matrix.

contigs_info
colSums(is.na(contigs_info)) 
```

## Removal of the contigs with high abundance only in control

Control water and buffer samples have been removed from the abundance table. Additionally, contigs primarily found in these control samples will be removed under the following conditions:

1.  Present only in water and buffer samples.

2.  Present in high abundance in water and buffer samples but in low abundance in other samples.

```{r control cleaning, message=FALSE, warning=FALSE}

bad_taxa = c("T_V_1000k_NODE_13_length_23093_cov_15.072107",
             "T_V_1000k_NODE_56_length_13019_cov_7.392491",
              "T_V_1000k_NODE_9_length_24886_cov_6.334182",
              "T_V_NODE_7_length_34739_cov_78.425344",
              "T_V_NODE_4_length_48176_cov_31.113759",
              "T_V_NODE_104_length_14793_cov_104.961271",
              "T_V_NODE_13_length_29414_cov_39.305391",
              "T_V_1000k_NODE_12_length_23419_cov_3.415679",
              "W1_V_500k_NODE_21_length_17263_cov_3.932773",
              "CA_0B_V_NODE_56_length_4940_cov_1.522543",
              "T_V_500k_NODE_24_length_14406_cov_1.990335",
              "T_V_NODE_112_length_13957_cov_84.059870",
              "T_V_NODE_52_length_20374_cov_58.245172",
              "T_V_100k_NODE_25_length_5848_cov_0.829226",
              "T_V_1000k_NODE_81_length_11275_cov_2.917743",
              "W1_V_NODE_229_length_5961_cov_33.205348",
              "T_V_NODE_27_length_24720_cov_28.517180",
              "T_V_NODE_209_length_8647_cov_53.598826",
              "T_V_1000k_NODE_30_length_16713_cov_1.892319",
              "T_V_NODE_22_length_25365_cov_26.715152",
              "T_V_NODE_130_length_12660_cov_47.610149",
              "T_V_1000k_NODE_64_length_12619_cov_1.951409",
              "T_V_1000k_NODE_148_length_7765_cov_3.760932",
              "T_V_1000k_NODE_73_length_11855_cov_2.388216",
              "T_V_NODE_115_length_13725_cov_26.154508",
              "T_V_1000k_NODE_218_length_5631_cov_3.524709",
              "T_V_NODE_145_length_11681_cov_39.453782",
              "T_V_1000k_NODE_69_length_12070_cov_1.904044",
              "T_V_NODE_243_length_7869_cov_53.875743",
              "W1_V_NODE_117_length_9118_cov_85.232010",
              "T_V_NODE_86_length_16067_cov_12.726913",
              "T_V_1000k_NODE_154_length_7568_cov_2.658110",
              "T_V_1000k_NODE_100_length_9726_cov_2.064486",
              "W1_V_NODE_112_length_9370_cov_35.870605",
              "W1_V_NODE_96_length_10280_cov_70.804294",
              "W1_V_500k_NODE_164_length_4846_cov_8.287985",
              "T_V_NODE_156_length_10950_cov_19.953248",
              "T_V_NODE_162_length_10579_cov_15.557979",
              "W1_V_NODE_119_length_8957_cov_9.429445",
              "W1_V_NODE_21_length_19574_cov_13.497609",
              "T_V_NODE_200_length_8994_cov_13.492726",
              "T_V_NODE_276_length_6855_cov_20.919441",
              "T_V_NODE_333_length_5442_cov_22.528692",
              "T_V_500k_NODE_214_length_4196_cov_1.369624",
              "T_V_500k_NODE_214_length_4196_cov_1.369624",
              "T_V_1000k_NODE_285_length_4145_cov_0.829766",
              "T_V_NODE_405_length_4508_cov_3.446017",
              "W1_V_NODE_145_length_8297_cov_6.021053",
              "T_V_500k_NODE_443_length_2099_cov_0.447769",
              "T_V_NODE_285_length_6459_cov_5.294378",
              "T_V_NODE_290_length_6306_cov_3.606085",
              "T_V_NODE_338_length_5352_cov_2.890718",
              "T_V_NODE_603_length_2733_cov_2.466999",
              "T_V_NODE_539_length_3206_cov_1.638519",
              "T_V_NODE_630_length_2588_cov_0.894352",
              "W1_V_500k_NODE_19_length_17675_cov_2.110041",
              "W1_V_500k_NODE_108_length_6788_cov_0.705600",
              "W1_V_1000k_NODE_12_length_21902_cov_9.845144",
              "W1_V_1000k_NODE_32_length_16063_cov_9.498117",
              "W1_V_1000k_NODE_56_length_13422_cov_3.322527",
              "W1_V_1000k_NODE_58_length_13234_cov_3.604791",
              "W1_V_1000k_NODE_128_length_7714_cov_5.103598",
              "W1_V_1000k_NODE_136_length_7431_cov_1.195783",
              "W_V_1000k_NODE_12_length_11849_cov_4.785617",
              "W_V_1000k_NODE_33_length_5693_cov_0.906576",
              "W_V_1000k_NODE_35_length_5450_cov_4.528086",
              "W_V_1000k_NODE_56_length_3442_cov_19.229261",
              "W_V_1000k_NODE_84_length_2398_cov_3.914135",
              "W_V_1000k_NODE_102_length_2051_cov_8.119543",
              "SA_28B_V_NODE_293_length_4473_cov_1.719972",
              "W1_V_NODE_1_length_38460_cov_67.594240",
              "W1_V_NODE_10_length_25573_cov_27.776075",
              "W1_V_NODE_45_length_14229_cov_17.107928",
              "W1_V_NODE_81_length_11207_cov_8.631588",
              "W1_V_NODE_98_length_10196_cov_8.714867",
              "W1_V_NODE_107_length_9643_cov_2.562526",
              "W1_V_NODE_129_length_8690_cov_44.821091",
              "W1_V_NODE_147_length_8223_cov_1.733819",
              "W1_V_NODE_203_length_6476_cov_95.062057",
              "W1_V_NODE_445_length_3312_cov_6.226688",
              "W1_V_NODE_554_length_2658_cov_1.602924",
              "W_V_NODE_21_length_8713_cov_11.221989",
              "W_V_NODE_38_length_6710_cov_3.715327",
              "W_V_NODE_61_length_5028_cov_16.932259",
              "W_V_NODE_79_length_4534_cov_18.637622",
              "W_V_NODE_84_length_4337_cov_4.519952",
              "W_V_NODE_111_length_3588_cov_38.746605",
              "W_V_NODE_158_length_2685_cov_0.659891",
              "W_V_NODE_165_length_2601_cov_1.288601",
              "W_V_NODE_176_length_2424_cov_5.028298",
              "W_V_NODE_214_length_2062_cov_16.974418")


# Créez un vecteur avec tous les noms de lignes dans le dataframe d'origine
all_taxa <- rownames(count)

# Trouvez les noms de lignes qui ne sont pas dans bad_taxa
conserved_name <- setdiff(all_taxa,bad_taxa)

# Sélectionnez les lignes du dataframe d'origine en utilisant les noms à conserver
abundance_clean <- count[conserved_name,]

contigs_vOTU <- contigs_info[conserved_name,]

```


## Abundance table

### Abundance table for all samples
We created a Phyloseq object that contains the abundance table and metadata of all samples.
```{r all abundance table, message=FALSE, warning=FALSE}

#Removing control samples from the abundance table
control <- c("T_V", "W1_V","W2_V","W_V") #List of the sample name of control
All_sample <- colnames(abundance_clean) #Sample name of abundance table
conserved_col<- setdiff(All_sample,control) # comparer sample name of control and sample name of abundance table. Keep the difference.
All_abundance <- abundance_clean[,conserved_col] #extract DNA sample without control from abundance table

# Check consistency
all(colnames(All_abundance) %in% rownames(sample_metadata)) ## sample names between metadata and otu table
all(rownames(All_abundance) %in% rownames(contigs_info)) ## taxa names between taxonomic and otu tables


# phyloseq object of DNA sample without control sample
obj_All <- phyloseq(otu_table(All_abundance, taxa_are_rows = TRUE), sample_data(sample_metadata) , tax_table(contigs_info))
obj_All

```

### Abundance table for cabbage samples
We created a Phyloseq object that contains the abundance table and metadata of cabbage samples.
```{r cabbage abundance table, message=FALSE, warning=FALSE}
# Extract only Cabbage samples
Cabbage_metadata <- subset(sample_metadata, sample_metadata$Matrix == 'Cabbage')
Cabbage_sample <- rownames(Cabbage_metadata)
Cabbage_abundance <- abundance_clean[,Cabbage_sample,drop = FALSE]

#Removing contig not found into DNA samples
#Calcul the abundance sum of all contigs 
Cabbage_abundance_colsum <- Cabbage_abundance %>%
  mutate(row_sum = rowSums(pick(matches("^SA_\\d+[ABC]_V$"))))

#selected only contigs with abundance higher than 0, meaning present into at least 1 sample
Cabbage_abundance_clean <- subset(Cabbage_abundance_colsum, Cabbage_abundance_colsum$row_sum > 0)%>%
  .[,-19] #delete the sum column for the rest of the analyses

# Check consistency
all(colnames(Cabbage_abundance_clean) %in% rownames(sample_metadata)) ## sample names between metadata and otu table
all(rownames(Cabbage_abundance_clean) %in% rownames(contigs_info)) ## taxa names between taxonomic and otu tables

# phyloseq object of DNA sample without control sample
obj_Cabbage <- phyloseq(otu_table(Cabbage_abundance_clean, taxa_are_rows = TRUE), sample_data(sample_metadata) , tax_table(contigs_info))
obj_Cabbage
```

### Abundance table for carrot samples
We created a Phyloseq object that contains the abundance table and metadata of carrot samples.
```{r Carrot abundance table, message=FALSE, warning=FALSE}
# Extract only Carrot samples
Carrot_metadata <- subset(sample_metadata, sample_metadata$Matrix == 'Carrot')
Carrot_sample <- rownames(Carrot_metadata)
Carrot_abundance <- abundance_clean[,Carrot_sample,drop = FALSE]

#Removing contig not found into DNA samples
#Calcul the abundance sum of all contigs 
Carrot_abundance_colsum <- Carrot_abundance %>%
  mutate(row_sum = rowSums(pick(matches("^CA_\\d+[ABC]_V$"))))

#selected only contigs with abundance higher than 0, meaning present into at least 1 sample
Carrot_abundance_clean <- subset(Carrot_abundance_colsum, Carrot_abundance_colsum$row_sum > 0)%>%
  .[,-19] #delete the sum column for the rest of the analyses

# Check consistency
all(colnames(Carrot_abundance_clean) %in% rownames(sample_metadata)) ## sample names between metadata and otu table
all(rownames(Carrot_abundance_clean) %in% rownames(contigs_info)) ## taxa names between taxonomic and otu tables

# phyloseq object of DNA sample without control sample
obj_Carrot <- phyloseq(otu_table(Carrot_abundance_clean, taxa_are_rows = TRUE), sample_data(sample_metadata) , tax_table(contigs_info))
obj_Carrot
```

## Statistic part

#### Statistic of vOTU for all samples

```{r statistic on vOTU size for all samples, message=FALSE, warning=FALSE}
#Statistics of vOTUs length about abundance table without control contigs
abundance_clean_info <-rownames_to_column(abundance_clean,var="Contig")
Contigs_info_clean <- left_join(x=abundance_clean_info,y=contigs_info1,by=c("Contig"="vOTUs_ID"))
print("Statistics of vOTUs length about abundance table without control contigs")
summary(Contigs_info_clean$Length)
table(Contigs_info_clean$Type_vibrant)

#Selected contigs with a size higher than  100 kb
contigs_length_stat <- Contigs_info_clean %>%
  arrange(-Length) %>%
  dplyr::filter(Length >= 100000)

# Count of contigs greater than 10kb based on CheckV quality predictions
table(contigs_length_stat$checkv_quality)


# Draw density histogram of the distribution of contig sizes. 
p_dens <- ggplot(contigs_info1, aes(x=(Length/1000))) +
  geom_histogram(colour="black", fill="#a6bddb", binwidth = 2) + 
  theme_bw() +
  ylab("Number of vOTUs") + 
  xlab("Size (Kb)") +
  scale_x_continuous(breaks=seq(0, 2000000, 100)) +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 12),)
p_dens 


#Statistics of vOTUs length about raw data 
print("Statistique of vOTUs length about raw data")
summary(contigs_info1$Length)

```

#### Statistic of vOTU for cabbage  and carrot samples
```{r statistic on vOTU size for all samples, message=FALSE, warning=FALSE}
## Contigs length statistics

# Statistics of the vOTUs length for cabbage samples after control cleaning
contigs_info_cabbage_sample <- data.frame(tax_table(obj_Cabbage))
print("Statistique of vOTUs length about cabbage samples without control samples")
summary(as.numeric(contigs_info_cabbage_sample$Length))

contigs_info_cabbage_sample$Length <- as.numeric(contigs_info_cabbage_sample$Length)

#Selected contigs with a size higher than  100 kb
contigs_length_stat_cab <- contigs_info_cabbage_sample %>%
  arrange(-Length) %>%
  dplyr::filter(Length >= 100000)

table(contigs_info_cabbage_sample$group)


# Draw density histogram of the distribution of contig sizes. 
p_dens_cabbage <- ggplot(contigs_info_cabbage_sample, aes(x=(Length/1000))) +
  geom_histogram(colour="black", fill="#a6bddb", binwidth = 2) + 
  theme_bw() +
  ylab("Number of vOTUs") + 
  xlab("Size (Kb)") +
  scale_x_continuous(breaks=seq(0, 2000000, 100)) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14))
p_dens_cabbage 

# Statistics of the vOTUs length for carrot samples after control cleaning
contigs_info_carrot_sample <- data.frame(tax_table(obj_Carrot))
print("Statistique of vOTUs length about carrot samples without control samples")
summary(as.numeric(contigs_info_carrot_sample$Length))

contigs_info_carrot_sample$Length <- as.numeric(contigs_info_carrot_sample$Length)

#Selected contigs with a size higher than  100 kb
contigs_length_stat_cab <- contigs_info_carrot_sample %>%
  arrange(-Length) %>%
  dplyr::filter(Length >= 100000)

# Draw density histogram of the distribution of contig sizes. 
p_dens_carrot <- ggplot(contigs_info_carrot_sample, aes(x=(Length/1000))) +
  geom_histogram(colour="black", fill="#a6bddb", binwidth = 2) + 
  theme_bw() +
  ylab("Number of vOTUs") + 
  xlab("Size (Kb)") +
  scale_x_continuous(breaks=seq(0, 2000000, 100)) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14))
p_dens_carrot

p_dens_complet <- ggarrange(p_dens_cabbage,p_dens_carrot, labels = c("Cabbage","Carrot"),
                            common.legend = FALSE,
                            nrow = 2,
                            label.y = 1.10)+
  theme(plot.background = element_rect(fill = "white", color = NA),
           plot.margin = margin(1, 1, 1, 1, "cm"))
p_dens_complet


```

## Beta-diversity

Is there any differences regarding viral composition between samples
according to the vegetable matrix?

Methods:

-   calculate a distance matrix between samples with bc
    (presence/absence of contigs corrected by abundance) and jcc methods
    (presence/absence of contigs only)

-   apply a PERMANOVA test (permutational analysis of variance) to test
    the effect of the Kinetic factor

-   perform an ordination (PCoA or NMDS) for visualisation purpose. PCoA
    is a parametric method of ordination and NMDS is non-parametric
    (turn distances into ranked distance).

#### Normalisation

We transformed abundance data by TPM. We normalized by contig length and
transformed to proportions

```{r TPM transformation, message=FALSE, warning=FALSE}
# Extract abundance table 
abundance_table <- as.data.frame(otu_table(obj_All)) %>% rownames_to_column(var = "Contig")
#recovery the abundance table of obj_clean in the form of dataframe. The row name is the contigs identify.

# Add contig length and set the vector of contig length
abundance_table2 <- left_join(abundance_table, contigs_info1, by=c("Contig"="vOTUs_ID"))
vect_length <- Contigs_info_clean$Length 
# join the abundance table and the list of contig length into the data frame abundance_table2 based on the column contig identity.
# creat the vector vect_length that contain all length of the abundance table.
  
# Transform counts to TPM
## First divide the count matrix by the vector of contig length
obj_All_TPM <- obj_All
otu_table(obj_All_TPM) <- otu_table(obj_All_TPM) / vect_length 

## Second transform the count matrix into proportions 
count_to_prop_sub <- function(x) {return( x / sum(x) )}
obj_All_TPM <- transform_sample_counts(obj_All_TPM, count_to_prop_sub)
#sample_sums(obj_All_TPM)
# creat function count_to_prop that calculate the proportion based of the count of abundance divide by the sum of the abundance.
# transform_sample_counts: Transform abundance data in an otu_table, sample-by-sample.
# sample_sums: Returns the total number of individuals observed from each sample.

abun_test<-as.data.frame(otu_table(obj_All_TPM)) %>%
  rownames_to_column(.,var = "contig")

```

##### Matrix calcul

```{r Matrix calcul, message=FALSE, warning=FALSE}
dist_bc <- phyloseq::distance(obj_All_TPM, method = "bray") 
dist_jcc <- phyloseq::distance(obj_All_TPM, method = "jaccard")
```

##### Permanova test

Is there any differences regarding viral composition between samples
according to the vegetable matrix?

```{r Permanova test in function of Sampling_day, message=FALSE, warning=FALSE}
# on BC 
set.seed(030623)
adonis_bc_sample <- adonis2(formula = dist_bc~Sampling_day, data = data.frame(sample_data(obj_All_TPM)))
adonis_bc_sample
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample$R2[1],3), ", p =", adonis_bc_sample$`Pr(>F)`[1],"**") 

# on JCC
set.seed(030623)
adonis_jcc_sample <- adonis2(formula = dist_jcc~Sampling_day, data = data.frame(sample_data(obj_All_TPM)))
adonis_jcc_sample
res.adonis.jcc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_jcc_sample$R2[1],3), ", p =", adonis_jcc_sample$`Pr(>F)`[1],"**")
```
```{r Permanova test in function of Sampling_day and matrice, message=FALSE, warning=FALSE}
# on BC 
set.seed(030623)
adonis_bc_sample <- adonis2(formula = dist_bc~Sampling_day*Matrix, data = data.frame(sample_data(obj_All_TPM)))
adonis_bc_sample
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample$R2[1],3), ", p =", adonis_bc_sample$`Pr(>F)`[1],"**") 

# on JCC
set.seed(030623)
adonis_jcc_sample <- adonis2(formula = dist_jcc~Sampling_day*Matrix, data = data.frame(sample_data(obj_All_TPM)))
adonis_jcc_sample
res.adonis.jcc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_jcc_sample$R2[1],3), ", p =", adonis_jcc_sample$`Pr(>F)`[1],"**")
```
A significant effect of the matrix was observed on the composition of
the virome both with BC and Jaccard distances (p-value of 0.005, R2
values of 2.4031 and 1.8732, respectly). The viromes thus varied
according to the type of vegetables used.



## Cabbage analyses

#### Normalisation

We transformed abundance data by TPM. We normalized by contig length and
transformed to proportions

```{r TPM transformation, message=FALSE, warning=FALSE}
# Extract abundance table 
abundance_table_cabbage <- as.data.frame(otu_table(obj_Cabbage)) %>% rownames_to_column(var = "Contig")
#recovery the abundance table of obj_clean in the form of dataframe. The row name is the contigs identify.

# Add contig length and set the vector of contig length
abundance_table_cabbage2 <- left_join(abundance_table_cabbage, contigs_info1, by=c("Contig"="vOTUs_ID"))
vect_length <- abundance_table_cabbage2$Length 
# join the abundance table and the list of contig length into the data frame abundance_table2 based on the column contig identity.
# creat the vector vect_length that contain all length of the abundance table.
  
# Transform counts to TPM
## First divide the count matrix by the vector of contig length
obj_cabbage_TPM <- obj_Cabbage
otu_table(obj_cabbage_TPM) <- otu_table(obj_cabbage_TPM) / vect_length 

## Second transform the count matrix into proportions 
count_to_prop_sub <- function(x) {return( x / sum(x) )}
obj_cabbage_TPM <- transform_sample_counts(obj_cabbage_TPM, count_to_prop_sub)
#sample_sums(obj_DNA_TPM)
# creat function count_to_prop that calculate the proportion based of the count of abundance divide by the sum of the abundance.
# transform_sample_counts: Transform abundance data in an otu_table, sample-by-sample.
# sample_sums: Returns the total number of individuals observed from each sample.

obj_cabbage_TPM@sam_data$Sampling_day <- factor(obj_cabbage_TPM@sam_data$Sampling_day, 
                                                 levels = c("0", "2", "7","14","21","28"))

abun_cabbage<-as.data.frame(otu_table(obj_cabbage_TPM)) %>%
  rownames_to_column(.,var = "contig")

```

#Alpha-diversity cabbage virome
We measured the Shannon index to evaluate the effect of the kinetic on the overall phage diversity observed in the samples.
```{r alpha diversity, message=FALSE, warning=FALSE}
pal_lib <- c("0"="#3D5941FF","2"="#778868FF","7"="#B5B991FF", "14"="#DE8A5AFF","21"="#BE7245FF","28"="#8C311CFF")
alpha.diversity <- estimate_richness(obj_cabbage_TPM, measures = "Shannon")
alpha.diversity
summary(alpha.diversity)

res.kruskal <- alpha.diversity %>% kruskal_test(Shannon ~ sample_data(obj_cabbage_TPM)$Sampling_day)
res.kruskal
res.kruskal.short <- paste("Kruskal-Wallis test, Chi2 = ", round(res.kruskal$statistic[1],3), ", p = ", round(res.kruskal$p[1],3))

p1_SI_cab <- plot_richness(obj_cabbage_TPM, color = "Sampling_day", measures = "Shannon", x="Sampling_day") +
  # Create a richness plot using the `plot_richness` function
  # - `obj_cabbage_TPM`: The data object containing the richness or diversity metrics for cabbage samples.
  # - `color = "Sampling_day"`: Points/lines are colored by the 'Sampling_day' variable.
  # - `measures = "Shannon"`: The Shannon index is used to measure alpha diversity.
  # - `x = "Sampling_day"`: Sampling day is plotted on the x-axis.
  labs(subtitle = res.kruskal.short) +
  # Adds a subtitle to the plot, containing results from a Kruskal-Wallis test stored in `res.kruskal.short`.
  #scale_color_manual(values=pal_step) + 
  # (Commented out) Allows for manual adjustment of the color palette. `pal_step` would contain the color values.
  theme_bw()+ 
  # Applies a clean, black-and-white theme to the plot.
  #theme(strip.background = element_blank(), strip.text = element_blank()) +
  # Removes the background and text from any strip elements (facet labels).
  ggtitle("SHANNON INDEX")+
  # Sets the main title of the plot to "SHANNON INDEX".
  xlab("Time (day)") +
  theme(legend.position = "none", 
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 14),
        axis.text.x = element_text(size = 14,hjust=1),
        axis.ticks = element_line(size = 0))+
  scale_colour_manual(values = pal_lib)
  # Removes the legend from the plot.
p1_SI_cab <- add_summary(p1_SI_cab, "median_hilow", size = 0.5)
# Adds summary statistics (median and high/low range) to the plot using `add_summary`.
# The size of the summary points/lines is set to 0.5.

alpha.diversity$Sampling_day <- sample_data(obj_cabbage_TPM)$Sampling_day

# Modèle ANOVA
model <- aov(Shannon ~ Sampling_day, data = alpha.diversity)

# Test de Dunnett

dunnett <- glht(model, linfct = mcp(Sampling_day = "Dunnett"))
summary_dunnett <- summary(dunnett)

# Ajouter les annotations significatives
p1_SI_cab <- p1_SI_cab +
  geom_text(aes(
    x = as.numeric(sample_data(obj_cabbage_TPM)$Sampling_day),
    y = max(alpha.diversity$Shannon) + 0.2, # Ajuster selon l'échelle
    label = signif_levels[sample_data(obj_cabbage_TPM)$Sampling_day]
  ), color = "black", size = 4, inherit.aes = FALSE)

# Extraire les p-values
pvalues <- summary_dunnett$test$pvalues

# Convertir en niveaux significatifs (étoiles)
signif_levels <- cut(
  pvalues,
  breaks = c(0, 0.001, 0.01, 0.05, 1),
  labels = c("***", "**", "*", "ns")
)

# Ajouter des noms aux annotations
names(signif_levels) <- levels(factor(alpha.diversity$Sampling_day))[-1] # Exclut le contrôle

# Ajouter les annotations significatives au graphique
p1_SI_cab_dunn <- p1_SI_cab +
  geom_text(
    aes(
      x = as.numeric(as.factor(alpha.diversity$Sampling_day)), # Position x
      y = max(alpha.diversity$Shannon) + 0.2,                 # Position y ajustée
      label = signif_levels[Sampling_day]
    ),
    color = "black",
    size = 4,
    inherit.aes = FALSE
  )
p1_SI_cab_dunn

```
We observed a high Shannon index of the initial day of fermentation, meaning a high specific diversity specific 
a small decrease in the Shannon index from second day of fermentation but the effect of the kinetic step was not significant (Kruskal wallis test, p>0.05).

## Sparsity
We next calculated the sparisity of the abundance table :
```{r alpha diversity, message=FALSE, warning=FALSE}
sparsity <- 100 * sum(otu_table(obj_cabbage_TPM) == 0) / length(otu_table(obj_cabbage_TPM))
cat("Sparsity =", sparsity, "%")
```

## Beta diversity

##### Matrix calcul

```{r Matrix calcul, message=FALSE, warning=FALSE}
dist_bc <- phyloseq::distance(obj_cabbage_TPM, method = "bray") 
dist_jcc <- phyloseq::distance(obj_cabbage_TPM, method = "jaccard")
```

##### Permanova test

Is there any differences regarding viral composition between samples
according to the vegetable matrix?

```{r Permanova test in function of Sampling_day, message=FALSE, warning=FALSE}
# on BC 
set.seed(030623)
adonis_bc_sample <- adonis2(formula = dist_bc~Sampling_day, data = data.frame(sample_data(obj_cabbage_TPM)))
adonis_bc_sample
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample$R2[1],3), ", p =", adonis_bc_sample$`Pr(>F)`[1],"**") 

# on JCC
set.seed(030623)
adonis_jcc_sample <- adonis2(formula = dist_jcc~Sampling_day, data = data.frame(sample_data(obj_cabbage_TPM)))
adonis_jcc_sample
res.adonis.jcc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_jcc_sample$R2[1],3), ", p =", adonis_jcc_sample$`Pr(>F)`[1],"**")
```

##### Plot construction

```{r Plot construction, message=FALSE, warning=FALSE}
pal_lib <- c("0"="#3D5941FF","2"="#778868FF","7"="#B5B991FF", "14"="#DE8A5AFF","21"="#BE7245FF","28"="#8C311CFF")

# Si vous voulez ordonner Sampling_day selon un ordre spécifique
obj_cabbage_TPM@sam_data$Sampling_day <- factor(obj_cabbage_TPM@sam_data$Sampling_day, 
                                                 levels = c("0", "2", "7","14","21","28"))  # Spécifiez votre ordre ici

# Ordination et graphique
#PCoA with BC
ord_PCoA_bc <- ordinate(obj_cabbage_TPM, method = "PCoA", distance = dist_bc)

p1B_cab <- plot_ordination(obj_cabbage_TPM, ord_PCoA_bc,color = "Sampling_day",shape = "Matrix")+
  labs(title = "BRAY-CURTIS",
    subtitle = res.adonis.bc_sample)+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 10),
        plot.background = element_rect(fill = "white", color ="white"),
        panel.background = element_rect(fill = "white", color ="white"),
        axis.line = element_line(colour = "black"))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib)
p1B_cab


#PCoA with Jaccard
ord_PCoA_jcc <- ordinate(obj_cabbage_TPM, method = "PCoA", distance = dist_jcc)

p2B_cab <- plot_ordination(obj_cabbage_TPM, ord_PCoA_jcc,color = "Sampling_day")+
  labs(title = "JACCARD",
    subtitle = res.adonis.jcc_sample)+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 15),
        plot.background = element_rect(fill = "white", color ="white"),
        panel.background = element_rect(fill = "white", color ="white"),
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size=15),
        legend.text = element_text(size=14))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib,name = "Time (day)")+
  theme(axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 14),
        axis.text.x = element_text(size = 14,hjust=1),
        axis.ticks = element_line(size = 0))
p2B_cab

```
```{r Pheatmap normalized filtred, message=FALSE, warning=FALSE}
#Calculate the mean of relative abundance of each contigs to have the statistic of abundance.
sum_taxa <- taxa_sums(obj_cabbage_TPM)/18
summary(sum_taxa)

#transform log function
log_fun <- function(x){
  (log((x + 0.000001),10))
}

Ab_maj <- 0.001
x <- taxa_sums(obj_cabbage_TPM)/18 # Calculate the mean of relative abundance of each contigs
keepTaxa_maj <- x > Ab_maj
prunedSet_cabbage_maj <- prune_taxa(keepTaxa_maj, obj_cabbage_TPM)
#prunedSet_DNA_Med

vOTUs_prunedSet_cabbage_maj <-  data.frame(otu_table(prunedSet_cabbage_maj))
#transform log 
vOTUs_prunedSet_cabbage_norm_maj <- t(apply(vOTUs_prunedSet_cabbage_maj, 1, log_fun))

p6 <- fviz_nbclust(vOTUs_prunedSet_cabbage_norm_maj, kmeans, method = "gap_stat", k.max = 15) + theme_minimal() + ggtitle("gap_stat")
p6

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/", Sys.Date(), "_gap_stat_p6.rds")
saveRDS(p6,file_name)
ggsave(file_name, plot = p6, device = "svg", width = 8, height = 5, create.dir = TRUE)

k_choose <- 6

set.seed(123)
km <- kmeans(vOTUs_prunedSet_cabbage_norm_maj, centers = k_choose, nstart = 25)
km$cluster

vOTUs_annotations_maj <- data.frame(tax_table(prunedSet_cabbage_maj))
table(vOTUs_annotations_maj$Host_genus)
# Extract vOTU metadata
Lifestyle_group <- vOTUs_annotations_maj["Type_vibrant"]
virus_group <- vOTUs_annotations_maj["group"]
Host_genus <-  vOTUs_annotations_maj["Host_genus"]

# Creat dataframe `row_ann` for the row annotation
row_ann <- data.frame(Lifestyle_group,Host_genus,virus_group)
#row_ann <- data.frame(Host_genus)

sample_annotation_Med<-data.frame(sample_data(obj_cabbage_TPM))
col_ann <- sample_annotation_Med["Sampling_day"]


# Replace empty values with NA in data frame `row_ann`
row_ann$Type_vibrant[is.na(row_ann$'Life style')] <- 'No prediction'
row_ann$Host_genus[is.na(row_ann$genus)] <- 'No prediction'

table(row_ann$Host_genus)
# Convert the column into a factor with the appropriate levels
row_ann$Type_vibrant <- factor(row_ann$Type_vibrant, levels = c("Temperate", "Virulent", "unknown"))
row_ann$Host_genus <- factor(row_ann$Host_genus, levels = c("Citrobacter","Enterobacter","Erwinia","Escherichia","Ewingella","Klebsiella","Lactiplantibacillus","Lentilactobacillus","Leuconostoc","Levilactobacillus","Pantoea","Pediococcus","Providencia","Pseudomonas","Rahnella","Secundilactobacillus","No prediction"))

names(row_ann)[names(row_ann) == "Type_vibrant"] <- "Life style"
names(row_ann)[names(row_ann) == "group"] <- "Group"
names(row_ann)[names(row_ann) == "Host_genus"] <- "Host genus"
names(col_ann)[names(col_ann) == "Sampling_day"] <- "Sampling day"

custom_colnames <- c("Sauerkraut_0A","Sauerkraut_0B","Sauerkraut_0C","Sauerkraut_2A","Sauerkraut_2B","Sauerkraut_2C","Sauerkraut_7A","Sauerkraut_7B","Sauerkraut_7C","Sauerkraut_14A","Sauerkraut_14B","Sauerkraut_14C","Sauerkraut_21A","Sauerkraut_21B","Sauerkraut_21C",
"Sauerkraut_28A","Sauerkraut_28B","Sauerkraut_28C")

ann_color = list("Life style"=c("Temperate"="grey",
                          "Virulent"="black",
                          "No prediction"="white"),
                 "Sampling day"=c("0"="#1C3333FF" , 
                            "2"="#226060FF", 
                            "7"="#639CA4FF",
                            "14"="#D2AD7CFF",
                            "21"="#BE7245FF",
                            "28"="#8C311CFF"),
                 "Host genus"=c('Citrobacter' = '#33a02c',
                                'Enterobacter' = '#8e0152',
                                'Erwinia' = 'orchid',
                                'Escherichia' = '#e31a1c',
                                'Ewingella' = '#fdbf6f',
                                'Klebsiella' = 'rosybrown1',
                                'Lactiplantibacillus' = '#b2df8a',
                                'Lentilactobacillus' = '#6a3d9a',
                                'Leuconostoc' = '#a6cee3',
                                'Levilactobacillus' = '#fb9a99',
                                "Pantoea"= "#cab2d6",
                                "Pediococcus"= "#5e4fa2",
                                "Providencia"= "rosybrown",
                                "Pseudomonas"= "maroon3",
                                "Rahnella"= "#ff7f00",
                                "Secundilactobacillus"= "#1f78b4",
                                "No prediction"="white"),
                 "Group"=c("dsDNAphage"="powderblue",
                                "Lavidaviridae (virophages)" ="azure2",
                                "Nucleocytoviricota (giant viruses)" = "lightsteelblue2",
                                 "RNA virus" = "lightblue4",
                                "ssDNA viruses"="deepskyblue",
                             "Unknown"= "cornflowerblue"))


ph3_maj <- pheatmap(vOTUs_prunedSet_cabbage_norm_maj,
              color = colorRampPalette(rev(brewer.pal(11, "Spectral")))(100),
               show_rownames = FALSE,
               cluster_cols = FALSE,
               cutree_rows=6,
               annotation_row = row_ann,
               annotation_col = col_ann,
               annotation_colors = ann_color,
               angle_col = 45,
               fontsize_col = 10,
               fontsize = 10,
               labels_col = custom_colnames,
               border_color = NA,
               legend_breaks = c(-5, -3, -1, max(vOTUs_prunedSet_cabbage_norm_maj)),
               legend_labels = c("-5", "-3", "-1", "log10"))
ph3_maj <- as.ggplot(ph3_maj)+
  theme(plot.background = element_rect(fill = "white", color = "white"))
ph3_maj
setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/",Sys.Date(),"_Figure_4B.svg")
ggsave(file_name,plot =ph3_maj,device = "svg",width = 12, height = 15, dpi = 300)

```

## Carrot analyses

#### Normalisation

We transformed abundance data by TPM. We normalized by contig length and
transformed to proportions

```{r TPM transformation, message=FALSE, warning=FALSE}
# Extract abundance table 
abundance_table_carrot <- as.data.frame(otu_table(obj_Carrot)) %>% rownames_to_column(var = "Contig")
#recovery the abundance table of obj_clean in the form of dataframe. The row name is the contigs identify.

# Add contig length and set the vector of contig length
abundance_table_carrot2 <- left_join(abundance_table_carrot,contigs_info1, by=c("Contig"="vOTUs_ID"))
vect_length <- abundance_table_carrot2$Length 
# join the abundance table and the list of contig length into the data frame abundance_table2 based on the column contig identity.
# creat the vector vect_length that contain all length of the abundance table.
  
# Transform counts to TPM
## First divide the count matrix by the vector of contig length
obj_Carrot_TPM <- obj_Carrot
otu_table(obj_Carrot_TPM) <- otu_table(obj_Carrot_TPM) / vect_length 

## Second transform the count matrix into proportions 
count_to_prop_sub <- function(x) {return( x / sum(x) )}
obj_Carrot_TPM <- transform_sample_counts(obj_Carrot_TPM, count_to_prop_sub)
#sample_sums(obj_DNA_TPM)
# creat function count_to_prop that calculate the proportion based of the count of abundance divide by the sum of the abundance.
# transform_sample_counts: Transform abundance data in an otu_table, sample-by-sample.
# sample_sums: Returns the total number of individuals observed from each sample.

obj_Carrot_TPM@sam_data$Sampling_day <- factor(obj_Carrot_TPM@sam_data$Sampling_day, 
                                                 levels = c("0", "2", "7","14","21","28"))

abun_test<-as.data.frame(otu_table(obj_Carrot_TPM)) %>%
  rownames_to_column(.,var = "contig")
```


#Alpha-diversity virome
We measured the Shannon index to evaluate the effect of the kinetic on the overall phage diversity observed in the samples.
```{r alpha diversity, message=FALSE, warning=FALSE}
pal_lib <- c("0"="#3D5941FF","2"="#778868FF","7"="#B5B991FF", "14"="#DE8A5AFF","21"="#BE7245FF","28"="#8C311CFF")

alpha.diversity<-estimate_richness(obj_Carrot_TPM, measures = "Shannon")
alpha.diversity
summary(alpha.diversity)
res.kruskal <- alpha.diversity %>% kruskal_test(Shannon ~ sample_data(obj_Carrot_TPM)$Sampling_day)
res.kruskal
res.kruskal.short <- paste("Kruskal-Wallis test, Chi2 = ", round(res.kruskal$statistic[1],3), ", p = ", round(res.kruskal$p[1],3))
p1_SI_CAR <- plot_richness(obj_Carrot_TPM, color = "Sampling_day", measures = "Shannon", x="Sampling_day") +
  # Create a richness plot using the `plot_richness` function
  # - `obj_Carrot_TPM`: The data object containing the richness or diversity metrics for cabbage samples.
  # - `color = "Sampling_day"`: Points/lines are colored by the 'Sampling_day' variable.
  # - `measures = "Shannon"`: The Shannon index is used to measure alpha diversity.
  # - `x = "Sampling_day"`: Sampling day is plotted on the x-axis.

  labs(subtitle = res.kruskal.short) +
  # Adds a subtitle to the plot, containing results from a Kruskal-Wallis test stored in `res.kruskal.short`.

  #scale_color_manual(values=pal_step) + 
  # (Commented out) Allows for manual adjustment of the color palette. `pal_step` would contain the color values.

  theme_bw() + 
  # Applies a clean, black-and-white theme to the plot.

  #theme(strip.background = element_blank(), strip.text = element_blank()) +
  # Removes the background and text from any strip elements (facet labels).

  ggtitle("SHANNON INDEX")+
  # Sets the main title of the plot to "SHANNON INDEX".

  theme(legend.position = "none")+
  scale_colour_manual(values = pal_lib)+
  xlab("Time (day)") +
  theme(legend.position = "none", 
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        legend.title = element_text(size = 14),
        axis.text.x = element_text(size = 14,hjust=1),
        axis.ticks = element_line(size = 0))+
  scale_colour_manual(values = pal_lib)


p1_SI_CAR <- add_summary(p1_SI_CAR, "median_hilow", size = 0.5)
# Adds summary statistics (median and high/low range) to the plot using `add_summary`.
# The size of the summary points/lines is set to 0.5.

p1_SI_CAR
# Displays the plot `p1`.



#file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Shannon_index_carrot.tiff")
#ggsave(file_name,plot =p1_SI_CAR,device = "tiff",width = 8, height = 6, dpi = 300)
```
## Sparsity
We next calculated the sparisity of the abundance table :
```{r alpha diversity, message=FALSE, warning=FALSE}
sparsity <- 100 * sum(otu_table(obj_Carrot_TPM) == 0) / length(otu_table(obj_Carrot_TPM))
cat("Sparsity =", sparsity, "%")
```

## Beta diversity

##### Matrix calcul

```{r Matrix calcul, message=FALSE, warning=FALSE}
dist_bc <- phyloseq::distance(obj_Carrot_TPM, method = "bray") 
dist_jcc <- phyloseq::distance(obj_Carrot_TPM, method = "jaccard")
```

##### Permanova test

Is there any differences regarding viral composition between samples
according to the vegetable matrix?

```{r Permanova test in function of Sampling_day, message=FALSE, warning=FALSE}
# on BC 
set.seed(030623)
adonis_bc_sample <- adonis2(formula = dist_bc~Sampling_day, data = data.frame(sample_data(obj_Carrot_TPM)))
adonis_bc_sample
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample$R2[1],3), ", p =", adonis_bc_sample$`Pr(>F)`[1],"") 

# on JCC
set.seed(030623)
adonis_jcc_sample <- adonis2(formula = dist_jcc~Sampling_day, data = data.frame(sample_data(obj_Carrot_TPM)))
adonis_jcc_sample
res.adonis.jcc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_jcc_sample$R2[1],3), ", p =", adonis_jcc_sample$`Pr(>F)`[1],"")
```
##### Plot construction

```{r Plot construction, message=FALSE, warning=FALSE}
pal_lib <- c("0"="#3D5941FF","2"="#778868FF","7"="#B5B991FF", "14"="#DE8A5AFF","21"="#BE7245FF","28"="#8C311CFF")

# Si vous voulez ordonner Sampling_day selon un ordre spécifique
obj_Carrot_TPM@sam_data$Sampling_day <- factor(obj_Carrot_TPM@sam_data$Sampling_day, 
                                                 levels = c("0", "2", "7","14","21","28"))  # Spécifiez votre ordre ici

# Ordination et graphique
#PCoA with BC
ord_PCoA_bc <- ordinate(obj_Carrot_TPM, method = "PCoA", distance = dist_bc)

p1B <- plot_ordination(obj_Carrot_TPM, ord_PCoA_bc,color = "Sampling_day")+
  labs(title = "BRAY-CURTIS",
    subtitle = res.adonis.bc_sample)+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 10),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "black"))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib)
p1B


#PCoA with Jaccard
ord_PCoA_jcc <- ordinate(obj_Carrot_TPM, method = "PCoA", distance = dist_jcc)

p2B <- plot_ordination(obj_Carrot_TPM, ord_PCoA_jcc,color = "Sampling_day")+
  labs(title = "JACCARD",
    subtitle = res.adonis.jcc_sample)+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 10),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "black"))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib)
p2B

```
```{r Pheatmap normalized filtred, message=FALSE, warning=FALSE}
# Extraire la table d'abondance
otu_table <- as.data.frame(otu_table(obj_Carrot_TPM)) %>%
  rownames_to_column(var = "Contig")

#Extraire les metadata des échantillons
metadata <- sample_data(obj_Carrot_TPM)
metadata_df <- data.frame(SampleID = rownames(metadata), metadata, row.names = NULL)
dim(metadata_df)  # Vérifier si on a toujours 18 lignes
head(metadata_df)


# Fusion des tables d'abondance et des métadonnées
otu_long <- otu_table %>%
  pivot_longer(-Contig, names_to = "SampleID", values_to = "Abundance") %>%
  left_join(metadata_df, by = "SampleID")

# Définir un seuil d'abondance minimum (ajustez selon vos données)
seuil <- 0.0000001  # Exemple : Abondance ≥ 10

otu_long_test <- otu_long[c("Contig","Sampling_day","Abundance")]
otu_long_test$Sampling_day <- as.numeric(as.character(otu_long$Sampling_day))

otu_filtered <- otu_long %>%
  group_by(Contig, Sampling_day) %>% 
  dplyr::summarize(N_replicats = sum(Abundance >= seuil, na.rm = TRUE), .groups = "drop")%>% 
  filter(N_replicats == 3)  # Garde les contigs présents dans les 3 réplicats

# Liste des contigs sélectionnés
contigs_selectionnes <- unique(otu_filtered$Contig)

# Filtrer la table d'abondance initiale
otu_final_init <- otu_table %>%
  filter(Contig %in% contigs_selectionnes)

otu_final <- otu_final_init %>%
  column_to_rownames(.,"Contig")

vOTUs_carrot_select_norm <- t(apply(otu_final, 1, log_fun))

if ("Figure_label" %in% colnames(metadata_df)) {
  custom_colnames <- metadata_df$Figure_label
} else {
  stop(" 'NewNames' column doesn't exist in samples data")
}

#Extract vOTUs annotation data 
vOTUs_annotations_carrot <- data.frame(tax_table(obj_Carrot_TPM))%>%
  rownames_to_column(.,var="Contig")

vOTUs_annotations_carrot_top <- left_join(otu_final_init,vOTUs_annotations_carrot, by="Contig")%>%
  column_to_rownames(.,"Contig")

table(vOTUs_annotations_carrot_top$Type_vibrant)

# Extract vOTU metadata
Lifestyle_group <- vOTUs_annotations_carrot_top["Type_vibrant"]
Host_genus <-  vOTUs_annotations_carrot_top["Host_genus"]
#virus_group <- vOTUs_annotations_Med["group"]

# Creat dataframe `row_ann` for the row annotation
row_ann <- data.frame(Lifestyle_group,Host_genus)
sample_annotation_Med <- data.frame(sample_data(obj_Carrot_TPM))
col_ann <- sample_annotation_Med["Sampling_day"]

names(row_ann)[names(row_ann) == "Type_vibrant"] <- "Life style"
names(col_ann)[names(col_ann) == "Sampling_day"] <- "Sampling day"

custom_colnames <- c("Carrot_0A","Carrot_0B","Carrot_0C","Carrot_2A","Carrot_2B","Carrot_2C","Carrot_7A","Carrot_7B","Carrot_7C","Carrot_14A","Carrot_14B","Carrot_14C","Carrot_21A","Carrot_21B","Carrot_21C","Carrot_28A","Carrot_28B","Carrot_28C")

# Replace empty values with NA in data frame `row_ann`
row_ann$`Life style`[is.na(row_ann$`Life style`)] <- 'No prediction'
row_ann$Host_genus[is.na(row_ann$Host_genus)] <- 'No prediction'

# Convert the column into a factor with the appropriate levels
row_ann$`Life style` <- factor(row_ann$`Life style`, levels = c("Temperate", "Virulent"))
row_ann$Host_genus <- factor(row_ann$Host_genus, levels = c("Lactiplantibacillus","Leuconostoc","No prediction"))

ann_color = list("Life style"=c("Temperate"="grey",
                          "Virulent"="black",
                          "No prediction"="white"),
                 "Sampling day"=c("0"="#1C3333FF" , 
                            "2"="#226060FF", 
                            "7"="#639CA4FF",
                            "14"="#D2AD7CFF",
                            "21"="#BE7245FF",
                            "28"="#8C311CFF"),
                 "Host_genus"=c('Lactiplantibacillus' = '#b2df8a',
                                'Leuconostoc' = '#a6cee3',
                                "No prediction"="white"))


range(vOTUs_carrot_select_norm, na.rm = TRUE)
summary(as.vector(vOTUs_carrot_select_norm))

ph3_Med <- pheatmap(vOTUs_carrot_select_norm,
              color = colorRampPalette(rev(brewer.pal(11, "Spectral")))(100),
                show_rownames = TRUE,
               cluster_cols = FALSE,
               annotation_row = row_ann,
               annotation_col = col_ann,
               annotation_colors = ann_color,
               angle_col = 45,
               fontsize_col = 12,
               fontsize = 12,
               labels_col = custom_colnames,
               legend_breaks = c(-5, -3, -1, max(vOTUs_carrot_select_norm)),
               legend_labels = c("-5", "-3", "-1", "log10"))
ph3_Med <- as.ggplot(ph3_Med)+
  theme(plot.background = element_rect(fill = "white", color = "white"),
        plot.margin = margin(10, 60, 10, 10))
ph3_Med

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/",Sys.Date(),"_Figure_5B.svg")
ggsave(file_name,plot =ph3_Med,device = "svg",width = 15, height = 8, dpi = 300)

```
Temporal variation of the virome during carrot fermentation. For better clarity, only vOTUs consistently detected in all three replicates at least at one time point were represented. The color scale represents the log-transformed relative abundance and ranges from blue (low relative abundance) to red (high relative abundance). Host genus and lifestyle of the different vOTUs are indicated when available.

```{r figure article, message=FALSE, warning=FALSE}
otu_df <- as.data.frame(otu_table(obj_All))
if (taxa_are_rows(obj_All)) {
  otu_df <- otu_df %>% rownames_to_column("vOTU")
} else {
  otu_df <- otu_df %>% t() %>% as.data.frame() %>% rownames_to_column("vOTU")
}

sample_df <- as.data.frame(sample_data(obj_All))
sample_df <- rownames_to_column(sample_df, "SampleID")


sample_df <- as.data.frame(sample_data(obj_All), stringsAsFactors = FALSE)


sample_df$SampleID <- rownames(sample_df)

presence_df <- otu_df %>%
  pivot_longer(-vOTU, names_to = "SampleID", values_to = "abundance") %>%
  left_join(sample_df, by = "SampleID") %>%
  group_by(vOTU, Matrix) %>%
  summarise(present = any(abundance > 0), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = Matrix, values_from = present, values_fill = FALSE)

presence_df <- presence_df %>%
  mutate(matrix_presence = case_when(
    Cabbage & Carrot ~ "both",
    Carrot ~ "carrot",
    Cabbage ~ "cabbage",
    TRUE ~ "absent"
  ))

tax_df <- as.data.frame(tax_table(obj_All)) %>%
  rownames_to_column("vOTU")

tax_df <- left_join(tax_df, presence_df, by = "vOTU")

tax_table(obj_All) <- as.matrix(tax_df %>% column_to_rownames("vOTU"))

contigs_info_test <- as.data.frame(tax_table(obj_All))
contigs_info_test$Type_vibrant [is.na(contigs_info_test$Type_vibrant)] <- 'Unknown'

contigs_summary <- contigs_info_test %>%
  group_by(matrix_presence,Type_vibrant, group,Host_family,) %>%
  summarise(freq = n(), .groups = "drop")

contigs_set <- ggforce::gather_set_data(contigs_summary,c("matrix_presence","Type_vibrant", "group","Host_family"))
contigs_set$x <- factor(contigs_set$x,
                        levels = c("matrix_presence","Type_vibrant","group","Host_family" ))

sum_col <- sum(contigs_summary$freq, na.rm = TRUE)


new_names <- c(
  "matrix_presence" = "Matrix",
  "Type_vibrant" = "Phage life cycle",
  "group" = "Type of virus", 
  "Host_family" = "Host family"
)

summmary_virome <- ggplot(contigs_set, aes(x = x, id = id, split = y, value = freq)) +
  geom_parallel_sets(aes(fill = Type_vibrant), alpha = 0.7, axis.width = 0.15) +
  geom_parallel_sets_axes(axis.width = 0.15) +
  geom_parallel_sets_labels(angle = 0, size = 4, face = "bold",colour = "black", nudge_x = 0.1, hjust = 0) +
  scale_x_discrete(labels = new_names) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2", 
                    limits = c("Temperate", "Virulent", "Unknown")) +
  labs(fill = "Phage life cycle")+
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),
    plot.title = element_blank()
  )
summmary_virome

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/",Sys.Date(),"_Figure_3C.svg")
ggsave(file_name,plot =summmary_virome,device = "svg",width = 15, height = 8, dpi = 300)

```

Parallel plot of the viral composition of fermented cabbage and carrot. Each vOTU is represented according to the matrix in which it is present, its predicted life cycle (based on Vibrant prediction), virus type (according to VirSorter2 prediction), and host family (based on iPHoP prediction). The mean relative abundance of vOTUs per host family is shown in parentheses.

```{r Analyses of DNA viruses taxonomy, message=FALSE, warning=FALSE}
# Glomerate OTUs at the genus level (or another taxonomic rank)
obj_genus <- tax_glom(obj_All, taxrank = "Genus")
#Normalize abundances within each sample
obj_genus_rel <- transform_sample_counts(obj_genus, function(x) x / sum(x))

# Convert to a long-format dataframe for ggplot
df_genus <- psmelt(obj_genus_rel)
#any(is.na(df_genus$Genus))

# Identify the 10 most abundant taxa (at the "Genus" level)
top_10_genus_DNA <- df_genus %>%
  group_by(Genus) %>%
  summarise(total_abundance = sum(Abundance)) %>%
  arrange(desc(total_abundance)) %>%
  slice_head(n = 10) %>%
  pull(Genus)

# Replace genera not in the top 10 with "Other"
df_genus$Genus <- ifelse(df_genus$Genus %in% top_10_genus_DNA, df_genus$Genus, "Other")

df_genus$Genus <- factor(df_genus$Genus, levels = c("Brunovirus","Fattrevirus","Lederbergvirus","Prymnesiovirus","Heilongjiangvirus","Sandinevirus","Unaquatrovirus","Seongbukvirus","Sherbrookevirus","Unknown","Other"))

# Sum relative abundances across samples within each Matrix
# Note: at this stage, values are NOT percentages yet -
# the sum per Matrix equals the number of samples in that Matrix, not 1 or 100
df_genus_matrix <- df_genus %>%
  group_by(Matrix, Genus) %>%
  summarise(Abundance = sum(Abundance), .groups = "drop")

# Rescale so abundances sum to 100% within each Matrix
# This effectively computes the mean per-sample relative abundance (in %) for each genus,
# giving equal weight to each sample regardless of its sequencing depth
df_genus_matrix <- df_genus_matrix %>%
  group_by(Matrix) %>%
  mutate(Abundance = Abundance / sum(Abundance) * 100)


df_genus_matrix$Genus <- factor(
  df_genus_matrix$Genus,
  levels = c("Brunovirus","Fattrevirus","Lederbergvirus","Prymnesiovirus",
             "Heilongjiangvirus","Sandinevirus","Unaquatrovirus",
             "Seongbukvirus","Sherbrookevirus","Unknown","Other")
)

p_genus_viral_matrix <- ggplot(
  df_genus_matrix,
  aes(x = Matrix, y = Abundance, fill = Genus)
) +
  geom_bar(stat = "identity", position = "stack", width = 0.7) +
  labs(
    x = "",
    y = "Relative abundance (%)",
    fill = "Genus"
  ) +
  scale_fill_manual(values = c(
    "Brunovirus" ="#ffff99",
    "Fattrevirus"="#6a3d9a",
    "Lederbergvirus" ="#cab2d6",
    "Prymnesiovirus"="#ff7f00",
    "Heilongjiangvirus"="#fdbf6f",
    "Sandinevirus"="#e31a1c",
    "Unaquatrovirus"="#b2df8a",
    "Seongbukvirus"="#1f78b4",
    "Sherbrookevirus"="orange1",
    "Unknown"= "#a6cee3",
    "Other"="#d9d9d9"
  )) +
  theme(
    legend.text = element_text(size = 12, face = "italic"),
    legend.title = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 14))

p_genus_viral_matrix

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")
file_name <- paste0("results/graph/",Sys.Date(),"_Figure_3B.svg")
ggsave(file_name,plot =p_genus_viral_matrix,device = "svg",width = 10, height = 10, dpi = 300)
```
Barplot of  viral communities of lab fermented cabbage and carrot. Relative abundance of reads per matrix (%) at the genus level.



