---
title: "Analysis of microbial enumeration results"
author: "Julia Cantuti Gendre"
date: "2025-02-17"
output: html_document
---

## Description
This script calculates the mean and standard deviation of the microbial counts for each bacterium in order to plot them on a graph, and then performs statistical tests to highlight the effect of phages on the levels of each bacterium in each consortium.

```{r import library,message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(readr)
library(dplyr)
library(gridExtra)
library(cowplot)
library(ggpubr)
sessionInfo()
```

## Importing data

```{r import data, message=FALSE, warning=FALSE}
setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

Results_table <- read.table("results/experimental/Experimental_results.csv", header = TRUE, sep = ",", dec=",")
```

## Calculation of mean and standard deviation of each bacteria per consortium for each day 

```{r calcul mean and standard deviation bacteria, message=FALSE, warning=FALSE}
mean_R_aqua_df <- Results_table %>%
  group_by(Consortia_condition, Day) %>%
  summarise(
    mean_R_aqua = mean(R_aquatilis_.CFU.g., na.rm = TRUE),
    sd_R_aqua = sd(R_aquatilis_.CFU.g., na.rm = TRUE),
    mean_log10_R_aqua = log10(mean(R_aquatilis_.CFU.g., na.rm = TRUE)),   
    sd_log10_R_aqua = log10(mean(R_aquatilis_.CFU.g., na.rm = TRUE) + sd(R_aquatilis_.CFU.g., na.rm = TRUE)) - log10(mean(R_aquatilis_.CFU.g., na.rm = TRUE)),
    .groups = 'drop'
  )
mean_R_aqua_df

mean_R_aqua_df_day <- mean_R_aqua_df %>%
  group_by(Day) %>%
  summarise(
    mean_log10_R_aqua_day = mean(mean_log10_R_aqua, na.rm = TRUE),   
    sd_log10_R_aqua_day =  sd(mean_log10_R_aqua, na.rm = TRUE),
    .groups = 'drop'
)
mean_R_aqua_df_day

mean_L_planta_df <- Results_table %>%
  group_by(Consortia_condition, Day) %>%
  summarise(
    mean_L_planta = mean(L_plantatarum_.CFU.g., na.rm = TRUE),
    sd_L_planta = sd(L_plantatarum_.CFU.g., na.rm = TRUE),
    mean_log10_L_planta = log10(mean(L_plantatarum_.CFU.g., na.rm = TRUE)),   
    sd_log10_L_planta = log10(mean(L_plantatarum_.CFU.g., na.rm = TRUE) + sd(L_plantatarum_.CFU.g., na.rm = TRUE)) - log10(mean(L_plantatarum_.CFU.g., na.rm = TRUE)),
    .groups = 'drop'
  )
mean_L_planta_df

mean_L_planta_df_day <- mean_L_planta_df %>%
  group_by(Day) %>%
  summarise(
    mean_log10_L_planta_day = mean(mean_log10_L_planta, na.rm = TRUE),   
    sd_log10_L_planta_day =  sd(mean_log10_L_planta, na.rm = TRUE),
    .groups = 'drop'
)
mean_L_planta_df_day

mean_L_mes_df <- Results_table %>%
  group_by(Consortia_condition, Day) %>%
  summarise(
    mean_L_mes = mean(L_mesenteroides_.CFU.g., na.rm = TRUE),
    sd_L_mes = sd(L_mesenteroides_.CFU.g., na.rm = TRUE),
    mean_log10_L_mes = log10(mean(L_mesenteroides_.CFU.g., na.rm = TRUE)),   
    sd_log10_L_mes = log10(mean(L_mesenteroides_.CFU.g., na.rm = TRUE) + sd(L_mesenteroides_.CFU.g., na.rm = TRUE)) - log10(mean(L_mesenteroides_.CFU.g., na.rm = TRUE)),
    .groups = 'drop'
  )
mean_L_mes_df

mean_L_mes_df_day <- mean_L_mes_df %>%
  group_by(Day) %>%
  summarise(
    mean_log10_L_mes_day = mean(mean_log10_L_mes, na.rm = TRUE),   
    sd_log10_L_mes_day =  sd(mean_log10_L_mes, na.rm = TRUE),
    .groups = 'drop'
)
mean_L_mes_df_day


mean_PLE001_df <- Results_table %>%
  group_by(Consortia_condition,Day) %>%
  summarise(
    mean_PLE001 = mean(PLE001_.PFU.g., na.rm = TRUE),
    sd_PLE001 = sd(PLE001_.PFU.g., na.rm = TRUE),
    mean_log10_PLE001 = log10(mean(PLE001_.PFU.g., na.rm = TRUE)),   
    sd_log10_PLE001 = log10(mean(PLE001_.PFU.g., na.rm = TRUE) + sd(PLE001_.PFU.g., na.rm = TRUE)) - log10(mean(PLE001_.PFU.g., na.rm = TRUE)),
    .groups = 'drop'
  )
mean_PLE001_df

all_rep_phage <- subset(Results_table, Results_table$Phage=='Yes') 
all_rep_corrected <- all_rep_phage %>%
  mutate(PLE001_.PFU.g. = ifelse(PLE001_.PFU.g. == 1250, 1000, PLE001_.PFU.g.))

mean_PLE001_df_corrected <- all_rep_corrected %>%
  group_by(Consortia_condition,Day) %>%
  summarise(
    mean_PLE001 = mean(PLE001_.PFU.g., na.rm = TRUE),
    sd_PLE001 = sd(PLE001_.PFU.g., na.rm = TRUE),
    mean_log10_PLE001 = log10(mean(PLE001_.PFU.g., na.rm = TRUE)),   
    sd_log10_PLE001 = log10(mean(PLE001_.PFU.g., na.rm = TRUE) + sd(PLE001_.PFU.g., na.rm = TRUE)) - log10(mean(PLE001_.PFU.g., na.rm = TRUE)),
    .groups = 'drop'
  )
mean_PLE001_df_corrected



```

## Representation of the bacterial community for each bacterium constituting the consortia

### Barplot representation of *Rahnella aquatillis* during fermentation

```{r plot Rahnella, message=FALSE, warning=FALSE}
mean_R_aqua_df$Consortia_condition <- factor(mean_R_aqua_df$Consortia_condition, levels = c("NI", "C1","C1P", "C1R", "C1RP"))

graph_R_aqua <- ggplot(mean_R_aqua_df, aes(x = Consortia_condition, y = mean_log10_R_aqua, fill = Consortia_condition)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.5) +
  geom_errorbar(aes(ymin = mean_log10_R_aqua - sd_log10_R_aqua, ymax = mean_log10_R_aqua + sd_log10_R_aqua),
                width = 0.2,                     
                position = position_dodge(0.7)) + 
  scale_fill_manual(values = c("grey","#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  facet_wrap(~Day,  
             labeller = labeller(Day = c(`0` = "Day 0", `3` = "Day 3", `28` = "Day 28")))+
  labs(title = "",
       x = "",
       y = "Rahnella aquatilis \n log(10)CFU/g",
       fill = "Consortia") +
  scale_y_continuous(breaks = seq(0, 9, 0.5)) +
  coord_cartesian(ylim = c(2,9)) +
  
  # Horizontal line for Day 0 to y = 2
  geom_hline(data = subset(mean_R_aqua_df, Day == 0), 
             aes(yintercept = 2), 
             linetype = "dashed", color = "red") +
  
  # Horizontal line for Day 3 and Day 28 at y = 3.3
  geom_hline(data = subset(mean_R_aqua_df, Day %in% c(3, 28)), 
             aes(yintercept = 5), 
             linetype = "dashed", color = "red") +

  theme_bw() +
  theme(panel.background = element_rect(fill="white"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        strip.text = element_text(size = 12))

graph_R_aqua

```

Viable counts of Rahnella aquatilis CIRMBIA 2785.Bar indicates the average value of four replicates (independent laboratory-scale productions) in log CFU/g of vegetables, with error bars representing the standard deviation. The dashed red line indicates the detection limit for microbial enumeration (2 CFU/g for day 0 and 5 CFU/g for day 3 and 28).

### Barplot representation of *Leuconostoc mesenteroides* during fermentation

```{r plot L.mesenteroides, message=FALSE, warning=FALSE}
mean_L_mes_df$Consortia_condition <- factor(mean_L_mes_df$Consortia_condition, levels = c("NI", "C1","C1P","C1R", "C1RP"))


table(mean_L_mes_df$Consortia_condition)
graph_L_mes <- ggplot(mean_L_mes_df, aes(x = Consortia_condition, y = mean_log10_L_mes, fill = Consortia_condition)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.5) +
  geom_errorbar(aes(ymin = mean_log10_L_mes - sd_log10_L_mes, ymax = mean_log10_L_mes + sd_log10_L_mes),
                width = 0.2,                     
                position = position_dodge(0.7)) + 
  scale_fill_manual(values = c("grey","#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  facet_wrap(~Day,  
             labeller = labeller(Day = c(`0` = "Day 0", `3` = "Day 3", `28` = "Day 28")))+
  labs(title = "",
       x = "",
       y = "Leuconostoc mesenteroides \n log(10)CFU/g",
       fill = "Consortia") +
  scale_y_continuous(breaks = seq(0, 9, 0.5)) +
  coord_cartesian(ylim = c(2,9)) +
  
  # Horizontal line for Day 0 to y = 2
  geom_hline(data = subset(mean_L_mes_df, Day == 0), 
             aes(yintercept = 2), 
             linetype = "dashed", color = "red") +
  
  # Horizontal line for Day 3 and Day 28 at y = 3.3
  geom_hline(data = subset(mean_L_mes_df, Day %in% c(3, 28)), 
             aes(yintercept = 5), 
             linetype = "dashed", color = "red") +

  theme_bw() +
  theme(panel.background = element_rect(fill="white"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        strip.text = element_text(size = 12))

graph_L_mes

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

file_name <- paste0("results/graph/", Sys.Date(), "_Figure_6C.svg")
ggsave(file_name, plot = graph_L_mes, device = "svg", width = 10, height = 5, create.dir = TRUE)

```

Viable counts of Leuconostoc mesenteroides. Bar indicates the average value of four replicates (independent laboratory-scale productions) in log CFU/g of vegetables, with error bars representing the standard deviation. The dashed red line indicates the detection limit for microbial enumeration (2 CFU/g for day 0 and 5 CFU/g for day 3 and 28). 

### Barplot representation of *Lactiplantibacillus plantarum* during fermentation

```{r plot L.plantarum, message=FALSE, warning=FALSE}
mean_L_planta_df$Consortia_condition <- factor(mean_L_planta_df$Consortia_condition, levels = c("NI", "C1","C1P", "C1R", "C1RP"))

graph_L_planta <- ggplot(mean_L_planta_df, aes(x = Consortia_condition, y = mean_log10_L_planta, fill = Consortia_condition)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.5) +
  geom_errorbar(aes(ymin = mean_log10_L_planta - sd_log10_L_planta, ymax = mean_log10_L_planta + sd_log10_L_planta),
                width = 0.2,                     
                position = position_dodge(0.7)) + 
  scale_fill_manual(values = c("grey","#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  facet_wrap(~Day,  
             labeller = labeller(Day = c(`0` = "Day 0", `3` = "Day 3", `28` = "Day 28")))+
  labs(title = "",
       x = "",
       y = "Lactiplantibacillus plantarum \n log(10)CFU/g",
       fill = "Consortia") +
  scale_y_continuous(breaks = seq(0, 9, 0.5)) +
  coord_cartesian(ylim = c(2,9)) +
  
  # Horizontal line for Day 0 to y = 2
  geom_hline(data = subset(mean_L_planta_df, Day == 0), 
             aes(yintercept = 2), 
             linetype = "dashed", color = "red") +
  
  # Horizontal line for Day 3 and Day 28 at y = 3.3
  geom_hline(data = subset(mean_L_planta_df, Day %in% c(3, 28)), 
             aes(yintercept = 5), 
             linetype = "dashed", color = "red") +

  theme_bw() +
  theme(panel.background = element_rect(fill="white"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        strip.text = element_text(size = 12))

graph_L_planta
```

Viable counts of Lactiplantibacillus plantarum CIRM-BIA 2443.Bar indicates the average value of four replicates (independent laboratory-scale productions) in log CFU/g of vegetables, with error bars representing the standard deviation. The dashed red line indicates the detection limit for microbial enumeration (2 CFU/g for day 0 and 5 CFU/g for day 3 and 28).

## Representation of the phage PLE001 titration during the fermentation

### Barplot representation of phage PLE001 titration during fermentation

```{r bareplot PLE001, message=FALSE, warning=FALSE}
mean_PLE001_df$Consortia_condition <- factor(mean_PLE001_df$Consortia_condition, levels = c("NI", "C1","C1P", "C1R", "C1RP"))


graph_PLE001 <- ggplot(mean_PLE001_df, aes(x = Consortia_condition, y = mean_log10_PLE001, fill = Consortia_condition)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.5) +
  geom_errorbar(aes(ymin = mean_log10_PLE001 - sd_log10_PLE001, ymax = mean_log10_PLE001 + sd_log10_PLE001),
                width = 0.2,                     
                position = position_dodge(0.7)) + 
  scale_fill_manual(values = c("grey","#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  facet_wrap(~Day,  
             labeller = labeller(Day = c(`0` = "Day 0", `3` = "Day 3", `28` = "Day 28")))+
  labs(title = "",
       x = "",
       y = "Leuconostoc phage PLE001 \n log(10)PFU/g",
       fill = "Consortia") +
scale_y_continuous(breaks = seq(0, 6, 0.5)) +
  coord_cartesian(ylim = c(2,6))+
  geom_hline(aes(yintercept = 3), linetype = "dashed", color = "red")+
  theme_bw()+
  theme(panel.background = element_rect(fill="white"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        strip.text = element_text(size = 12))

  
  
graph_PLE001

```

Concentration of Leuconostoc phage PLE001 in log PFU/g during carrot fermentation. The dotted red bar represents the limit of detection for phage titration set at 3.1 log CFU/g.

### Phage production dynamics between D0 and D3 of fermentation consortia infected with phage PLE001. 

```{r plot PLE001, message=FALSE, warning=FALSE}
data_filtered <- mean_PLE001_df %>%
  filter(Consortia_condition %in% c("C1P", "C1RP")) %>%
  select(Consortia_condition, Day, mean_log10_PLE001, sd_log10_PLE001) %>%
  pivot_wider(names_from = Day, values_from = c(mean_log10_PLE001, sd_log10_PLE001)) %>%
  mutate(
    delta_log10_PLE001 = `mean_log10_PLE001_3` - `mean_log10_PLE001_0`, 
    delta_sd = sqrt(`sd_log10_PLE001_3`^2 + `sd_log10_PLE001_0`^2) 
  ) %>%
  select(Consortia_condition, delta_log10_PLE001, delta_sd)


# Creating a chart with standard deviations
delta_log_J3 <- ggplot(data_filtered, aes(x = Consortia_condition, y = delta_log10_PLE001, color = Consortia_condition)) +
  geom_segment(aes(x = Consortia_condition, xend = Consortia_condition, y = 0, yend = delta_log10_PLE001), color = "gray") +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = delta_log10_PLE001 - delta_sd, ymax = delta_log10_PLE001 + delta_sd), width = 0.2) +  
  scale_color_manual(values = c("C1P" = "#1f78b4",
                                "C1RP" = "#33a02c")) +
  theme_minimal() +
  labs(y = "Δlog10 (PFU/g)", x = "",
       title = "") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  theme(legend.position = "none",
        plot.background = element_rect(fill = "white", color = NA),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 20))
delta_log_J3


setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

file_name <- paste0("results/graph/", Sys.Date(), "_Figure_6D.svg")
ggsave(file_name, plot = delta_log_J3, device = "svg", width = 10, height = 8, create.dir = TRUE)

```

Phage production dynamics between D0 and D3 of fermentation consortia infected with phage PLE001. The ∆log10 represents the ratio between PFU counts at D0 and D3.


## Statistical analysis about the phage effect on the microbial community

The test answers the following biological question: ‘Are there significant differences in microbial community during fermentation between samples with and without phage, for a specific species and on a given day?’ The Wilcoxon test compares the medians of the two groups (with and without phage).

```{r Wilcoxon test function, message=FALSE, warning=FALSE}
# Function to perform the Wilcoxon test for a given variable
test_wilcoxon <- function(df, consortia, sampling_day, variable) {
  data_temp <- df %>% filter(Day == sampling_day, Consortia == consortia)
  print(nrow(data_temp)) 
  # Check if Phage has exactly two levels
  if (length(unique(data_temp$Phage)) != 2) {
    print(paste("There are not two levels for Phage in", consortia, "on day", sampling_day))
    return(NA)
  }
  
  # Select the variable dynamically
  data_temp <- data_temp %>% select(Phage, value = {{ variable }})

  # Check and convert to numeric if necessary
  if (!is.numeric(data_temp$value)) {
    data_temp$value <- as.numeric(data_temp$value)
  }

  # Wilcoxon test
  test_result <- wilcox.test(value ~ Phage, data = data_temp)

  p_value <- test_result$p.value
  print(paste("Wilcoxon test | Consortia:", consortia, "| Day:", sampling_day, "| Variable:", deparse(substitute(variable)), "| p-value:", p_value))  # Debug

  return(p_value)
}
```

```{r Statistical tests, message=FALSE, warning=FALSE}

# Retirer le consortium "NI" de all_rep
all_rep_filtered <- Results_table %>% filter(Consortia != "NI")

# Vérification des niveaux de Phage avant tout
all_rep_filtered$Phage <- factor(all_rep_filtered$Phage, levels = c("Yes", "No"))

# List of days and consortia
jours <- unique(all_rep_filtered$Day)
consortia_list <- unique(all_rep_filtered$Consortia)

resultats_RA <- expand.grid(Consortia = consortia_list, Day = jours) %>%
  rowwise() %>%
  mutate(
    p_value = test_wilcoxon(all_rep_filtered, Consortia, Day, !!sym("R_aquatilis_.CFU.g.")),
    test = "Wilcoxon",
    variable_tested = "Phages-L.mesenteroides enumeration"
  ) %>%
  ungroup()

resultats_LP <- expand.grid(Consortia = consortia_list, Day = jours) %>%
  rowwise() %>%
  mutate(
    p_value = test_wilcoxon(all_rep_filtered, Consortia, Day, !!sym("L_plantatarum_.CFU.g.")),
    test = "Wilcoxon",
    variable_tested = "Phages-L.plantarum enumeration"
  ) %>%
  ungroup()


resultats_LM <- expand.grid(Consortia = consortia_list, Day = jours) %>%
  rowwise() %>%
  mutate(
    p_value = test_wilcoxon(all_rep_filtered, Consortia, Day, !!sym("L_mesenteroides_.CFU.g.")),
    test = "Wilcoxon",
    variable_tested = "Phages-L.mesenteroides enumeration"
  ) %>%
  ungroup()



# Display results
print(resultats_RA)
print(resultats_LP)
print(resultats_LM)


```

The only statistically significant effect of phage addition on bacterial population levels was observed on day 3 in the viable counts of *Leuconostoc mesenteroides* in consortium C1 (Wilcoxon test, p \< 0.05).
