---
title: "Analysis of titratable acidity, pH and lactic acid measurement"
author: "Julia Cantuti Gendre"
date: "2025-01-27"
output: html_document
---

## Description

This script calculates the mean and standard deviation of pH and acidity measurements for each sample in each replicate in order to plot them on a graph, and then performs statistical tests to highlight the effect of phages on these variables. This script then analyzes the biochemical measurement of lactic acid during fermentation and performs statistical tests to identify any potential effect of the phage on lactic acid production.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import library,message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(readr)
library(dplyr)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(purrr)
options(bitmapType = "cairo")
sessionInfo()
```

## Importing data

```{r import data, message=FALSE, warning=FALSE}
setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

Results_table <- read.table("results/experimental/Experimental_results.csv", header = TRUE, sep = ",", dec=",")

table(Results_table$Samples)

```

## Calculation of mean acidity and standard deviation of each consortium for each day

```{r calcul mean and standard deviation acidity, message=FALSE, warning=FALSE}
mean_Ac_df <- Results_table %>%
  group_by(Consortia_condition, Day,Phage) %>%
  summarise(
    mean_Ac = mean(Acidity_perc, na.rm = TRUE),   # Calculer la moyenne du pH
    sd_Ac = sd(Acidity_perc, na.rm = TRUE),       # Calculer l'écart-type du pH
    .groups = 'drop'
  )
mean_Ac_df

```

## Graph showing titratable acidity during carrot fermentation

```{r plot acidity, message=FALSE, warning=FALSE}
Results_table$Consortia_condition <- factor(Results_table$Consortia_condition, levels = c("NI", "C1","C1P","C1R", "C1RP"))


consortia_colors <- c("NI"   = "grey",
                      "C1"   = "#a6cee3",
                      "C1P"  = "#1f78b4",
                      "C1R"  = "#b2df8a",
                      "C1RP" = "#33a02c")


# Création du boxplot avec les points colorés par consortium
AC_boxplot <- ggplot(Results_table, aes(x = as.factor(Day), y = Acidity_perc)) +
  geom_boxplot(fill = "white", color = "black", outlier.shape = NA) +  # Boxplot sans outliers
  geom_jitter(aes(fill = Consortia_condition),
              width = 0.2, size = 3, shape = 21, color = "black", stroke = 1)+  # Points avec jitter pour éviter le chevauchement
  scale_fill_manual(values = consortia_colors)+
  theme_minimal() +
  labs(x = "Time (day)", y = "Titratable acidity\n (g lactic acid/100mL)", title = "") +
  theme(legend.title = element_blank())  # Supprime le titre de la légende
AC_boxplot


#ggsave("mean_acidity_J0.png", mean_acidity_J0, width = 5.8, height = 3.5, dpi = 1000)

```

## Calculation of mean pH and standard deviation of each consortium for each day

```{r calcul mean and standard deviation pH, message=FALSE, warning=FALSE}
mean_pH_df <- Results_table %>%
  group_by(Consortia_condition, Day,Phage) %>%
  summarise(
    mean_pH = mean(pH, na.rm = TRUE),   # Calculer la moyenne du pH
    sd_pH = sd(pH, na.rm = TRUE),       # Calculer l'écart-type du pH
    .groups = 'drop'
  )
mean_pH_df

```

## Graph showing pH during carrot fermentation

```{r plot pH, message=FALSE, warning=FALSE}
Results_table$Consortia_condition <- factor(Results_table$Consortia_condition,
                                            levels = c("NI", "C1", "C1P", "C1R", "C1RP"))

consortia_colors <- c("NI"   = "grey",
                      "C1"   = "#a6cee3",
                      "C1P"  = "#1f78b4",
                      "C1R"  = "#b2df8a",
                      "C1RP" = "#33a02c")

pH_boxplot <- ggplot(Results_table, aes(x = as.factor(Day), y = pH)) +
  geom_boxplot(fill = "white", color = "black", outlier.shape = NA) +
  geom_jitter(aes(fill = Consortia_condition),
              width = 0.2, size = 3, shape = 21, color = "black", stroke = 1) +
  scale_fill_manual(values = consortia_colors) +
  theme_minimal() +
  labs(x = "Time (day)", y = "pH", title = "", fill = "Consortia condition")+
  theme(legend.title = element_blank())  # Supprime le titre de la légende

pH_boxplot
```

#### Grouping plots

```{r plot combined acidity, message=FALSE, warning=FALSE}
Acidity_group_BP <- ggarrange(pH_boxplot,AC_boxplot,labels = c("A", "B"),
  common.legend = TRUE,
  legend = "right",
  ncol = 2)+
  theme(plot.background = element_rect(fill = "white", color = NA),
           plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))
Acidity_group_BP 

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

file_name <- paste0("results/graph/", Sys.Date(), "_Figure_6AB.svg")
ggsave(file_name, plot = Acidity_group_BP, device = "svg", width = 10, height = 5, create.dir = TRUE)

```

A)  & B) Boxplots represent the distribution of the pH (A) and titratable acidity (B) values of each replicate for each of the four tested consortia and the control, in the absence ( ) and presence (⦁) of phage. Titratable acidity is expressed in grams of lactic acid equivalent per 100 mL of exudate.

## Statistic test about the phage effect on acidity during the fermentation

This test was used to check whether phages significantly modify pH or acidity at different stages of fermentation for each consortium. A significant p-value would suggest that phages influence the acid-base dynamics of the fermentation system.

```{r Wilcoxon test function , message=FALSE, warning=FALSE}

test_wilcoxon <- function(df, consortia, sampling_day, variable) {
  data_temp <- df %>% filter(Day == sampling_day, Consortia == consortia)
  
  # Check if Phage has exactly two levels
  if (length(unique(data_temp$Phage)) != 2) {
    print(paste("There are not two levels for Phage in", consortia, "on day", sampling_day))
    return(NA)
  }
  
  # Select the variable dynamically
  data_temp <- data_temp %>% select(Phage, value = {{ variable }})

  # Check and convert to numeric if necessary
  if (!is.numeric(data_temp$value)) {
    data_temp$value <- as.numeric(data_temp$value)
  }

  # Wilcoxon test
  test_result <- wilcox.test(value ~ Phage, data = data_temp)

  p_value <- test_result$p.value
  print(paste("Wilcoxon test | Consortia:", consortia, "| Day:", sampling_day, "| Variable:", deparse(substitute(variable)), "| p-value:", p_value))  # Debug

  return(p_value)
}

```

```{r Statistic test of phage impact on pH and acidity, message=FALSE, warning=FALSE}
# Delete NI samples from dataframe
all_rep_filtered <- Results_table %>% filter(Consortia != "NI")

# Check phage level
all_rep_filtered$Phage <- factor(all_rep_filtered$Phage, levels = c("Yes", "No"))

# List of days and consortia
days <- unique(all_rep_filtered$Day)
consortia_list <- unique(all_rep_filtered$Consortia)

resultats_pH_phage <- expand.grid(Consortia = consortia_list, Day = days) %>%
  rowwise() %>%
  mutate(Variable = "pH",result = list(test_wilcoxon(all_rep_filtered, Consortia, Day, !!sym("pH"))))%>% 
  unnest(result, names_sep = "_") %>%  
  ungroup()

resultats_AC_phage <- expand.grid(Consortia = consortia_list, Day = days) %>%
  rowwise() %>%
  mutate(Variable = "Acidity_perc",result = list(test_wilcoxon(all_rep_filtered, Consortia, Day, !!sym("Acidity_perc"))))%>% 
  unnest(result, names_sep = "_") %>%  
  ungroup()

```

No significant effect of phage PLE001 addition on pH or titratable acidity values was observed (Wilcoxon test, p \> 0.05).

## Calculation of mean and standard deviation of lactic acid concentration of each consortium for each day

```{r calcul mean and standard deviation bacteria, message=FALSE, warning=FALSE}
mean_Lactic_acid_df <- Results_table %>%
  group_by(Consortia_condition, Day) %>%
  summarise(
    mean_LA = mean(Lactic_acid_.g.L., na.rm = TRUE),
    sd_LA = sd(Lactic_acid_.g.L., na.rm = TRUE),
    .groups = 'drop'
  )
mean_Lactic_acid_df

mean_Lactic_acid_df_day <- Results_table %>%
  group_by( Day) %>%
  summarise(
    mean_LA = mean(Lactic_acid_.g.L., na.rm = TRUE),
    sd_LA = sd(Lactic_acid_.g.L., na.rm = TRUE),
    .groups = 'drop'
  )
mean_Lactic_acid_df_day
```

## Representation of the lactic acid concentration for each consortia during the fermentation

```{r plot Lactic acid, message=FALSE, warning=FALSE}
mean_Lactic_acid_df$Consortia_condition <- factor(mean_Lactic_acid_df$Consortia_condition, levels = c("NI", "C1","C1P","C1R", "C1RP"))


mean_Lactic_acid_ph <- ggplot(mean_Lactic_acid_df, aes(x = Consortia_condition, y = mean_LA, fill = Consortia_condition)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.5) +
  geom_errorbar(aes(ymin = mean_LA - sd_LA, ymax = mean_LA + sd_LA),
                width = 0.2,                     
                position = position_dodge(0.7)) + 
  scale_fill_manual(values = c("grey","#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  labs(title = "",
       x = "",
       y = "Lactic acid \n (g/L)") +
scale_y_continuous(breaks = seq(0, 20, 5)) +
  coord_cartesian(ylim = c(0,20))+
  facet_wrap(~Day,  
             labeller = labeller(Day = c(`0` = "Day 0", `3` = "Day 3", `28` = "Day 28")))+
  theme_bw()+
  theme(panel.background = element_rect(fill="white"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        strip.text = element_text(size = 12))+
    geom_hline(data = filter(mean_Lactic_acid_df, Day == 0), aes(yintercept = 3.23), linetype = "dashed", color = "red") +
  geom_hline(data = filter(mean_Lactic_acid_df, Day == 3), aes(yintercept = 3.23), linetype = "dashed", color = "red") +
  geom_hline(data = filter(mean_Lactic_acid_df, Day == 28), aes(yintercept = 3.23), linetype = "dashed", color = "red")
mean_Lactic_acid_ph

```

Evolution of lactic acid production during carrot fermentation. Concentrations are expressed in g/L of exudate.

## Statistic test about the phage effect on lactic acid concentration

The test answers the following biological question: ‘Are there significant differences in lactic acid concentration during fermentation between samples with and without phage, for a specific consortium and on a given day?’ The Wilcoxon test compares the medians of the two groups (with and without phage).

```{r Statistic for lactic acid, message=FALSE, warning=FALSE}
# Retirer le consortium "NI" de all_rep
all_rep_filtered <- Results_table %>% filter(Consortia != "NI")

# Vérification des niveaux de Phage avant tout
all_rep_filtered$Phage <- factor(all_rep_filtered$Phage, levels = c("Yes", "No"))

# List of days and consortia
jours <- unique(all_rep_filtered$Day)
consortia_list <- unique(all_rep_filtered$Consortia)

resultats_LA_phage <- expand.grid(Consortia = consortia_list, Day = jours) %>%
  rowwise() %>%
  mutate(Variable = "Lactic_acid_.g.L.",result = list(test_wilcoxon(all_rep_filtered, Consortia, Day, !!sym("Lactic_acid_.g.L."))))%>% 
  unnest(result, names_sep = "_") %>%  
  ungroup()
```

The addition of phage PLE001 did not significantly affect lactic acid production in all consortia (Wilcoxon test, p \> 0.05).
