---
title: "pH & Acidity distribution"
author: "Julia Gendre"
date: "2023-11-09"
output: html_document
---

### Description

XXXXX

### Packages loading

```{r directory, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Redaction/Phage_dynamic_fermented_food")
```

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(ggpubr)
library(rstatix)
library(gridExtra)
options(bitmapType = "cairo")
sessionInfo()
```

### Data import
```{r Import data, message=FALSE, warning=FALSE}
setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

Fermentation_metadata <- read.csv("results/pH_acidity_microbial_results.txt", sep = "\t",dec= ",",header = TRUE,
                                    fileEncoding = "latin1")

Fermentation_metadata$Day_sampling [Fermentation_metadata$Day_sampling=="22"]<- "21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T0"]<- "Day 0"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T2"]<- "Day 2"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T7"]<- "Day 7"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T14"]<- "Day 14"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T21"]<- "Day 21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T22"]<- "Day 21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T28"]<- "Day 28"


Fermentation_metadata$Day_sampling <- as.numeric(Fermentation_metadata$Day_sampling)

Fermentation_metadata$Day_sampling_text <- factor(Fermentation_metadata$Day_sampling_text, levels = c("Day 0", "Day 2", "Day 7","Day 14","Day 21","Day 28"))
```

## Dynamic of pH during vegetables' fermentation

#### Calculation of mean pH and standard deviation of each sampling day

```{r}
mean_pH_df <- Fermentation_metadata %>%
  group_by(Matrix, Day_sampling_text) %>%
  summarise(
    mean_acidity = mean(pH, na.rm = TRUE),   # Calculer la moyenne du pH
    sd_acidity = sd(pH, na.rm = TRUE),       # Calculer l'écart-type du pH
    .groups = 'drop'
  )
mean_pH_df
```

#### Representation of the pH curve during vegetables' fermentation

```{r pH distribution, message=FALSE, warning=FALSE}
color_matrix <- c('Cabbage'='darkolivegreen4','Carrot'='darkorange')

# Convertir Matrix en facteur si ce n'est pas déjà le cas
Fermentation_metadata$Matrix <- factor(Fermentation_metadata$Matrix)

# Vérification des niveaux
print(levels(Fermentation_metadata$Matrix))


pH_graph2 <- ggplot(Fermentation_metadata, aes(x = Day_sampling, y = pH, colour = Matrix)) +
  geom_point() +
  geom_smooth(method = "loess", level = 0.95) +
  scale_x_continuous(breaks = unique(Fermentation_metadata$Day_sampling)) +
  labs(x = "Time (day)",
       y = "pH",
       colour = "Matrix") +
  scale_colour_manual(values = c("Cabbage" = "darkolivegreen4", "Carrot" = "darkorange")) + # Couleurs personnalisées
  theme_minimal()+
  theme(axis.title = element_text(size=16),
        axis.text = element_text(size=15),
        legend.text = element_text(size=16),
        legend.title = element_text(size = 18))
pH_graph2


### Statistic
mean_pH_carrot <- Fermentation_metadata %>%
  filter(Matrix == "carrot") %>%    # Sélectionner uniquement les carottes
  group_by(Day_sampling) %>%               # Grouper les données par jour
  summarise(mean_pH = mean(pH))

mean_pH_cabbage <- Fermentation_metadata %>%
  filter(Matrix == "cabbage") %>%    # Sélectionner uniquement les carottes
  group_by(Day_sampling) %>%               # Grouper les données par jour
  summarise(mean_pH = mean(pH))
```

## Dynamic of titrable acidity during vegetables' fermentation

#### Calculation of mean titrable acidity and standard deviation of each sampling day

```{r}
mean_acidity_df <- Fermentation_metadata %>%
  group_by(Matrix, Day_sampling_text) %>%
  summarise(
    mean_acidity = mean(Acidity, na.rm = TRUE),   # Calculer la moyenne du pH
    sd_acidity = sd(Acidity, na.rm = TRUE),       # Calculer l'écart-type du pH
    .groups = 'drop'
  )
mean_acidity_df
```

#### Representation of the titrable acidity curve during vegetables' fermentation

```{r acidity distribution, message=FALSE, warning=FALSE}
Ac_graph <- ggplot(Fermentation_metadata, aes(x = Day_sampling, y = Acidity, colour = Matrix)) +
  geom_point() +
  geom_smooth(method = "loess", level = 0.95) +
  scale_x_continuous(breaks=Fermentation_metadata$Day_sampling)+
  labs(x = "Time (day)",
       y = "Titratable acidity\n (g lactic acid/100mL)",
       colour = "Matrix") +
  scale_colour_manual(values = c("Cabbage" = "darkolivegreen4", "Carrot" = "darkorange")) + # Couleurs personnalisées
  theme_minimal()+
  theme(axis.title = element_text(size=16),
        axis.text = element_text(size=15),
        legend.text = element_text(size=16),
        legend.title = element_text(size = 18))
Ac_graph

### Statistic
mean_Ac_carrot <- Fermentation_metadata %>%
  filter(Matrix == "carrot") %>%    # Sélectionner uniquement les carottes
  group_by(Day_sampling) %>%               # Grouper les données par jour
  summarise(mean_Ac = mean(Acidity))

mean_Ac_cabbage <- Fermentation_metadata %>%
  filter(Matrix == "cabbage") %>%    # Sélectionner uniquement les carottes
  group_by(Day_sampling) %>%               # Grouper les données par jour
  summarise(mean_Ac = mean(Acidity))

```

##### Combining the graphs for the final figure

```{r acidity & pH distribution, message=FALSE, warning=FALSE}

pH_acid_combin <- ggarrange(pH_graph2, Ac_graph,labels = c("A", "B"),
  common.legend = TRUE,
  legend = "right",
  label.y = 1.00)+
  theme(plot.background = element_rect(fill = "white", color = NA),
           plot.margin = margin(1, 1, 1, 1, "cm"))
pH_acid_combin

ggsave("pH_and_acidity_distribution.tiff",pH_acid_combin,width = 10, height = 6, dpi = 300)
```

## Representation of microbial enumeration during fermentation

```{r Microbial enumeration, message=FALSE, warning=FALSE}

metadata_log <-  Fermentation_metadata
metadata_log$MRS_enumeration <- as.numeric(Fermentation_metadata$MRS_enumeration)
metadata_log$MRS <- log10(metadata_log$MRS_enumeration)
metadata_log$PCA_enumeration <- as.numeric(Fermentation_metadata$PCA_enumeration)
metadata_log$PCA <- log10(metadata_log$PCA_enumeration)
metadata_log$YEGC_enumeration <- as.numeric(Fermentation_metadata$YEGC_enumeration)
metadata_log$YEGC <- log10(metadata_log$YEGC_enumeration)

metadata_log_cabbage <- subset(metadata_log, subset=metadata_log$Matrix == 'Cabbage')
metadata_long_cabbage <- pivot_longer(metadata_log_cabbage, 
                              cols = c(MRS, PCA, YEGC),
                              names_to = "Enumeration_Type", 
                              values_to = "Enumeration_Log")


B_cabbage <- ggplot(metadata_long_cabbage, aes(x = Day_sampling, y = Enumeration_Log, shape = Enumeration_Type)) +
  geom_jitter(aes(fill = "fixed"), width = 0.2, size = 3, color = "black", stroke = 1) +  # Contour noir + remplissage coloré
  geom_line(aes(group = interaction(Matrix, Enumeration_Type), color = "fixed"), 
            stat = "summary", fun = mean) + 
  geom_errorbar(stat = "summary", fun.data = function(x) {
    data.frame(ymin = mean(x) - sd(x), ymax = mean(x) + sd(x))
 }, width = 0.1, aes(color = "fixed")) +  
  scale_x_continuous(breaks = unique(metadata_long_carrot$Day_sampling)) +
  scale_color_manual(values = "darkolivegreen4", guide = "none") +  
  scale_fill_manual(values = "darkolivegreen4", guide = "none") +  # Remplissage orange
  scale_shape_manual(values = c(21, 22, 24)) +  # Formes qui acceptent `fill`
  labs(x = "Time (day)", y = "Log (CFU/g)") +
  theme_minimal() +
  theme(axis.title = element_text(size=16),
        axis.text = element_text(size=15),
        legend.text = element_text(size=16),
        legend.title = element_blank())
B_cabbage


metadata_log_carrot <- subset(metadata_log, subset=metadata_log$Matrix == 'Carrot')
metadata_long_carrot <- pivot_longer(metadata_log_carrot, 
                              cols = c(MRS, PCA, YEGC),
                              names_to = "Enumeration_Type", 
                              values_to = "Enumeration_Log")

B_carrot <- ggplot(metadata_long_carrot, aes(x = Day_sampling, y = Enumeration_Log, shape = Enumeration_Type)) +
  geom_jitter(aes(fill = "fixed"), width = 0.2, size = 3, color = "black", stroke = 1) +  # Contour noir + remplissage coloré
  geom_line(aes(group = interaction(Matrix, Enumeration_Type), color = "fixed"), 
            stat = "summary", fun = mean) + 
  geom_errorbar(stat = "summary", fun.data = function(x) {
    data.frame(ymin = mean(x) - sd(x), ymax = mean(x) + sd(x))
  }, width = 0.1, aes(color = "fixed")) +  
  scale_x_continuous(breaks = unique(metadata_long_carrot$Day_sampling)) +
  scale_color_manual(values = "darkorange", guide = "none") +  
  scale_fill_manual(values = "darkorange", guide = "none") +  # Remplissage orange
  scale_shape_manual(values = c(21, 22, 24)) +  # Formes qui acceptent `fill`
  labs(x = "Time (day)", y = "Log (CFU/g)") +
  theme_minimal() +
  theme(axis.title = element_text(size=16),
        axis.text = element_text(size=15),
        legend.text = element_text(size=16),
        legend.title = element_blank())

B_carrot




Microbio_combin <- ggarrange(B_cabbage, B_carrot, 
                            labels = c("Cabbage", "Carrot"),
                            common.legend = TRUE,
                            legend = "right",
                            ncol = 2,
                            label.y = 1) +  # Ajuste la position verticale des labels
  theme(plot.background = element_rect(fill = "white", color = NA),
           plot.margin = margin(1.2, 1.2, 1.2, 1.2, "cm"))

Microbio_combin



#ggsave("microbio_combined.tiff",Microbio_combin,width = 8, height = 6, dpi = 300)

```

#### Microbial enumeration statistics

```{r Microbial statistics, message=FALSE, warning=FALSE}
mean_MRS_df <- metadata_log %>%
  group_by(Matrix,Day_sampling_text) %>%
  summarise(
    mean_acidity = mean(MRS, na.rm = TRUE),   # Calculer la moyenne du pH
    sd_acidity = sd(MRS, na.rm = TRUE),       # Calculer l'écart-type du pH
    .groups = 'drop'
  )
mean_MRS_df

mean_yeast_df <- metadata_log %>%
  group_by(Matrix, Day_sampling_text) %>%
  summarise(
    mean_acidity = mean(YEGC, na.rm = TRUE),   # Calculer la moyenne du pH
    sd_acidity = sd(YEGC, na.rm = TRUE),       # Calculer l'écart-type du pH
    .groups = 'drop'
  )
mean_yeast_df

mean_MRS_carrot <- metadata_log_carrot %>%   
  group_by(Day_sampling) %>%               # Grouper les données par jour
  summarise(mean_MRS = mean(MRS))
mean_MRS_carrot

mean_MRS_cabbage <- metadata_log_cabbage %>%   
  group_by(Day_sampling) %>%               # Grouper les données par jour
  summarise(mean_MRS = mean(MRS))
mean_MRS_carrot
```


## VLP estimation by Epifluorescence microscopy

```{r}
metadata_log$VLP_estimation <- as.numeric(Fermentation_metadata$VLP_estimation)
metadata_log$VLP <- log10(metadata_log$VLP_estimation)

# Convertir Matrix en facteur si ce n'est pas déjà le cas
Fermentation_metadata$Matrix <- factor(Fermentation_metadata$Matrix)

# Vérification des niveaux
print(levels(Fermentation_metadata$Matrix))

mean_VLP_df <- metadata_log %>%
  group_by(Matrix, Day_sampling_text) %>%
  summarise(
    mean_VLP = mean(VLP, na.rm = TRUE),   # Calculer la moyenne du pH
    sd_VLP = sd(VLP, na.rm = TRUE),       # Calculer l'écart-type du pH
    .groups = 'drop'
  )
mean_VLP_df
```

#### Representation of the VLP estimation during vegetables' fermentation

```{r VLP distribution, message=FALSE, warning=FALSE}
color_matrix <- c('Cabbage'='darkolivegreen4','Carrot'='darkorange')

# Création du graphique
VLP_graph <- ggplot(metadata_log, aes(x = Day_sampling, y = VLP, colour = Matrix)) +
  geom_point() +
  geom_smooth(method = "loess", level = 0.95) +
  scale_x_continuous(breaks = unique(metadata_log$Day_sampling)) +
  labs(x = "Time (day)",
       y = "VLP estimation log(VLP/g)",
       colour = "Matrix") +
  scale_colour_manual(values = c("Cabbage" = "darkolivegreen4", "Carrot" = "darkorange")) + # Couleurs personnalisées
  theme_minimal()+
  theme(plot.margin = margin(2, 2, 2, 2, "cm"),
        axis.title = element_text(size=16),
        axis.text = element_text(size=15),
        legend.text = element_text(size=16),
        legend.title = element_text(size = 18),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA))
  
VLP_graph

```
##### Combining the graphs for the final figure

```{r arrangement pH, acidité, microbio, message=FALSE, warning=FALSE}

pH_all2.2 <- ggarrange(pH_acid_combin, Microbio_combin,VLP_graph, labels = c("", "C","D"),
  common.legend = FALSE,
  nrow = 3,
  label.x = 0.03)
pH_all2.2

setwd("~/Documents/Redaction/Phage_dynamic_fermented_food")

file_name <- paste0("results/graph/", Sys.Date(), "_Figure_2.svg")
ggsave(file_name, plot = pH_all2.2, device = "svg", width = 15, height = 15, create.dir = TRUE)

```

