---
title: "Metagenome_analyses"
author: "Julia Gendre"
date: "2024-10-28"
output: html_document
header-includes:
  - \usepackage{fvextra}
  - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.show = "asis",message = FALSE,warning = FALSE, tidy = FALSE )
knitr::opts_knit$set(root.dir = "/work_projet/metasimfood/metavirome/Phage_dynamic_fermented_food/")
```


## Packages loading

```{r packages, message=FALSE, warning=FALSE}
library(phyloseq)
library(readxl)  
library(phyloseq.extended)
library(vegan)
library(gridExtra)
library(rstatix)
library(gplots)
library(tidyverse)
library(DT)
library(pheatmap)
library(DESeq2)
library(ggplotify)
library(cowplot)
library(ggpubr)
library(ggforce)
library(ggplot2)
library(ggrepel)
library(ggbreak)
library(VennDiagram)
library(grid)
library(factoextra)
library(pairwiseAdonis)
library(Hmisc)
library(dplyr)
sessionInfo()

```
# Import fermentation metadata
```{r Import data, message=FALSE, warning=FALSE}
Fermentation_metadata <- read.table("Results_table.csv", sep = ',',dec= ".",header = TRUE)
Fermentation_metadata$Day_sampling [Fermentation_metadata$Day_sampling=="22"]<- "21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T0"]<- "Day 0"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T2"]<- "Day 2"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T7"]<- "Day 7"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T14"]<- "Day 14"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T21"]<- "Day 21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T22"]<- "Day 21"
Fermentation_metadata$Day_sampling_text [Fermentation_metadata$Day_sampling_text=="T28"]<- "Day 28"


Fermentation_metadata$Day_sampling <- as.numeric(Fermentation_metadata$Day_sampling)

Fermentation_metadata$Day_sampling_text <- factor(Fermentation_metadata$Day_sampling_text, levels = c("Day 0", "Day 2", "Day 7","Day 14","Day 21","Day 28"))

sample_metag_metadata <- read.table("METADATA/Sample_metadata.txt",sep="\t",header = TRUE)%>%
  column_to_rownames(.,var="Samples")

```

# Kaiju data analyses
## Function creation to treat Kaiju data
```{r function, message=FALSE, warning=FALSE}
# Function to treat kaiju data 
process_kaiju_data <- function(data, Samples_name) {
  # Add colonne 'Samples' to raw data vith Samples_name variable and select column : percentage of reads belong to taxa, number of reads belongt to taxa, taxon_id et taxon_name.
  data <- data %>%
    mutate(Samples = Samples_name) %>%
    select(Samples, percent, reads, taxon_id, taxon_name)
  
  # split taxon_name column according to different taxonomic rank.
  data <- data %>%
    separate(taxon_name, into = c("superkingdom", "phylum", "class", "order", "family", "genus", "species"), sep = ";")
  
  # Apply conditional modifications using `mutate` and `case_when`.
  data <- data %>%
    mutate(
      superkingdom = case_when(
        superkingdom == "cannot be assigned to a (non-viral) genus" ~ "unclassified",
        TRUE ~ superkingdom
      ),
      genus = case_when(
        superkingdom == "unclassified" ~ "unclassified",
        TRUE ~ genus
      ),
      family = case_when(
        superkingdom == "unclassified" ~ "unclassified",
        TRUE ~ family
      ),
      order = case_when(
        superkingdom == "unclassified" ~ "unclassified",
        TRUE ~ order
      )
    )
  
  # Filtrate Bacteria and Eukaryota data
  bacteria_data <- filter(data, superkingdom == "Bacteria")
  euka_data <- filter(data, superkingdom == "Eukaryota")
  viruses_data <- filter(data, superkingdom == "Viruses")
  
  # The function returns processed data and the filtered subsets.
  return(list(
    processed_data = data,
    bacteria = bacteria_data,
    eukaryota = euka_data,
    viruses = viruses_data
  ))
}

#Function to summary of domain taxon for all Samples
abundance_summary_ki <- function(data, group_column, value_column, new_column_name) {
  # Summary of abundance
  summary_data <- data %>%
    select({{group_column}}, {{value_column}}) %>%
    group_by({{group_column}}) %>%
    summarise(total_count = sum({{value_column}}, na.rm = TRUE))
  
  # Rename columns
  names(summary_data) <- c(as.character(substitute(group_column)), new_column_name)
  
  return(summary_data)
}

# Function to summary genus taxon for all Samples  
summarize_genus_abundance <- function(data, new_column_name, genus_col = "genus", reads_col = "reads") {
  
  # Selecte column
  summary <- data %>%
    select(!!sym(genus_col), !!sym(reads_col)) %>%
    
    # Group by 'genus' and sum 'reads'
    group_by_at(genus_col) %>%
    summarise(total_count = sum(!!sym(reads_col), na.rm = TRUE)) %>%
    ungroup()
  
  # Sum total 'reads'
  column_sum <- sum(summary$total_count, na.rm = TRUE)
  
  # renames columns
  names(summary) <- c("genus", "reads")
  
  # Calcul percentage of reads per genus
  summary <- summary %>%
    mutate(percent = (reads / column_sum) * 100)
  
  # group per genu and sum percentage
  summary_per <- summary %>%
    group_by(genus) %>%
    summarise(total_count = sum(percent, na.rm = TRUE)) %>%
    ungroup()
  
  # rename columns
  names(summary_per) <- c("genus", new_column_name)

  
  # final dataframe
  return(summary_per)
}  
  
# Function to summary species taxon for all Samples  
summarize_species_abundance <- function(data, new_column_name, species_col = "species", reads_col = "reads") {
  
  # Selecte column
  summary <- data %>%
    select(!!sym(species_col), !!sym(reads_col)) %>%
    
    # Group by 'species' and sum 'reads'
    group_by_at(species_col) %>%
    summarise(total_count = sum(!!sym(reads_col), na.rm = TRUE)) %>%
    ungroup()
  
  # Sum total 'reads'
  column_sum <- sum(summary$total_count, na.rm = TRUE)
  
  # renames columns
  names(summary) <- c("species", "reads")
  
  # Calcul percentage of reads per species
  summary <- summary %>%
    mutate(percent = (reads / column_sum) * 100)
  
  # group per genu and sum percentage
  summary_per <- summary %>%
    group_by(species) %>%
    summarise(total_count = sum(percent, na.rm = TRUE)) %>%
    ungroup()
  
  # rename columns
  names(summary_per) <- c("species", new_column_name)

  
  # final dataframe
  return(summary_per)
}
```

## Analyses of the cabbage fermentation
### Import kaiju results from cabbage fermentation
```{r Import data sauerkraut, message=FALSE, warning=FALSE}
SA_0A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_0A_libLAH8279_summary_viruses.tsv",sep="\t",header = TRUE)
SA_0B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_0B_libLAH8280_summary_viruses.tsv",sep="\t",header = TRUE)
SA_0C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_0C_libLAH8281_summary_viruses.tsv",sep="\t",header = TRUE)

SA_2A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_2A_libLAH8282_summary_viruses.tsv", sep="\t",header = TRUE)
SA_2B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_2B_libLAH8283_summary_viruses.tsv", sep="\t",header = TRUE)
SA_2C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_2C_libLAH8284_summary_viruses.tsv", sep="\t",header = TRUE)

SA_7A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_7A_libLAH8285_summary_viruses.tsv", sep="\t",header = TRUE)
SA_7B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_7B_libLAH8286_summary_viruses.tsv", sep="\t",header = TRUE)
SA_7C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_7C_libLAH8287_summary_viruses.tsv", sep="\t",header = TRUE)

SA_14A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_14A_libLAH8288_summary_viruses.tsv", sep="\t",header = TRUE)
SA_14B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_14B_libLAH8289_summary_viruses.tsv", sep="\t",header = TRUE)
SA_14C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_14C_libLAH8290_summary_viruses.tsv", sep="\t",header = TRUE)

SA_21A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_21A_libLAH8291_summary_viruses.tsv", sep="\t",header = TRUE)
SA_21B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_21B_libLAH8292_summary_viruses.tsv", sep="\t",header = TRUE)
SA_21C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_21C_libLAH8293_summary_viruses.tsv", sep="\t",header = TRUE)

SA_28A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_28A_libLAH8294_summary_viruses.tsv", sep="\t",header = TRUE)
SA_28B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_28B_libLAH8295_summary_viruses.tsv", sep="\t",header = TRUE)
SA_28C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_SA_28C_libLAH8296_summary_viruses.tsv", sep="\t",header = TRUE)

res_SA_0A <- process_kaiju_data(SA_0A_Kaiju, "SA_0A")
res_SA_0B <- process_kaiju_data(SA_0B_Kaiju, "SA_0B")
res_SA_0C <- process_kaiju_data(SA_0C_Kaiju, "SA_0C")
res_SA_2A <- process_kaiju_data(SA_2A_Kaiju, "SA_2A")
res_SA_2B <- process_kaiju_data(SA_2B_Kaiju, "SA_2B")
res_SA_2C <- process_kaiju_data(SA_2C_Kaiju, "SA_2C")
res_SA_7A <- process_kaiju_data(SA_7A_Kaiju, "SA_7A")
res_SA_7B <- process_kaiju_data(SA_7B_Kaiju, "SA_7B")
res_SA_7C <- process_kaiju_data(SA_7C_Kaiju, "SA_7C")
res_SA_14A <- process_kaiju_data(SA_14A_Kaiju, "SA_14A")
res_SA_14B <- process_kaiju_data(SA_14B_Kaiju, "SA_14B")
res_SA_14C <- process_kaiju_data(SA_14C_Kaiju, "SA_14C")
res_SA_21A <- process_kaiju_data(SA_21A_Kaiju, "SA_21A")
res_SA_21B <- process_kaiju_data(SA_21B_Kaiju, "SA_21B")
res_SA_21C <- process_kaiju_data(SA_21C_Kaiju, "SA_21C")
res_SA_28A <- process_kaiju_data(SA_28A_Kaiju, "SA_28A")
res_SA_28B <- process_kaiju_data(SA_28B_Kaiju, "SA_28B")
res_SA_28C <- process_kaiju_data(SA_28C_Kaiju, "SA_28C")


SA_0A_Kaiju_process <- res_SA_0A$processed_data
SA_0B_Kaiju_process <- res_SA_0B$processed_data
SA_0C_Kaiju_process <- res_SA_0C$processed_data
SA_2A_Kaiju_process <- res_SA_2A$processed_data
SA_2B_Kaiju_process <- res_SA_2B$processed_data
SA_2C_Kaiju_process <- res_SA_2C$processed_data
SA_7A_Kaiju_process <- res_SA_7A$processed_data
SA_7B_Kaiju_process <- res_SA_7B$processed_data
SA_7C_Kaiju_process <- res_SA_7C$processed_data
SA_14A_Kaiju_process <- res_SA_14A$processed_data
SA_14B_Kaiju_process <- res_SA_14B$processed_data
SA_14C_Kaiju_process <- res_SA_14C$processed_data
SA_21A_Kaiju_process <- res_SA_21A$processed_data
SA_21B_Kaiju_process <- res_SA_21B$processed_data
SA_21C_Kaiju_process <- res_SA_21C$processed_data
SA_28A_Kaiju_process <- res_SA_28A$processed_data
SA_28B_Kaiju_process <- res_SA_28B$processed_data
SA_28C_Kaiju_process <- res_SA_28C$processed_data

abundance_kaiju_cabbage <- full_join (SA_0A_Kaiju_process,SA_0B_Kaiju_process)%>%
  full_join(.,SA_0C_Kaiju_process)%>%
  full_join(.,SA_2A_Kaiju_process)%>%
  full_join(.,SA_2B_Kaiju_process)%>%
  full_join(.,SA_2C_Kaiju_process)%>%
  full_join(.,SA_7A_Kaiju_process)%>%
  full_join(.,SA_7B_Kaiju_process)%>%
  full_join(.,SA_7C_Kaiju_process)%>%
  full_join(.,SA_14A_Kaiju_process)%>%
  full_join(.,SA_14B_Kaiju_process)%>%
  full_join(.,SA_14C_Kaiju_process)%>%
  full_join(.,SA_21A_Kaiju_process)%>%
  full_join(.,SA_21B_Kaiju_process)%>%
  full_join(.,SA_21C_Kaiju_process)%>%
  full_join(.,SA_28A_Kaiju_process)%>%
  full_join(.,SA_28B_Kaiju_process)%>%
  full_join(.,SA_28C_Kaiju_process)

```
### Microbial community at the domain level
Relative abundance for reads for each kingdom were calculated and represented by bar plot.  
```{r sauerkraut kingdom plot, message=FALSE, warning=FALSE}

abundance_SA_0A_kingdom <- abundance_summary_ki(SA_0A_Kaiju_process, superkingdom, percent, "SA_0A")
abundance_SA_0B_kingdom <- abundance_summary_ki(SA_0B_Kaiju_process, superkingdom, percent, "SA_0B")
abundance_SA_0C_kingdom <- abundance_summary_ki(SA_0C_Kaiju_process, superkingdom, percent, "SA_0C")
abundance_SA_2A_kingdom <- abundance_summary_ki(SA_2A_Kaiju_process, superkingdom, percent, "SA_2A")
abundance_SA_2B_kingdom <- abundance_summary_ki(SA_2B_Kaiju_process, superkingdom, percent, "SA_2B")
abundance_SA_2C_kingdom <- abundance_summary_ki(SA_2C_Kaiju_process, superkingdom, percent, "SA_2C")
abundance_SA_7A_kingdom <- abundance_summary_ki(SA_7A_Kaiju_process, superkingdom, percent, "SA_7A")
abundance_SA_7B_kingdom <- abundance_summary_ki(SA_7B_Kaiju_process, superkingdom, percent, "SA_7B")
abundance_SA_7C_kingdom <- abundance_summary_ki(SA_7C_Kaiju_process, superkingdom, percent, "SA_7C")
abundance_SA_14A_kingdom <- abundance_summary_ki(SA_14A_Kaiju_process, superkingdom, percent, "SA_14A")
abundance_SA_14B_kingdom <- abundance_summary_ki(SA_14B_Kaiju_process, superkingdom, percent, "SA_14B")
abundance_SA_14C_kingdom <- abundance_summary_ki(SA_14C_Kaiju_process, superkingdom, percent, "SA_14C")
abundance_SA_21A_kingdom <- abundance_summary_ki(SA_21A_Kaiju_process, superkingdom, percent, "SA_21A")
abundance_SA_21B_kingdom <- abundance_summary_ki(SA_21B_Kaiju_process, superkingdom, percent, "SA_21B")
abundance_SA_21C_kingdom <- abundance_summary_ki(SA_21C_Kaiju_process, superkingdom, percent, "SA_21C")
abundance_SA_28A_kingdom <- abundance_summary_ki(SA_28A_Kaiju_process, superkingdom, percent, "SA_28A")
abundance_SA_28B_kingdom <- abundance_summary_ki(SA_28B_Kaiju_process, superkingdom, percent, "SA_28B")
abundance_SA_28C_kingdom <- abundance_summary_ki(SA_28C_Kaiju_process, superkingdom, percent, "SA_28C")

abundance_kaiju_kingdom_cabb <- full_join(abundance_SA_0A_kingdom,abundance_SA_0B_kingdom)%>%
  full_join(.,abundance_SA_0C_kingdom)%>%
  full_join(.,abundance_SA_2A_kingdom)%>%
  full_join(.,abundance_SA_2B_kingdom)%>%
  full_join(.,abundance_SA_2C_kingdom)%>%
  full_join(.,abundance_SA_7A_kingdom)%>%
  full_join(.,abundance_SA_7B_kingdom)%>%
  full_join(.,abundance_SA_7C_kingdom)%>%
  full_join(.,abundance_SA_14A_kingdom)%>%
  full_join(.,abundance_SA_14B_kingdom)%>%
  full_join(.,abundance_SA_14C_kingdom)%>%
  full_join(.,abundance_SA_21A_kingdom)%>%
  full_join(.,abundance_SA_21B_kingdom)%>%
  full_join(.,abundance_SA_21C_kingdom)%>%
  full_join(.,abundance_SA_28A_kingdom)%>%
  full_join(.,abundance_SA_28B_kingdom)%>%
  full_join(.,abundance_SA_28C_kingdom)

# Convertir en format long pour ggplot2
abundance_long_cabb <- abundance_kaiju_kingdom_cabb %>%
  pivot_longer(cols = -superkingdom, names_to = "Samples", values_to = "abundance")

abundance_long_cabb$Samples <- factor(abundance_long_cabb$Samples, levels = c("SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C"))

abundance_long_cabb$superkingdom <- factor(abundance_long_cabb$superkingdom , levels = c("Archaea","Bacteria","Eukaryota","Viruses","unclassified"))
kingdom_col <- c("Archaea"='#d7191c',"Bacteria"='#fdae61',"Eukaryota"='#ffffbf',"unclassified"='#abd9e9', "Viruses"='#2c7bb6' )

abundance_long_superkingdom_cabbage2 <- left_join(abundance_long_cabb,Fermentation_metadata, by="Samples")


per_read_kingdom_cabb <- ggplot(abundance_long_superkingdom_cabbage2, aes(fill=superkingdom, y=abundance, x=Samples)) +
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Kingdom")+
  scale_fill_manual(values = kingdom_col)+
  theme(legend.text = element_text(size = 11,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 11))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_kingdom_cabb


```
### Viral community at the species level
```{r cabbage viruses plot, message=FALSE, warning=FALSE}

SA_0A_Kaiju_viruses <- res_SA_0A$viruses
SA_0B_Kaiju_viruses <- res_SA_0B$viruses
SA_0C_Kaiju_viruses <- res_SA_0C$viruses
SA_2A_Kaiju_viruses <- res_SA_2A$viruses
SA_2B_Kaiju_viruses <- res_SA_2B$viruses
SA_2C_Kaiju_viruses <- res_SA_2C$viruses
SA_7A_Kaiju_viruses <- res_SA_7A$viruses
SA_7B_Kaiju_viruses <- res_SA_7B$viruses
SA_7C_Kaiju_viruses <- res_SA_7C$viruses
SA_14A_Kaiju_viruses <- res_SA_14A$viruses
SA_14B_Kaiju_viruses <- res_SA_14B$viruses
SA_14C_Kaiju_viruses <- res_SA_14C$viruses
SA_21A_Kaiju_viruses <- res_SA_21A$viruses
SA_21B_Kaiju_viruses <- res_SA_21B$viruses
SA_21C_Kaiju_viruses <- res_SA_21C$viruses
SA_28A_Kaiju_viruses <- res_SA_28A$viruses
SA_28B_Kaiju_viruses <- res_SA_28B$viruses
SA_28C_Kaiju_viruses <- res_SA_28C$viruses


abundance_SA_0A_v_genus_per <- summarize_species_abundance(SA_0A_Kaiju_viruses,"SA_0A")
abundance_SA_0B_v_genus_per <- summarize_species_abundance(SA_0B_Kaiju_viruses,"SA_0B")
abundance_SA_0C_v_genus_per <- summarize_species_abundance(SA_0C_Kaiju_viruses,"SA_0C")
abundance_SA_2A_v_genus_per <- summarize_species_abundance(SA_2A_Kaiju_viruses,"SA_2A")
abundance_SA_2B_v_genus_per <- summarize_species_abundance(SA_2B_Kaiju_viruses,"SA_2B")
abundance_SA_2C_v_genus_per <- summarize_species_abundance(SA_2C_Kaiju_viruses,"SA_2C")
abundance_SA_7A_v_genus_per <- summarize_species_abundance(SA_7A_Kaiju_viruses,"SA_7A")
abundance_SA_7B_v_genus_per <- summarize_species_abundance(SA_7B_Kaiju_viruses,"SA_7B")
abundance_SA_7C_v_genus_per <- summarize_species_abundance(SA_7C_Kaiju_viruses,"SA_7C")
abundance_SA_14A_v_genus_per <- summarize_species_abundance(SA_14A_Kaiju_viruses,"SA_14A")
abundance_SA_14B_v_genus_per <- summarize_species_abundance(SA_14B_Kaiju_viruses,"SA_14B")
abundance_SA_14C_v_genus_per <- summarize_species_abundance(SA_14C_Kaiju_viruses,"SA_14C")
abundance_SA_21A_v_genus_per <- summarize_species_abundance(SA_21A_Kaiju_viruses,"SA_21A")
abundance_SA_21B_v_genus_per <- summarize_species_abundance(SA_21B_Kaiju_viruses,"SA_21B")
abundance_SA_21C_v_genus_per <- summarize_species_abundance(SA_21C_Kaiju_viruses,"SA_21C")
abundance_SA_28A_v_genus_per <- summarize_species_abundance(SA_28A_Kaiju_viruses,"SA_28A")
abundance_SA_28B_v_genus_per <- summarize_species_abundance(SA_28B_Kaiju_viruses,"SA_28B")
abundance_SA_28C_v_genus_per <- summarize_species_abundance(SA_28C_Kaiju_viruses,"SA_28C")

abundance_kaiju_viruses <- full_join(abundance_SA_0A_v_genus_per,abundance_SA_0B_v_genus_per)%>%
  full_join(.,abundance_SA_0C_v_genus_per)%>%
  full_join(.,abundance_SA_2A_v_genus_per)%>%
  full_join(.,abundance_SA_2B_v_genus_per)%>%
  full_join(.,abundance_SA_2C_v_genus_per)%>%
  full_join(.,abundance_SA_7A_v_genus_per)%>%
  full_join(.,abundance_SA_7B_v_genus_per)%>%
  full_join(.,abundance_SA_7C_v_genus_per)%>%
  full_join(.,abundance_SA_14A_v_genus_per)%>%
  full_join(.,abundance_SA_14B_v_genus_per)%>%
  full_join(.,abundance_SA_14C_v_genus_per)%>%
  full_join(.,abundance_SA_21A_v_genus_per)%>%
  full_join(.,abundance_SA_21B_v_genus_per)%>%
  full_join(.,abundance_SA_21C_v_genus_per)%>%
  full_join(.,abundance_SA_28A_v_genus_per)%>%
  full_join(.,abundance_SA_28B_v_genus_per)%>%
  full_join(.,abundance_SA_28C_v_genus_per)

abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='NA'] <- 'Bacteriophage sp.'
abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='Bacteriophage sp.'] <- 'unknown'
abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='Caudoviricetes sp.'] <- 'unknown Caudoviricetes'
#abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='Escherichia phage KW1E_UTAR'] <- 'unclassified Caudoviricetes'
#abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='Lactobacillus phage Silenus'] <- 'unclassified Coetzeevirus'
#abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='Mitoviridae sp.'] <- 'unknown Mitoviridae'
#abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='Lactobacillus phage LL-K'] <- 'unclassified Caudoviricetes'
#abundance_kaiju_viruses$species[abundance_kaiju_viruses$species=='Xanthomonas phage vB_XooS_NR08'] <- 'unclassified Caudoviricetes'

abundance_kaiju_viruses [is.na(abundance_kaiju_viruses)] <- 0

abundance_kaiju_viruses_2 <- abundance_kaiju_viruses %>%
  group_by(species) %>% 
  summarise(across(everything(), sum, na.rm = TRUE)) %>% 
  column_to_rownames("species")

abundance_kaiju_viruses_2 <- abundance_kaiju_viruses_2%>%
  mutate(sum=rowSums(.)) %>%
  arrange(-sum) %>%
  rownames_to_column(var ="species")

abundance_kaiju_viruses_3 <- mutate(abundance_kaiju_viruses_2,species = ifelse(sum < 10, "Other", species)) %>%
  group_by(species)

abundance_kaiju_viruses_4 <- abundance_kaiju_viruses_3[c("species","SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C")]%>%
  summarise(across(contains("0") | contains("2") | contains("7") | contains("14") | contains("21") | contains("28") , sum, na.rm = TRUE))

abundance_long_viruses <- abundance_kaiju_viruses_4 %>%
  pivot_longer(cols = -species, names_to = "Samples", values_to = "abundance")



print(abundance_kaiju_viruses_4$species)

	
abundance_long_viruses$species <- factor(abundance_long_viruses$species, levels = c("Anamdongvirus LBR48","Duamitovirus hulu1","Escherichia phage KW1E_UTAR","Heilongjiangvirus Lb","Lactobacillus phage Silenus","Mitoviridae sp.","Lactobacillus phage LL-K","Salchichonvirus LP65","Xanthomonas phage vB_XooS_NR08","unknown Caudoviricetes","unknown","Other"))
abundance_long_viruses$Samples <- factor(abundance_long_viruses$Samples, levels = c("SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C"))



abundance_long_viruses_cabb <- left_join(abundance_long_viruses,Fermentation_metadata, by="Samples")



genus_col_vir <- c("Anamdongvirus LBR48" = '#a6cee3' , "Duamitovirus hulu1"="#1f78b4","Escherichia phage KW1E_UTAR"='#fdbf6f',"Heilongjiangvirus Lb"='#ff7f00',"Lactobacillus phage Silenus"='#5e4fa2',"Mitoviridae sp."='#cab2d6',"Lactobacillus phage LL-K"="#fee08b","Salchichonvirus LP65"="#de77ae","Xanthomonas phage vB_XooS_NR08"="#8e0152","unknown"="#b15928", "unknown Caudoviricetes"="lavenderblush3","Other" ='#d9d9d9')

per_read_genus_vir_cabb <- ggplot(abundance_long_viruses_cabb, aes(fill=species, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Species")+
  scale_fill_manual(values = genus_col_vir)+
  theme(legend.text = element_text(size = 11,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 11))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_genus_vir_cabb



#file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_virus_cabbage.tiff")
#ggsave(file_name,plot =per_read_genus_vir_cabb,device = "tiff",width = 15, height = 6, dpi = 300)

```

### Bacterial community at the genus level
Relative abundance for reads for each kingdom were calculated and represented by bar plot. 
```{r sauerkraut bacteria plot, message=FALSE, warning=FALSE}

SA_0A_Kaiju_bacteria <- res_SA_0A$bacteria
SA_0B_Kaiju_bacteria <- res_SA_0B$bacteria
SA_0C_Kaiju_bacteria <- res_SA_0C$bacteria
SA_2A_Kaiju_bacteria <- res_SA_2A$bacteria
SA_2B_Kaiju_bacteria <- res_SA_2B$bacteria
SA_2C_Kaiju_bacteria <- res_SA_2C$bacteria
SA_7A_Kaiju_bacteria <- res_SA_7A$bacteria
SA_7B_Kaiju_bacteria <- res_SA_7B$bacteria
SA_7C_Kaiju_bacteria <- res_SA_7C$bacteria
SA_14A_Kaiju_bacteria <- res_SA_14A$bacteria
SA_14B_Kaiju_bacteria <- res_SA_14B$bacteria
SA_14C_Kaiju_bacteria <- res_SA_14C$bacteria
SA_21A_Kaiju_bacteria <- res_SA_21A$bacteria
SA_21B_Kaiju_bacteria <- res_SA_21B$bacteria
SA_21C_Kaiju_bacteria <- res_SA_21C$bacteria
SA_28A_Kaiju_bacteria <- res_SA_28A$bacteria
SA_28B_Kaiju_bacteria <- res_SA_28B$bacteria
SA_28C_Kaiju_bacteria <- res_SA_28C$bacteria


abundance_SA_0A_b_genus_per <- summarize_genus_abundance(SA_0A_Kaiju_bacteria,"SA_0A")
abundance_SA_0B_b_genus_per <- summarize_genus_abundance(SA_0B_Kaiju_bacteria,"SA_0B")
abundance_SA_0C_b_genus_per <- summarize_genus_abundance(SA_0C_Kaiju_bacteria,"SA_0C")
abundance_SA_2A_b_genus_per <- summarize_genus_abundance(SA_2A_Kaiju_bacteria,"SA_2A")
abundance_SA_2B_b_genus_per <- summarize_genus_abundance(SA_2B_Kaiju_bacteria,"SA_2B")
abundance_SA_2C_b_genus_per <- summarize_genus_abundance(SA_2C_Kaiju_bacteria,"SA_2C")
abundance_SA_7A_b_genus_per <- summarize_genus_abundance(SA_7A_Kaiju_bacteria,"SA_7A")
abundance_SA_7B_b_genus_per <- summarize_genus_abundance(SA_7B_Kaiju_bacteria,"SA_7B")
abundance_SA_7C_b_genus_per <- summarize_genus_abundance(SA_7C_Kaiju_bacteria,"SA_7C")
abundance_SA_14A_b_genus_per <- summarize_genus_abundance(SA_14A_Kaiju_bacteria,"SA_14A")
abundance_SA_14B_b_genus_per <- summarize_genus_abundance(SA_14B_Kaiju_bacteria,"SA_14B")
abundance_SA_14C_b_genus_per <- summarize_genus_abundance(SA_14C_Kaiju_bacteria,"SA_14C")
abundance_SA_21A_b_genus_per <- summarize_genus_abundance(SA_21A_Kaiju_bacteria,"SA_21A")
abundance_SA_21B_b_genus_per <- summarize_genus_abundance(SA_21B_Kaiju_bacteria,"SA_21B")
abundance_SA_21C_b_genus_per <- summarize_genus_abundance(SA_21C_Kaiju_bacteria,"SA_21C")
abundance_SA_28A_b_genus_per <- summarize_genus_abundance(SA_28A_Kaiju_bacteria,"SA_28A")
abundance_SA_28B_b_genus_per <- summarize_genus_abundance(SA_28B_Kaiju_bacteria,"SA_28B")
abundance_SA_28C_b_genus_per <- summarize_genus_abundance(SA_28C_Kaiju_bacteria,"SA_28C")

abundance_kaiju_bacteria_cab <- full_join(abundance_SA_0A_b_genus_per,abundance_SA_0B_b_genus_per)%>%
  full_join(.,abundance_SA_0C_b_genus_per)%>%
  full_join(.,abundance_SA_2A_b_genus_per)%>%
  full_join(.,abundance_SA_2B_b_genus_per)%>%
  full_join(.,abundance_SA_2C_b_genus_per)%>%
  full_join(.,abundance_SA_7A_b_genus_per)%>%
  full_join(.,abundance_SA_7B_b_genus_per)%>%
  full_join(.,abundance_SA_7C_b_genus_per)%>%
  full_join(.,abundance_SA_14A_b_genus_per)%>%
  full_join(.,abundance_SA_14B_b_genus_per)%>%
  full_join(.,abundance_SA_14C_b_genus_per)%>%
  full_join(.,abundance_SA_21A_b_genus_per)%>%
  full_join(.,abundance_SA_21B_b_genus_per)%>%
  full_join(.,abundance_SA_21C_b_genus_per)%>%
  full_join(.,abundance_SA_28A_b_genus_per)%>%
  full_join(.,abundance_SA_28B_b_genus_per)%>%
  full_join(.,abundance_SA_28C_b_genus_per)


abundance_kaiju_bacteria_cab [is.na(abundance_kaiju_bacteria_cab)] <- 0

abundance_kaiju_bacteria_cab_2 <- column_to_rownames(abundance_kaiju_bacteria_cab,"genus")%>%
  mutate(sum=rowSums(.)) %>%
  arrange(-sum) %>%
  rownames_to_column(var ="genus")

abundance_kaiju_bacteria_cab_3 <- mutate(abundance_kaiju_bacteria_cab_2,genus = ifelse(sum < 5, "Other", genus)) %>%
  group_by(genus)

abundance_kaiju_bacteria_cab_4 <- abundance_kaiju_bacteria_cab_3[c("genus","SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C")]

abundance_long_bacteria_cabbage <- abundance_kaiju_bacteria_cab_4 %>%
  pivot_longer(cols = -genus, names_to = "Samples", values_to = "abundance")

#print(abundance_kaiju_bacteria_4$genus)

abundance_long_bacteria_cabbage$genus <- factor(abundance_long_bacteria_cabbage$genus, levels = c("Leuconostoc","Secundilactobacillus","Lactiplantibacillus","Levilactobacillus","Lactobacillus","Pantoea","Pediococcus","Rahnella","Weissella","Enterobacter","Other"))

abundance_long_bacteria_cabbage$Samples <- factor(abundance_long_bacteria_cabbage$Samples, levels = c("SA_0A","SA_0B","SA_0C","SA_2A","SA_2B","SA_2C","SA_7A","SA_7B","SA_7C","SA_14A","SA_14B","SA_14C","SA_21A","SA_21B","SA_21C", "SA_28A", "SA_28B", "SA_28C"))


abundance_long_bacteria_cabbage2 <- left_join(abundance_long_bacteria_cabbage,Fermentation_metadata, by="Samples")



genus_col <- c("Leuconostoc" = '#a6cee3' ,"Secundilactobacillus"= "#1f78b4" ,"Lactiplantibacillus" = '#b2df8a',"Levilactobacillus" = '#fb9a99',"Lactobacillus"= '#e31a1c',"Pantoea"= '#cab2d6',"Pediococcus"=  '#5e4fa2',"Rahnella"='#ff7f00' ,"Weissella" ='#de77ae',"Enterobacter" ='#8e0152',"Other" ='#d9d9d9')

per_read_genus_bact_cabb <- ggplot(abundance_long_bacteria_cabbage2, aes(fill=genus, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Genus")+
  scale_fill_manual(values = genus_col)+
  theme(legend.text = element_text(size = 11,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 11))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_genus_bact_cabb


#file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_bacteria_cabbage.tiff")
#ggsave(file_name,plot =per_read_genus_bact_cabb,device = "tiff",width = 15, height = 6, dpi = 300)
```
#### Statistic analyses of the bacterial community variation during cabbage fermentation
```{r Cabbage Bray-curtis/permanova, message=FALSE, warning=FALSE}
abundance_kaiju_bacteria_cabbage_obj <-abundance_kaiju_bacteria_cab%>%
  column_to_rownames(.,"genus")

Fermentation_metadata_obj <- Fermentation_metadata
Fermentation_metadata_obj$Figure_label <- Fermentation_metadata$Samples
Fermentation_metadata_obj<-column_to_rownames(Fermentation_metadata_obj,"Samples")

obj_Cabbage_metag_bact <- phyloseq(otu_table(abundance_kaiju_bacteria_cabbage_obj, taxa_are_rows = TRUE), sample_data(Fermentation_metadata_obj))
obj_Cabbage_metag_bact

dist_bc_metag_bact_cab <- phyloseq::distance(obj_Cabbage_metag_bact, method = "bray") 
#dist_jcc_metag_bact <- phyloseq::distance(obj_Carrot_metag_bact, method = "jaccard")

# on BC 
set.seed(030623)
adonis_bc_sample_cab <- adonis2(formula = dist_bc_metag_bact_cab~Day_sampling_text, data = data.frame(sample_data(obj_Cabbage_metag_bact)))
adonis_bc_sample_cab
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample_cab$R2[1],3), ", p =", adonis_bc_sample_cab$`Pr(>F)`[1],"***") 



pal_lib <- c("Day 0"="#3D5941FF","Day 2"="#778868FF","Day 7"="#B5B991FF", "Day 14"="#DE8A5AFF","Day 21"="#BE7245FF","Day 28"="#8C311CFF")

# Si vous voulez ordonner sampling_day selon un ordre spécifique
#obj_Carrot_metag_bact@sam_data$sampling_day <- factor(obj_Cabbage_metag_bact@sam_data$sampling_day, 
                                                 #levels = c("0", "2", "7","14","21","28"))  # Spécifiez votre ordre ici

#PCoA with BC
ord_PCoA_bc <- ordinate(obj_Cabbage_metag_bact, method = "PCoA", distance = dist_bc_metag_bact_cab)

p1_metag_bact_cabb_BC <- plot_ordination(obj_Cabbage_metag_bact, ord_PCoA_bc,color = "Day_sampling_text")+
  labs(title = "BRAY-CURTIS",
    subtitle = res.adonis.bc_sample,
    color="Sampling day")+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 10),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 12),  # Taille du texte de la légende
    legend.title = element_text(size = 14))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib)
  
p1_metag_bact_cabb_BC

```

#### Combining the graphs for the final figure
```{r combine cabbage plot, message=FALSE, warning=FALSE}

p_kaiju_cabb <- ggarrange(per_read_kingdom_cabb, per_read_genus_bact_cabb, labels = c("A", "B"),
          ncol = 2, nrow = 1, common.legend = FALSE,legend = "right",
          heights = c(1, 1), widths = c(5, 5),
          align = "hv")+
  theme(plot.background = element_rect(fill = "white", color = NA),
        panel.background = element_rect(fill = "white", color = NA),
           plot.margin = margin(1, 1, 1, 1, "cm")) # Ajuster les marges
p_kaiju_cabb

file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_combine_Kaiju_cabbage.tiff")
ggsave(file_name,plot =p_kaiju_cabb,device = "tiff",width = 20, height = 6, dpi = 300)
```

#### Statistics of results
```{r Cabbage stat, message=FALSE, warning=FALSE}

abundance_kaiju_cabbage <- abundance_kaiju_cabbage %>%
  mutate(Day = case_when(
    grepl("0[A-C]$", Samples) ~ 0,
    grepl("2[A-C]$", Samples) ~ 2,
    grepl("7[A-C]$", Samples) ~ 7,
    grepl("14[A-C]$", Samples) ~ 14,
    grepl("21[A-C]$", Samples) ~ 21,
    grepl("28[A-C]$", Samples) ~ 28,
    TRUE ~ NA_real_  # Valeur par défaut si aucun motif ne correspond
  ))


# Calculate number of reads per day and per superkingdom
reads_summary_cabb <- abundance_kaiju_cabbage %>%
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads)) 


reads_summary_cabb_without_unclassifies <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "Unclassified") %>%  # delete "Unclassified"
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads)) 


## Bacteria stat
reads_summary_cabb_bacteria <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>%  # delete "Unclassified"
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  # Total of classified reads (classifiés (without Unclassified)
    bacteria_reads = sum(reads[superkingdom == "Bacteria"], na.rm = TRUE)  # Reads of bacteria only
    ) %>%
  mutate(proportion_bacteria = bacteria_reads / total_reads_classified)  # Proportion of bacteria on classified reads

# Affichage du résultat
print(reads_summary_cabb_bacteria)


## Eukaryota stat
reads_summary_cabb_euka <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>% 
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE), 
    euka_reads = sum(reads[superkingdom == "Eukaryota"], na.rm = TRUE) 
  ) %>%
  mutate(proportion_euka = euka_reads / total_reads_classified) 

print(reads_summary_cabb_euka)

## Viruses stat
reads_summary_cabb_Viruses <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>% 
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  
    Viruses_reads = sum(reads[superkingdom == "Viruses"], na.rm = TRUE)
    ) %>%
  mutate(proportion_Viruses = Viruses_reads / total_reads_classified)  

print(reads_summary_cabb_Viruses)



## Archaea stat
reads_summary_cabb_Archaea <- abundance_kaiju_cabbage %>%
  filter(superkingdom != "unclassified") %>% 
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE), 
    Archaea_reads = sum(reads[superkingdom == "Archaea"], na.rm = TRUE) 
  ) %>%
  mutate(proportion_Archaea = Archaea_reads / total_reads_classified)  

# Affichage du résultat
print(reads_summary_cabb_Archaea)

```



## Analyses of the carrot fermentation
## Import kaiju results from carrot fermentation
```{r Import data carrot, message=FALSE, warning=FALSE}


CA_0A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_0A_libLAH8297_summary_viruses.tsv",sep="\t",header = TRUE)
CA_0B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_0B_libLAH8298_summary_viruses.tsv",sep="\t",header = TRUE)
CA_0C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_0C_libLAH8299_summary_viruses.tsv",sep="\t",header = TRUE)

CA_2A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_2A_libLAH8300_summary_viruses.tsv", sep="\t",header = TRUE)
CA_2B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_2B_libLAH8301_summary_viruses.tsv", sep="\t",header = TRUE)
CA_2C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_2C_libLAH8302_summary_viruses.tsv", sep="\t",header = TRUE)

CA_7A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_7A_libLAH8303_summary_viruses.tsv", sep="\t",header = TRUE)
CA_7B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_7B_libLAH8304_summary_viruses.tsv", sep="\t",header = TRUE)
CA_7C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_7C_libLAH8305_summary_viruses.tsv", sep="\t",header = TRUE)

CA_14A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_14A_libLAH8306_summary_viruses.tsv", sep="\t",header = TRUE)
CA_14B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_14B_libLAH8307_summary_viruses.tsv", sep="\t",header = TRUE)
CA_14C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_14C_libLAH8308_summary_viruses.tsv", sep="\t",header = TRUE)

CA_21A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_21A_libLAH8309_summary_viruses.tsv", sep="\t",header = TRUE)
CA_21B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_21B_libLAH8310_summary_viruses.tsv", sep="\t",header = TRUE)
CA_21C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_21C_libLAH8311_summary_viruses.tsv", sep="\t",header = TRUE)

CA_28A_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_28A_libLAH8312_summary_viruses.tsv", sep="\t",header = TRUE)
CA_28B_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_28B_libLAH8313_summary_viruses.tsv", sep="\t",header = TRUE)
CA_28C_Kaiju <- read.delim("results/Kaiju_metag/kaiju_NG-A2640_CA_28C_libLAH8314_libLAI5154_summary_viruses.tsv", sep="\t",header = TRUE)

res_CA_0A <- process_kaiju_data(CA_0A_Kaiju, "CA_0A")
res_CA_0B <- process_kaiju_data(CA_0B_Kaiju, "CA_0B")
res_CA_0C <- process_kaiju_data(CA_0C_Kaiju, "CA_0C")
res_CA_2A <- process_kaiju_data(CA_2A_Kaiju, "CA_2A")
res_CA_2B <- process_kaiju_data(CA_2B_Kaiju, "CA_2B")
res_CA_2C <- process_kaiju_data(CA_2C_Kaiju, "CA_2C")
res_CA_7A <- process_kaiju_data(CA_7A_Kaiju, "CA_7A")
res_CA_7B <- process_kaiju_data(CA_7B_Kaiju, "CA_7B")
res_CA_7C <- process_kaiju_data(CA_7C_Kaiju, "CA_7C")
res_CA_14A <- process_kaiju_data(CA_14A_Kaiju, "CA_14A")
res_CA_14B <- process_kaiju_data(CA_14B_Kaiju, "CA_14B")
res_CA_14C <- process_kaiju_data(CA_14C_Kaiju, "CA_14C")
res_CA_21A <- process_kaiju_data(CA_21A_Kaiju, "CA_21A")
res_CA_21B <- process_kaiju_data(CA_21B_Kaiju, "CA_21B")
res_CA_21C <- process_kaiju_data(CA_21C_Kaiju, "CA_21C")
res_CA_28A <- process_kaiju_data(CA_28A_Kaiju, "CA_28A")
res_CA_28B <- process_kaiju_data(CA_28B_Kaiju, "CA_28B")
res_CA_28C <- process_kaiju_data(CA_28C_Kaiju, "CA_28C")

CA_0A_Kaiju_process <- res_CA_0A$processed_data
CA_0B_Kaiju_process <- res_CA_0B$processed_data
CA_0C_Kaiju_process <- res_CA_0C$processed_data
CA_2A_Kaiju_process <- res_CA_2A$processed_data
CA_2B_Kaiju_process <- res_CA_2B$processed_data
CA_2C_Kaiju_process <- res_CA_2C$processed_data
CA_7A_Kaiju_process <- res_CA_7A$processed_data
CA_7B_Kaiju_process <- res_CA_7B$processed_data
CA_7C_Kaiju_process <- res_CA_7C$processed_data
CA_14A_Kaiju_process <- res_CA_14A$processed_data
CA_14B_Kaiju_process <- res_CA_14B$processed_data
CA_14C_Kaiju_process <- res_CA_14C$processed_data
CA_21A_Kaiju_process <- res_CA_21A$processed_data
CA_21B_Kaiju_process <- res_CA_21B$processed_data
CA_21C_Kaiju_process <- res_CA_21C$processed_data
CA_28A_Kaiju_process <- res_CA_28A$processed_data
CA_28B_Kaiju_process <- res_CA_28B$processed_data
CA_28C_Kaiju_process <- res_CA_28C$processed_data

abundance_kaiju_carrot <- full_join (CA_0A_Kaiju_process,CA_0B_Kaiju_process)%>%
  full_join(.,CA_0C_Kaiju_process)%>%
  full_join(.,CA_2A_Kaiju_process)%>%
  full_join(.,CA_2B_Kaiju_process)%>%
  full_join(.,CA_2C_Kaiju_process)%>%
  full_join(.,CA_7A_Kaiju_process)%>%
  full_join(.,CA_7B_Kaiju_process)%>%
  full_join(.,CA_7C_Kaiju_process)%>%
  full_join(.,CA_14A_Kaiju_process)%>%
  full_join(.,CA_14B_Kaiju_process)%>%
  full_join(.,CA_14C_Kaiju_process)%>%
  full_join(.,CA_21A_Kaiju_process)%>%
  full_join(.,CA_21B_Kaiju_process)%>%
  full_join(.,CA_21C_Kaiju_process)%>%
  full_join(.,CA_28A_Kaiju_process)%>%
  full_join(.,CA_28B_Kaiju_process)%>%
  full_join(.,CA_28C_Kaiju_process)

write.table(as.data.frame(abundance_kaiju_carrot), file = "Carrot_read_kaiju.txt", sep = "\t", row.names = T, quote = F, col.names = NA)



options(scipen = 999)
mean_proportion_df <- abundance_kaiju_carrot %>%
  group_by(superkingdom) %>%
  summarise(
    mean_reads = mean(reads, na.rm = TRUE),
    sum_reads = sum(reads),
    .groups = 'drop'
  ) %>%
  mutate(
    proportion = (mean_reads / sum(mean_reads))*100  # Calculer la proportion des moyennes => ligne de code fausse
  )

mean_proportion_df

```
### Microbial community at the domain level
```{r carrot kingdom plot, message=FALSE, warning=FALSE}
abundance_CA_0A_kingdom <- abundance_summary_ki(CA_0A_Kaiju_process, superkingdom, percent, "CA_0A")
abundance_CA_0B_kingdom <- abundance_summary_ki(CA_0B_Kaiju_process, superkingdom, percent, "CA_0B")
abundance_CA_0C_kingdom <- abundance_summary_ki(CA_0C_Kaiju_process, superkingdom, percent, "CA_0C")
abundance_CA_2A_kingdom <- abundance_summary_ki(CA_2A_Kaiju_process, superkingdom, percent, "CA_2A")
abundance_CA_2B_kingdom <- abundance_summary_ki(CA_2B_Kaiju_process, superkingdom, percent, "CA_2B")
abundance_CA_2C_kingdom <- abundance_summary_ki(CA_2C_Kaiju_process, superkingdom, percent, "CA_2C")
abundance_CA_7A_kingdom <- abundance_summary_ki(CA_7A_Kaiju_process, superkingdom, percent, "CA_7A")
abundance_CA_7B_kingdom <- abundance_summary_ki(CA_7B_Kaiju_process, superkingdom, percent, "CA_7B")
abundance_CA_7C_kingdom <- abundance_summary_ki(CA_7C_Kaiju_process, superkingdom, percent, "CA_7C")
abundance_CA_14A_kingdom <- abundance_summary_ki(CA_14A_Kaiju_process, superkingdom, percent, "CA_14A")
abundance_CA_14B_kingdom <- abundance_summary_ki(CA_14B_Kaiju_process, superkingdom, percent, "CA_14B")
abundance_CA_14C_kingdom <- abundance_summary_ki(CA_14C_Kaiju_process, superkingdom, percent, "CA_14C")
abundance_CA_21A_kingdom <- abundance_summary_ki(CA_21A_Kaiju_process, superkingdom, percent, "CA_21A")
abundance_CA_21B_kingdom <- abundance_summary_ki(CA_21B_Kaiju_process, superkingdom, percent, "CA_21B")
abundance_CA_21C_kingdom <- abundance_summary_ki(CA_21C_Kaiju_process, superkingdom, percent, "CA_21C")
abundance_CA_28A_kingdom <- abundance_summary_ki(CA_28A_Kaiju_process, superkingdom, percent, "CA_28A")
abundance_CA_28B_kingdom <- abundance_summary_ki(CA_28B_Kaiju_process, superkingdom, percent, "CA_28B")
abundance_CA_28C_kingdom <- abundance_summary_ki(CA_28C_Kaiju_process, superkingdom, percent, "CA_28C")

abundance_kaiju_kingdom <- full_join(abundance_CA_0A_kingdom,abundance_CA_0B_kingdom)%>%
  full_join(.,abundance_CA_0C_kingdom)%>%
  full_join(.,abundance_CA_2A_kingdom)%>%
  full_join(.,abundance_CA_2B_kingdom)%>%
  full_join(.,abundance_CA_2C_kingdom)%>%
  full_join(.,abundance_CA_7A_kingdom)%>%
  full_join(.,abundance_CA_7B_kingdom)%>%
  full_join(.,abundance_CA_7C_kingdom)%>%
  full_join(.,abundance_CA_14A_kingdom)%>%
  full_join(.,abundance_CA_14B_kingdom)%>%
  full_join(.,abundance_CA_14C_kingdom)%>%
  full_join(.,abundance_CA_21A_kingdom)%>%
  full_join(.,abundance_CA_21B_kingdom)%>%
  full_join(.,abundance_CA_21C_kingdom)%>%
  full_join(.,abundance_CA_28A_kingdom)%>%
  full_join(.,abundance_CA_28B_kingdom)%>%
  full_join(.,abundance_CA_28C_kingdom)

# Convertir en format long pour ggplot2
abundance_long_carrot <- abundance_kaiju_kingdom %>%
  pivot_longer(cols = -superkingdom, names_to = "Samples", values_to = "abundance")

abundance_long_carrot$superkingdom <- factor(abundance_long_carrot$superkingdom , levels = c("Archaea","Bacteria","Eukaryota","Viruses","unclassified"))
kingdom_col <- c("Archaea"='#d7191c',"Bacteria"='#fdae61',"Eukaryota"='#ffffbf',"unclassified"='#abd9e9', "Viruses"='#2c7bb6' )

abundance_long_carrot2 <- left_join(abundance_long_carrot,Fermentation_metadata, by="Samples")


per_read_kingdom_carrot <- ggplot(abundance_long_carrot2, aes(fill=superkingdom, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Kingdom")+
  scale_fill_manual(values = kingdom_col)+
  theme(legend.text = element_text(size = 12),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 12))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
  
per_read_kingdom_carrot

```
### Viral community at the species level
```{r carrot viruses plot, message=FALSE, warning=FALSE}

CA_0A_Kaiju_viruses <- res_CA_0A$viruses
CA_0B_Kaiju_viruses <- res_CA_0B$viruses
CA_0C_Kaiju_viruses <- res_CA_0C$viruses
CA_2A_Kaiju_viruses <- res_CA_2A$viruses
CA_2B_Kaiju_viruses <- res_CA_2B$viruses
CA_2C_Kaiju_viruses <- res_CA_2C$viruses
CA_7A_Kaiju_viruses <- res_CA_7A$viruses
CA_7B_Kaiju_viruses <- res_CA_7B$viruses
CA_7C_Kaiju_viruses <- res_CA_7C$viruses
CA_14A_Kaiju_viruses <- res_CA_14A$viruses
CA_14B_Kaiju_viruses <- res_CA_14B$viruses
CA_14C_Kaiju_viruses <- res_CA_14C$viruses
CA_21A_Kaiju_viruses <- res_CA_21A$viruses
CA_21B_Kaiju_viruses <- res_CA_21B$viruses
CA_21C_Kaiju_viruses <- res_CA_21C$viruses
CA_28A_Kaiju_viruses <- res_CA_28A$viruses
CA_28B_Kaiju_viruses <- res_CA_28B$viruses
CA_28C_Kaiju_viruses <- res_CA_28C$viruses


abundance_CA_0A_v_genus_per <- summarize_species_abundance(CA_0A_Kaiju_viruses,"CA_0A")
abundance_CA_0B_v_genus_per <- summarize_species_abundance(CA_0B_Kaiju_viruses,"CA_0B")
abundance_CA_0C_v_genus_per <- summarize_species_abundance(CA_0C_Kaiju_viruses,"CA_0C")
abundance_CA_2A_v_genus_per <- summarize_species_abundance(CA_2A_Kaiju_viruses,"CA_2A")
abundance_CA_2B_v_genus_per <- summarize_species_abundance(CA_2B_Kaiju_viruses,"CA_2B")
abundance_CA_2C_v_genus_per <- summarize_species_abundance(CA_2C_Kaiju_viruses,"CA_2C")
abundance_CA_7A_v_genus_per <- summarize_species_abundance(CA_7A_Kaiju_viruses,"CA_7A")
abundance_CA_7B_v_genus_per <- summarize_species_abundance(CA_7B_Kaiju_viruses,"CA_7B")
abundance_CA_7C_v_genus_per <- summarize_species_abundance(CA_7C_Kaiju_viruses,"CA_7C")
abundance_CA_14A_v_genus_per <- summarize_species_abundance(CA_14A_Kaiju_viruses,"CA_14A")
abundance_CA_14B_v_genus_per <- summarize_species_abundance(CA_14B_Kaiju_viruses,"CA_14B")
abundance_CA_14C_v_genus_per <- summarize_species_abundance(CA_14C_Kaiju_viruses,"CA_14C")
abundance_CA_21A_v_genus_per <- summarize_species_abundance(CA_21A_Kaiju_viruses,"CA_21A")
abundance_CA_21B_v_genus_per <- summarize_species_abundance(CA_21B_Kaiju_viruses,"CA_21B")
abundance_CA_21C_v_genus_per <- summarize_species_abundance(CA_21C_Kaiju_viruses,"CA_21C")
abundance_CA_28A_v_genus_per <- summarize_species_abundance(CA_28A_Kaiju_viruses,"CA_28A")
abundance_CA_28B_v_genus_per <- summarize_species_abundance(CA_28B_Kaiju_viruses,"CA_28B")
abundance_CA_28C_v_genus_per <- summarize_species_abundance(CA_28C_Kaiju_viruses,"CA_28C")

abundance_kaiju_viruses_car <- full_join(abundance_CA_0A_v_genus_per,abundance_CA_0B_v_genus_per)%>%
  full_join(.,abundance_CA_0C_v_genus_per)%>%
  full_join(.,abundance_CA_2A_v_genus_per)%>%
  full_join(.,abundance_CA_2B_v_genus_per)%>%
  full_join(.,abundance_CA_2C_v_genus_per)%>%
  full_join(.,abundance_CA_7A_v_genus_per)%>%
  full_join(.,abundance_CA_7B_v_genus_per)%>%
  full_join(.,abundance_CA_7C_v_genus_per)%>%
  full_join(.,abundance_CA_14A_v_genus_per)%>%
  full_join(.,abundance_CA_14B_v_genus_per)%>%
  full_join(.,abundance_CA_14C_v_genus_per)%>%
  full_join(.,abundance_CA_21A_v_genus_per)%>%
  full_join(.,abundance_CA_21B_v_genus_per)%>%
  full_join(.,abundance_CA_21C_v_genus_per)%>%
  full_join(.,abundance_CA_28A_v_genus_per)%>%
  full_join(.,abundance_CA_28B_v_genus_per)%>%
  full_join(.,abundance_CA_28C_v_genus_per)

abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='NA'] <- 'Bacteriophage sp.'
abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='Bacteriophage sp.'] <- 'unknown'
abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='Caudoviricetes sp.'] <- 'unknown Caudoviricetes'
#abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='Escherichia phage KW1E_UTAR'] <- 'unclassified Caudoviricetes'
#abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='Lactobacillus phage Silenus'] <- 'unclassified Coetzeevirus'
#abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='Mitoviridae sp.'] <- 'unknown Mitoviridae'
#abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='Lactobacillus phage LL-K'] <- 'unclassified Caudoviricetes'
#abundance_kaiju_viruses_car$species[abundance_kaiju_viruses_car$species=='Xanthomonas phage vB_XooS_NR08'] <- 'unclassified Caudoviricetes'

abundance_kaiju_viruses_car [is.na(abundance_kaiju_viruses_car)] <- 0

abundance_kaiju_viruses_car2 <- abundance_kaiju_viruses_car %>%
  group_by(species) %>% 
  summarise(across(everything(), sum, na.rm = TRUE)) %>% 
  column_to_rownames("species")

abundance_kaiju_viruses_car2 <- abundance_kaiju_viruses_car2%>%
  mutate(sum=rowSums(.)) %>%
  arrange(-sum) %>%
  rownames_to_column(var ="species")

abundance_kaiju_viruses_car3 <- mutate(abundance_kaiju_viruses_car2,species = ifelse(sum < 10, "Other", species)) %>%
  group_by(species)

abundance_kaiju_viruses_car4 <- abundance_kaiju_viruses_car3[c("species","CA_0A","CA_0B","CA_0C","CA_2A","CA_2B","CA_2C","CA_7A","CA_7B","CA_7C","CA_14A","CA_14B","CA_14C","CA_21A","CA_21B","CA_21C", "CA_28A", "CA_28B", "CA_28C")]%>%
  summarise(across(contains("0") | contains("2") | contains("7") | contains("14") | contains("21") | contains("28") , sum, na.rm = TRUE))

abundance_long_viruses_car <- abundance_kaiju_viruses_car4 %>%
  pivot_longer(cols = -species, names_to = "Samples", values_to = "abundance")



print(abundance_kaiju_viruses_car4$species)

	
abundance_long_viruses_car$species <- factor(abundance_long_viruses_car$species, levels = c("Leuconostoc phage P974","Leuconostoc phage phiLNTR2","Unaquatrovirus uv1A4","Unaquatrovirus LN25","Unaquatrovirus LN34","Unaquatrovirus Ln9","unknown Caudoviricetes","unknown","Other"))
abundance_long_viruses_car$Samples <- factor(abundance_long_viruses_car$Samples, levels = c("CA_0A","CA_0B","CA_0C","CA_2A","CA_2B","CA_2C","CA_7A","CA_7B","CA_7C","CA_14A","CA_14B","CA_14C","CA_21A","CA_21B","CA_21C", "CA_28A", "CA_28B", "CA_28C"))


abundance_long_viruses_carr <- left_join(abundance_long_viruses_car,Fermentation_metadata, by="Samples")



#b2df8a
#33a02c
#fb9a99
#e31a1c
#fdbf6f
#ff7f00
#cab2d6

genus_col_vir <- c("Leuconostoc phage P974" = '#a6cee3' , "Leuconostoc phage phiLNTR2"="#1f78b4","Unaquatrovirus uv1A4"='#b2df8a',"Unaquatrovirus LN25"='#33a02c',"Unaquatrovirus LN34"='#fb9a99',"Unaquatrovirus Ln9"='#e31a1c',"unknown"="#fdbf6f", "unknown Caudoviricetes"="#ff7f00","Other" ='#cab2d6')

per_read_genus_vir_carr <- ggplot(abundance_long_viruses_carr, aes(fill=species, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Species")+
  scale_fill_manual(values = genus_col_vir)+
  theme(legend.text = element_text(size = 12,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 12))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_genus_vir_carr



#file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_virus_cabbage.tiff")
#ggsave(file_name,plot =per_read_genus_vir_cabb,device = "tiff",width = 15, height = 6, dpi = 300)

```
### Bacterial community at the genus level
```{r carrot bacteria plot, message=FALSE, warning=FALSE}

CA_0A_Kaiju_bacteria <- res_CA_0A$bacteria
CA_0B_Kaiju_bacteria <- res_CA_0B$bacteria
CA_0C_Kaiju_bacteria <- res_CA_0C$bacteria
CA_2A_Kaiju_bacteria <- res_CA_2A$bacteria
CA_2B_Kaiju_bacteria <- res_CA_2B$bacteria
CA_2C_Kaiju_bacteria <- res_CA_2C$bacteria
CA_7A_Kaiju_bacteria <- res_CA_7A$bacteria
CA_7B_Kaiju_bacteria <- res_CA_7B$bacteria
CA_7C_Kaiju_bacteria <- res_CA_7C$bacteria
CA_14A_Kaiju_bacteria <- res_CA_14A$bacteria
CA_14B_Kaiju_bacteria <- res_CA_14B$bacteria
CA_14C_Kaiju_bacteria <- res_CA_14C$bacteria
CA_21A_Kaiju_bacteria <- res_CA_21A$bacteria
CA_21B_Kaiju_bacteria <- res_CA_21B$bacteria
CA_21C_Kaiju_bacteria <- res_CA_21C$bacteria
CA_28A_Kaiju_bacteria <- res_CA_28A$bacteria
CA_28B_Kaiju_bacteria <- res_CA_28B$bacteria
CA_28C_Kaiju_bacteria <- res_CA_28C$bacteria


abundance_CA_0A_b_genus_per <- summarize_genus_abundance(CA_0A_Kaiju_bacteria,"CA_0A")
abundance_CA_0B_b_genus_per <- summarize_genus_abundance(CA_0B_Kaiju_bacteria,"CA_0B")
abundance_CA_0C_b_genus_per <- summarize_genus_abundance(CA_0C_Kaiju_bacteria,"CA_0C")
abundance_CA_2A_b_genus_per <- summarize_genus_abundance(CA_2A_Kaiju_bacteria,"CA_2A")
abundance_CA_2B_b_genus_per <- summarize_genus_abundance(CA_2B_Kaiju_bacteria,"CA_2B")
abundance_CA_2C_b_genus_per <- summarize_genus_abundance(CA_2C_Kaiju_bacteria,"CA_2C")
abundance_CA_7A_b_genus_per <- summarize_genus_abundance(CA_7A_Kaiju_bacteria,"CA_7A")
abundance_CA_7B_b_genus_per <- summarize_genus_abundance(CA_7B_Kaiju_bacteria,"CA_7B")
abundance_CA_7C_b_genus_per <- summarize_genus_abundance(CA_7C_Kaiju_bacteria,"CA_7C")
abundance_CA_14A_b_genus_per <- summarize_genus_abundance(CA_14A_Kaiju_bacteria,"CA_14A")
abundance_CA_14B_b_genus_per <- summarize_genus_abundance(CA_14B_Kaiju_bacteria,"CA_14B")
abundance_CA_14C_b_genus_per <- summarize_genus_abundance(CA_14C_Kaiju_bacteria,"CA_14C")
abundance_CA_21A_b_genus_per <- summarize_genus_abundance(CA_21A_Kaiju_bacteria,"CA_21A")
abundance_CA_21B_b_genus_per <- summarize_genus_abundance(CA_21B_Kaiju_bacteria,"CA_21B")
abundance_CA_21C_b_genus_per <- summarize_genus_abundance(CA_21C_Kaiju_bacteria,"CA_21C")
abundance_CA_28A_b_genus_per <- summarize_genus_abundance(CA_28A_Kaiju_bacteria,"CA_28A")
abundance_CA_28B_b_genus_per <- summarize_genus_abundance(CA_28B_Kaiju_bacteria,"CA_28B")
abundance_CA_28C_b_genus_per <- summarize_genus_abundance(CA_28C_Kaiju_bacteria,"CA_28C")

abundance_kaiju_bacteria_carrot <- full_join(abundance_CA_0A_b_genus_per,abundance_CA_0B_b_genus_per)%>%
  full_join(.,abundance_CA_0C_b_genus_per)%>%
  full_join(.,abundance_CA_2A_b_genus_per)%>%
  full_join(.,abundance_CA_2B_b_genus_per)%>%
  full_join(.,abundance_CA_2C_b_genus_per)%>%
  full_join(.,abundance_CA_7A_b_genus_per)%>%
  full_join(.,abundance_CA_7B_b_genus_per)%>%
  full_join(.,abundance_CA_7C_b_genus_per)%>%
  full_join(.,abundance_CA_14A_b_genus_per)%>%
  full_join(.,abundance_CA_14B_b_genus_per)%>%
  full_join(.,abundance_CA_14C_b_genus_per)%>%
  full_join(.,abundance_CA_21A_b_genus_per)%>%
  full_join(.,abundance_CA_21B_b_genus_per)%>%
  full_join(.,abundance_CA_21C_b_genus_per)%>%
  full_join(.,abundance_CA_28A_b_genus_per)%>%
  full_join(.,abundance_CA_28B_b_genus_per)%>%
  full_join(.,abundance_CA_28C_b_genus_per)


abundance_kaiju_bacteria_carrot [is.na(abundance_kaiju_bacteria_carrot)] <- 0

abundance_kaiju_bacteria_carrot2 <- column_to_rownames(abundance_kaiju_bacteria_carrot,"genus")%>%
  mutate(sum=rowSums(.)) %>%
  arrange(-sum) %>%
  rownames_to_column(var ="genus")

abundance_kaiju_bacteria_carrot3 <- mutate(abundance_kaiju_bacteria_carrot2,genus = ifelse(sum < 5, "Other", genus)) %>%
  group_by(genus)

abundance_kaiju_bacteria_carrot4 <- abundance_kaiju_bacteria_carrot3[c("genus","CA_0A","CA_0B","CA_0C","CA_2A","CA_2B","CA_2C","CA_7A","CA_7B","CA_7C","CA_14A","CA_14B","CA_14C","CA_21A","CA_21B","CA_21C", "CA_28A", "CA_28B", "CA_28C")]

abundance_long_bacteria_cabbage2 <- left_join(abundance_long_bacteria_cabbage,Fermentation_metadata, by="Samples")


abundance_long_bacteria_carrot <- abundance_kaiju_bacteria_carrot4 %>%
  pivot_longer(cols = -genus, names_to = "Samples", values_to = "abundance")

print(abundance_kaiju_bacteria_carrot2$genus)

abundance_long_bacteria_carrot$genus <- factor(abundance_long_bacteria_carrot$genus, levels = c("Leuconostoc","Lactiplantibacillus","Levilactobacillus","Weissella","Other"))

abundance_long_bacteria_carrot$Samples <- factor(abundance_long_bacteria_carrot$Samples, levels = c("CA_0A","CA_0B","CA_0C","CA_2A","CA_2B","CA_2C","CA_7A","CA_7B","CA_7C","CA_14A","CA_14B","CA_14C","CA_21A","CA_21B","CA_21C", "CA_28A", "CA_28B", "CA_28C"))

genus_col <- c("Leuconostoc" = '#a6cee3',"Lactiplantibacillus" = '#b2df8a',"Levilactobacillus" = '#fb9a99',"Weissella" ='#de77ae',"Other" ='#d9d9d9')

abundance_long_bacteria_carrot2 <- left_join(abundance_long_bacteria_carrot,Fermentation_metadata, by="Samples")

per_read_genus_bact_carr <- ggplot(abundance_long_bacteria_carrot2, aes(fill=genus, y=abundance, x=Samples)) + 
  geom_bar(stat="identity", position="stack") +
  labs(x="",y="Relative abundance (%)"
  )+
  labs(fill = "Genus")+
  scale_fill_manual(values = genus_col)+
  theme(legend.text = element_text(size = 12,face = 'italic'),  # Taille du texte de la légende
    legend.title = element_text(size = 14),# Taille du titre de la légende
    axis.text.x = element_blank(),
    panel.background = element_rect(fill = "white"),
    axis.ticks = element_blank(),
    strip.text = element_text(size = 12))+
  facet_wrap(~ Day_sampling_text, scales = "free_x", nrow = 1 )
per_read_genus_bact_carr



file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_bacteria_carrot.tiff")
ggsave(file_name,plot =per_read_genus_bact_carr,device = "tiff",width = 15, height = 6, dpi = 300)
```
#### Statistic analyses of the bacterial community variation during carrot fermentation
```{r Carrot Bray-curtis/permanova, message=FALSE, warning=FALSE}
abundance_kaiju_bacteria_carrot_obj <-abundance_kaiju_bacteria_carrot%>%
  column_to_rownames(.,"genus")

Fermentation_metadata_obj <- Fermentation_metadata
Fermentation_metadata_obj$Figure_label <- Fermentation_metadata$Samples
Fermentation_metadata_obj<-column_to_rownames(Fermentation_metadata_obj,"Samples")

obj_Carrot_metag_bact <- phyloseq(otu_table(abundance_kaiju_bacteria_carrot_obj, taxa_are_rows = TRUE), sample_data(Fermentation_metadata_obj))
obj_Carrot_metag_bact

dist_bc_metag_bact <- phyloseq::distance(obj_Carrot_metag_bact, method = "bray") 
#dist_jcc_metag_bact <- phyloseq::distance(obj_Carrot_metag_bact, method = "jaccard")

# on BC 
set.seed(030623)
adonis_bc_sample <- adonis2(formula = dist_bc_metag_bact~Day_sampling_text, data = data.frame(sample_data(obj_Carrot_metag_bact)))
adonis_bc_sample
res.adonis.bc_sample <- paste("Permanova, ", "R2", " = ", round(adonis_bc_sample$R2[1],3), ", p =", adonis_bc_sample$`Pr(>F)`[1],"**") 



pal_lib <- c("Day 0"="#3D5941FF","Day 2"="#778868FF","Day 7"="#B5B991FF", "Day 14"="#DE8A5AFF","Day 21"="#BE7245FF","Day 28"="#8C311CFF")

# Si vous voulez ordonner sampling_day selon un ordre spécifique
#obj_Carrot_metag_bact@sam_data$sampling_day <- factor(obj_Carrot_metag_bact@sam_data$sampling_day, 
                                                 #levels = c("0", "2", "7","14","21","28"))  # Spécifiez votre ordre ici

#PCoA with BC
ord_PCoA_bc <- ordinate(obj_Carrot_metag_bact, method = "PCoA", distance = dist_bc_metag_bact)

p1_metag_bact_BC <- plot_ordination(obj_Carrot_metag_bact, ord_PCoA_bc,color = "Day_sampling_text")+
  labs(title = "BRAY-CURTIS",
    subtitle = res.adonis.bc_sample,
    color="Sampling day")+
  theme(plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 10),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "black"),
        legend.text = element_text(size = 12),  # Taille du texte de la légende
    legend.title = element_text(size = 14))+
  geom_point(size=3)+
  geom_text_repel(aes(label=Figure_label),size=4,vjust = 1.5)+
  scale_colour_manual(values = pal_lib)
  
p1_metag_bact_BC


file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_Kaiju_carrot_PCoA_BC_.tiff")
ggsave(file_name,plot =p1_metag_bact_BC,device = "tiff",width = 10, height = 6, dpi = 300)

```

#### Combining the graphs for the final figure
```{r combine carrot plot, message=FALSE, warning=FALSE}

p_kaiju_carrot <- ggarrange(per_read_kingdom_carrot, per_read_genus_bact_carr, labels = c("A", "B"),
          ncol = 2, nrow = 1, common.legend = FALSE,legend = "right",
          heights = c(1, 1), widths = c(5, 5),
          align = "hv")+
  theme(plot.background = element_rect(fill = "white", color = NA),
        panel.background = element_rect(fill = "white", color = NA),
           plot.margin = margin(1, 1, 1, 1, "cm")) # Ajuster les marges
p_kaiju_carrot

file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_combine_Kaiju_carrot.tiff")
ggsave(file_name,plot =p_kaiju_carrot,device = "tiff",width = 20, height = 6, dpi = 300)
```

```{r Carrot figure, message=FALSE, warning=FALSE}
p_kaiju_carrot <- ggarrange(per_read_kingdom_carrot, per_read_genus_vir_carr, labels = c("A", "B"),
          ncol = 2, nrow = 1, common.legend = FALSE,legend = "right",
          heights = c(1, 1), widths = c(5, 5),
          align = "hv")
p_kaiju_carrot

p_compo_microb <- ggarrange(per_read_genus_bact_carr,p1_metag_bact_BC, labels = c("C","D"),
          ncol = 2, nrow = 1, common.legend = FALSE,legend = "right",
          heights = c(1, 1), widths = c(5, 5),
          align = "hv")
p_compo_microb

p_compo_microb_VF <- ggarrange(p_kaiju_carrot,p_compo_microb,
          ncol = 1, nrow = 2, common.legend = FALSE,legend = "right",
          heights = c(1, 1), widths = c(5, 5),
          align = "hv")+
  theme(plot.background = element_rect(fill = "white", color = NA),
        panel.background = element_rect(fill = "white", color = NA),
           plot.margin = margin(1, 1, 1, 1, "cm")) # Ajuster les marges
p_compo_microb_VF

file_name <- paste0("/work_projet/metasimfood/metavirome/2024_spontaneous_fermentation/results/graph/",Sys.Date(),"_combine_microb_carrot.tiff")
ggsave(file_name,plot =p_compo_microb_VF,device = "tiff",width = 18, height = 10, dpi = 300)

```

#### Statistics of results
```{r Carrot stat, message=FALSE, warning=FALSE}
abundance_kaiju_carrot
abundance_kaiju_carrot <- abundance_kaiju_carrot %>%
  mutate(Day = case_when(
    grepl("0[A-C]$", Samples) ~ 0,
    grepl("2[A-C]$", Samples) ~ 2,
    grepl("7[A-C]$", Samples) ~ 7,
    grepl("14[A-C]$", Samples) ~ 14,
    grepl("21[A-C]$", Samples) ~ 21,
    grepl("28[A-C]$", Samples) ~ 28,
    TRUE ~ NA_real_  # Valeur par défaut si aucun motif ne correspond
  ))


reads_summary_carr <- abundance_kaiju_carrot %>%
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads)) 


reads_summary_carr_without_unclassifies <- abundance_kaiju_carrot %>%
  filter(superkingdom != "Unclassified") %>%  
  group_by(Day, superkingdom) %>%
  summarise(total_reads = sum(reads, na.rm = TRUE), .groups = "drop") %>%
  mutate(proportion = total_reads / sum(total_reads)) 


## Bacteria stat
reads_summary_carr_bacteria <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>%  
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  
    bacteria_reads = sum(reads[superkingdom == "Bacteria"], na.rm = TRUE) 
  ) %>%
  mutate(proportion_bacteria = bacteria_reads / total_reads_classified)  


print(reads_summary_carr_bacteria)


## Eukaryota stat
reads_summary_carr_euka <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>%  
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE), 
    euka_reads = sum(reads[superkingdom == "Eukaryota"], na.rm = TRUE) 
  ) %>%
  mutate(proportion_euka = euka_reads / total_reads_classified) 


print(reads_summary_carr_euka)

## Viruses stat
reads_summary_carr_Viruses <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>%  
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE), 
    Viruses_reads = sum(reads[superkingdom == "Viruses"], na.rm = TRUE) 
  ) %>%
  mutate(proportion_Viruses = Viruses_reads / total_reads_classified)  


print(reads_summary_carr_Viruses)



## Archaea stat
reads_summary_carr_Archaea <- abundance_kaiju_carrot %>%
  filter(superkingdom != "unclassified") %>% 
  group_by(Day) %>%
  summarise(
    total_reads_classified = sum(reads, na.rm = TRUE),  
    Archaea_reads = sum(reads[superkingdom == "Archaea"], na.rm = TRUE)
  ) %>%
  mutate(proportion_Archaea = Archaea_reads / total_reads_classified)
  

# Affichage du résultat
print(reads_summary_carr_Archaea)

```


# Analyses of the MAGs abundance during cabbage and carrot fermentation.

## Data import Carrot

```{r Import data, message=FALSE, warning=FALSE}
#Import count matrix of carrot samples
count_metag_carrot <- read.table("results/SnakeMAGs/MAPPING_BWA-MEM2/Carrot/presence_counts.csv") 
count_metag_carrot <- count_metag_carrot[-1:-2,] #delete 2 firsts line that don't corresponding to contigs.

samples_sum_reads<-colSums(count_metag_carrot)

count_metag_carrot<- count_metag_carrot%>%
  rownames_to_column(.,var="Contigs")

list_reference_MAG_carrot <- c("CA_14A_1","CA_14A_5","CA_2A_4","CA_2C_2","CA_7C_7")

info_contigs_carrot <- read.table("METADATA/Contig_MAG_info_carrot.txt",sep="\t",header = TRUE) 
info_MAGs_carrot0 <- read.table("METADATA/MAG_info_carrot.txt",sep="\t",header = TRUE)
info_MAGs_carrot0 <- info_MAGs_carrot0 %>% filter(MAG_ID %in% list_reference_MAG_carrot)
info_MAGs_carrot <- info_MAGs_carrot0%>%
  column_to_rownames(.,var="MAG_ID")



count_metag_carrot_info <- right_join(count_metag_carrot,info_contigs_carrot,by="Contigs")

# Supposons que df_abundance contient les abondances des contigs par échantillon
df_MAG_abundance_carr <- count_metag_carrot_info %>%
  group_by(MAGs) %>% 
  summarise(CA_0A = sum(CA_0A, na.rm = TRUE),
            CA_0B = sum(CA_0B, na.rm = TRUE),
            CA_0C = sum(CA_0C, na.rm = TRUE),
            CA_2A = sum(CA_2A, na.rm = TRUE),
            CA_2B = sum(CA_2B, na.rm = TRUE),
            CA_2C = sum(CA_2C, na.rm = TRUE),
            CA_7A = sum(CA_7A, na.rm = TRUE),
            CA_7B = sum(CA_7B, na.rm = TRUE),
            CA_7C = sum(CA_7C, na.rm = TRUE),
            CA_14A = sum(CA_14A, na.rm = TRUE),
            CA_14B = sum(CA_14B, na.rm = TRUE),
            CA_14C = sum(CA_14C, na.rm = TRUE),
            CA_21A = sum(CA_21A, na.rm = TRUE),
            CA_21B = sum(CA_21B, na.rm = TRUE),
            CA_21C = sum(CA_21C, na.rm = TRUE),
            CA_28A = sum(CA_28A, na.rm = TRUE),
            CA_28B = sum(CA_28B, na.rm = TRUE),
            CA_28C = sum(CA_28C, na.rm = TRUE))%>%
  column_to_rownames(.,var="MAGs")
 

```

### Abundance table for Carrot
We created a Phyloseq object that contains the abundance table and metadata of all samples.
```{r all abundance table, message=FALSE, warning=FALSE}
# Check consistency 
identical(rownames(df_MAG_abundance_carr), rownames(info_MAGs_carrot))  # Doit retourner TRUE

df_MAG_abundance_carr <- df_MAG_abundance_carr[order(rownames(df_MAG_abundance_carr)), , drop = FALSE]
info_MAGs_carrot <- info_MAGs_carrot[order(rownames(info_MAGs_carrot)), , drop = FALSE]

identical(rownames(df_MAG_abundance_carr), rownames(info_MAGs_carrot))  # Devrait être TRUE maintenant

# Vérifier si les noms des échantillons correspondent
identical(colnames(df_MAG_abundance_carr), rownames(sample_metag_metadata))  # Doit


# Check consistency
all(colnames(df_MAG_abundance_carr) %in% rownames(sample_metag_metadata)) ## sample names between metadata and otu table
all(rownames(df_MAG_abundance_carr) %in% rownames(info_MAGs_carrot)) ## taxa names between taxonomic and otu tables

otu_tab <- otu_table(as.matrix(df_MAG_abundance_carr), taxa_are_rows = TRUE)
tax_tab <- tax_table(as.matrix(info_MAGs_carrot))

obj_metag_carrot <- phyloseq(otu_tab, sample_data(sample_metag_metadata), tax_tab)

otu_table(obj_metag_carrot)
tax_table(obj_metag_carrot)

```
#### Normalisation of abundance table for carrot
We transformed abundance data by TPM. We normalized by contig length and
transformed to proportions
```{r TPM transformation, message=FALSE, warning=FALSE}
# Extract abundance table 
abundance_table_carrot <- as.data.frame(otu_table(obj_metag_carrot)) %>% rownames_to_column(var = "MAG_ID")
#recovery the abundance table of obj_clean in the form of dataframe. The row name is the contigs identify.

# Add contig length and set the vector of contig length
abundance_table_carrot2 <- left_join(abundance_table_carrot, info_MAGs_carrot0, by = "MAG_ID")
vect_length_carrot <- abundance_table_carrot2$Genome.size..bp. 
# join the abundance table and the list of contig length into the data frame abundance_table2 based on the column contig identity.
# creat the vector vect_length that contain all length of the abundance table.
  
# Transform counts to TPM
## First divide the count matrix by the vector of contig length
obj_metag_carrot_TPM <- obj_metag_carrot
otu_table(obj_metag_carrot_TPM) <- otu_table(obj_metag_carrot_TPM) / vect_length_carrot 

## Second transform the count matrix into proportions 
count_to_prop_sub <- function(x) {return( x / sum(x) )}
obj_metag_carrot_TPM <- transform_sample_counts(obj_metag_carrot_TPM, count_to_prop_sub)
#sample_sums(obj_All_TPM)
# creat function count_to_prop that calculate the proportion based of the count of abundance divide by the sum of the abundance.
# transform_sample_counts: Transform abundance data in an otu_table, sample-by-sample.
# sample_sums: Returns the total number of individuals observed from each sample.


#sample_sums(obj_DNA)
```


```{r Pheatmap normalized filtred, message=FALSE, warning=FALSE}

#transform log function
log_fun <- function(x){
  (log((x + 0.000001),10))
}

MAGs_carrot_abundance <-  data.frame(otu_table(obj_metag_carrot_TPM))
MAGs_carrot_abundance_norm <- t(apply(MAGs_carrot_abundance, 1, log_fun))
sample_annotation_metag_carr <- data.frame(sample_data(obj_metag_carrot_TPM))

#Extract annotation data 
MAG_annotations <- data.frame(tax_table(obj_metag_carrot_TPM))
table(MAG_annotations$Species)

# Extract vOTU metadata
Classification_carr <- MAG_annotations["Species"]

# Creat dataframe `row_ann` for the row annotation
row_ann_carr <- data.frame(Classification)
col_ann_carr <- sample_annotation_metag_carr["sampling_day"]
col_ann_carr$sampling_day <- as.character(col_ann_carr$sampling_day)
col_ann_carr$sampling_day <- factor (col_ann_carr$sampling_day, levels = c("0","2","7","14","21","28"))

custom_colnames <- 
  if ("Figure_label" %in% colnames(sample_annotation_metag_carr)) {
  custom_colnames_MAGS_cabb <- sample_annotation_metag_carr$Figure_label
} else {
  stop(" 'NewNames' column doesn't exist in samples data")
}

names(col_ann_carr)[names(col_ann_carr) == "sampling_day"] <- "Sampling day"

ann_color_carr = list("Sampling day"=c("0"="#1C3333FF" , 
                            "2"="#226060FF", 
                            "7"="#639CA4FF",
                            "14"="#D2AD7CFF",
                            "21"="#BE7245FF",
                            "28"="#8C311CFF"),
                 "Species"=c('Lactiplantibacillus plantarum' = '#cab2d6',
                                'Leuconostoc inhae' = '#1f78b4',
                                'Levilactobacillus brevis' = '#ffff99',
                                "Leuconostoc mesenteroides"= "#a6cee3"))


### Representation of MAGS after reduced-center normalization with respect to the MAGS.
t.MAGs_carrot_abundance.scale <- scale(t(MAGs_carrot_abundance)) # center-reduced normalization based on the transposed frequency table so that it applies to the MAG rather than per sample
MAGs_carrot_abundance.scale <- t(t.MAGs_carrot_abundance.scale) #Transposition of the normalized table to create the heat map

# Define the breaks for the color scale
breaks <- seq(-2.7, 2.7, length.out = 101)
color_palette <- colorRampPalette(c("#2c7bb6","white","#d7191c"))(length(breaks))

# Creat the heatmap
ph_MAGs_carr_Z <- pheatmap(MAGs_carrot_abundance.scale,
               show_rownames = TRUE,
               cluster_cols = FALSE,
               annotation_row = row_ann_carr,
               annotation_col = col_ann_carr,
               annotation_colors = ann_color_carr,
               fontsize_col = 7,
               fontsize = 6,
               angle_col = 45,
               labels_col = custom_colnames,
               legend_breaks = c(-2.5, 0, 2, max(MAGs_carrot_abundance.scale)),
               legend_labels = c("-2.5", "0", "2", "Z-score"),
               color = color_palette,
               breaks = breaks) 
ph_MAGs_carr_Z


## A graph showing the relative abundance of MAGs during fermentation.
# Extract and transpose the abundance table (species in columns)
otu_table_df <- as.data.frame(t(otu_table(obj_metag_carrot_TPM)))
otu_table_df$SampleID <- rownames(otu_table_df)

# Metadata extraction
metadata_df <- as.data.frame(sample_data(obj_metag_carrot_TPM))
metadata_df$SampleID <- rownames(metadata_df)

# Extraction of the taxonomy 
tax_df <- as.data.frame(tax_table(obj_metag_carrot_TPM))
tax_df$MAG_ID <- rownames(tax_df)
colnames(tax_df) <- trimws(colnames(tax_df))  # Remove extra spaces
names(tax_df)

# keep only species column
tax_df <- tax_df %>%
  select(MAG_ID, Species)

otu_long <- otu_table_df %>%
  pivot_longer(cols = -SampleID, names_to = "MAG_ID", values_to = "Abundance") %>%
  left_join(metadata_df, by = "SampleID") %>%
  left_join(tax_df, by = "MAG_ID")  # Ajouter les noms des espèces

# Calculate relative abundance
otu_long <- otu_long %>%
  group_by(SampleID) %>%
  mutate(Rel_Abundance = (Abundance / sum(Abundance, na.rm = TRUE))*100)

# Calculate the mean and standard deviation for each species on each day
otu_summary <- otu_long %>%
  group_by(sampling_day, Species,MAG_ID) %>%
  summarise(mean_abund = mean(Rel_Abundance, na.rm = TRUE),
            sd_abund = sd(Rel_Abundance, na.rm = TRUE), .groups = "drop")

otu_summary$MAG_species <- paste(otu_summary$MAG_ID, " (", otu_summary$Species, ")", sep = "")

table (otu_summary$MAG_species)

species_color_carr_intense <- list("Species" = c(
  "CA_2A_4 (Leuconostoc mesenteroides)" = "#42A5F5",
  "CA_2C_2 (Leuconostoc mesenteroides)" = "paleturquoise4", 
  'CA_7C_7 (Leuconostoc inhae)' = "#004D99",  
  'CA_14A_1 (Lactiplantibacillus plantarum)' = "#8E24AA",  
  'CA_14A_5 (Levilactobacillus brevis)' = "#FFD700"  
))

g_MAGs_carr_species <- ggplot(otu_summary, aes(x = sampling_day, y = mean_abund, color = MAG_species, group = MAG_species)) +
  geom_line() +  
  geom_point() +  
  geom_ribbon(aes(ymin = mean_abund - sd_abund, ymax = mean_abund + sd_abund, fill = MAG_species), 
              alpha = 0.2, color = NA) + 
  theme_minimal() +
  scale_color_manual(values = species_color_carr_intense$Species) +  
  scale_fill_manual(values = species_color_carr_intense$Species) +  
  labs(x = "Fermentation time (Day)", y = "Relative abundance (%)", color = "MAG (Species)", fill = "MAG (Species)") +
  scale_x_continuous(breaks = unique(otu_summary$sampling_day)) +
  theme(legend.position = "right")
g_MAGs_carr_species

```
## Data import

```{r Import data, message=FALSE, warning=FALSE}

#Import count matrix of all samples
count_metag_cabbage <- read.table("results/SnakeMAGs/MAPPING_BWA-MEM2/Cabbage/presence_counts.csv") 
count_metag_cabbage <- count_metag_cabbage[-1:-2,] #delete 2 firsts line that don't corresponding to contigs.

samples_sum_reads_cabbage <-colSums(count_metag_cabbage)


count_metag_cabbage <- count_metag_cabbage%>%
  rownames_to_column(.,var="Contigs")


info_contigs_cabbage <- read.table("METADATA/Contig_MAG_info_cabbage.txt",sep="\t",header = TRUE) 

count_metag_cabbage_info <- right_join(count_metag_cabbage,info_contigs_cabbage,by="Contigs")


common <- intersect(count_metag_cabbage$Contigs, info_contigs_cabbage$Contigs)
cat("Contigs dans count_metag :", nrow(count_metag_cabbage), "\n")
cat("Contigs dans info_contigs :", nrow(info_contigs_cabbage), "\n")
cat("Contigs en commun :", length(common), "\n")


info_MAGs_cabbage0 <- read.table("METADATA/MAG_info_cabbage.txt",sep="\t",header = TRUE)
info_MAGs_cabbage <- info_MAGs_cabbage0%>%
  column_to_rownames(.,var="MAG_ID")




# Supposons que df_abundance contient les abondances des contigs par échantillon
df_MAG_abundance_cabb <- count_metag_cabbage_info %>%
  group_by(MAGs) %>% 
  summarise(SA_0A = sum(SA_0A, na.rm = TRUE),
            SA_0B = sum(SA_0B, na.rm = TRUE),
            SA_0C = sum(SA_0C, na.rm = TRUE),
            SA_2A = sum(SA_2A, na.rm = TRUE),
            SA_2B = sum(SA_2B, na.rm = TRUE),
            SA_2C = sum(SA_2C, na.rm = TRUE),
            SA_7A = sum(SA_7A, na.rm = TRUE),
            SA_7B = sum(SA_7B, na.rm = TRUE),
            SA_7C = sum(SA_7C, na.rm = TRUE),
            SA_14A = sum(SA_14A, na.rm = TRUE),
            SA_14B = sum(SA_14B, na.rm = TRUE),
            SA_14C = sum(SA_14C, na.rm = TRUE),
            SA_21A = sum(SA_21A, na.rm = TRUE),
            SA_21B = sum(SA_21B, na.rm = TRUE),
            SA_21C = sum(SA_21C, na.rm = TRUE),
            SA_28A = sum(SA_28A, na.rm = TRUE),
            SA_28B = sum(SA_28B, na.rm = TRUE),
            SA_28C = sum(SA_28C, na.rm = TRUE))%>%
  column_to_rownames(.,var="MAGs")
 
```



### Abundance table for Cabbage
We created a Phyloseq object that contains the abundance table and metadata of all samples.
```{r all abundance table, message=FALSE, warning=FALSE}

# Check thaht MAG name correspond between to files
identical(rownames(df_MAG_abundance_cabb), rownames(info_MAGs_cabbage))  # Must be : TRUE

df_MAG_abundance_cabb <- df_MAG_abundance_cabb[order(rownames(df_MAG_abundance_cabb)), , drop = FALSE]
info_MAGs_cabbage <- info_MAGs_cabbage[order(rownames(info_MAGs_cabbage)), , drop = FALSE]

identical(rownames(df_MAG_abundance_cabb), rownames(info_MAGs_cabbage))  

# Vérifier si les noms des échantillons correspondent
identical(colnames(df_MAG_abundance_cabb), rownames(sample_metag_metadata))  


# Check consistency
all(colnames(df_MAG_abundance_cabb) %in% rownames(sample_metag_metadata)) ## sample names between metadata and otu table
all(rownames(df_MAG_abundance_cabb) %in% rownames(info_MAGs_cabbage)) ## taxa names between taxonomic and otu tables

otu_tab_cabbage <- otu_table(as.matrix(df_MAG_abundance_cabb), taxa_are_rows = TRUE)
tax_tab_cabbage <- tax_table(as.matrix(info_MAGs_cabbage))

obj_metag_cabbage <- phyloseq(otu_tab_cabbage, sample_data(sample_metag_metadata), tax_tab_cabbage)

otu_table(obj_metag_cabbage)
tax_table(obj_metag_cabbage)

```


#### Normalisation

We transformed abundance data by TPM. We normalized by contig length and
transformed to proportions

```{r TPM transformation, message=FALSE, warning=FALSE}
# Extract abundance table 
abundance_table_cabb <- as.data.frame(otu_table(obj_metag_cabbage)) %>% rownames_to_column(var = "MAG_ID")
#recovery the abundance table of obj_clean in the form of dataframe. The row name is the contigs identify.

# Add contig length and set the vector of contig length
abundance_table_cabb2 <- left_join(abundance_table_cabb, info_MAGs_cabbage0, by = "MAG_ID")
vect_length_cabb <- abundance_table_cabb2$Genome.size..bp. 
# join the abundance table and the list of contig length into the data frame abundance_table2 based on the column contig identity.
# creat the vector vect_length that contain all length of the abundance table.
  
# Transform counts to TPM
## First divide the count matrix by the vector of contig length
obj_metag_cabbage_TPM <- obj_metag_cabbage
otu_table(obj_metag_cabbage_TPM) <- otu_table(obj_metag_cabbage_TPM) / vect_length_cabb 

## Second transform the count matrix into proportions 
count_to_prop_cabb_sub <- function(x) {return( x / sum(x) )}
obj_metag_cabbage_TPM <- transform_sample_counts(obj_metag_cabbage_TPM, count_to_prop_cabb_sub)
#sample_sums(obj_All_TPM)
# creat function count_to_prop that calculate the proportion based of the count of abundance divide by the sum of the abundance.
# transform_sample_counts: Transform abundance data in an otu_table, sample-by-sample.
# sample_sums: Returns the total number of individuals observed from each sample.

abundance_cabb_TPM <- otu_table(obj_metag_cabbage_TPM)

```


```{r Pheatmap normalized filtred, message=FALSE, warning=FALSE}

#transform log function
log_fun <- function(x){
  (log((x + 0.000001),10))
}

MAGs_cabbage_abundance <-  data.frame(otu_table(obj_metag_cabbage_TPM))
MAGs_cabbage_abundance_norm <- t(apply(MAGs_cabbage_abundance, 1, log_fun))
sample_annotation_cabb <- data.frame(sample_data(obj_metag_cabbage_TPM))

#Extract annotation data 
MAG_annotations_cabb <- data.frame(tax_table(obj_metag_cabbage_TPM))
table(MAG_annotations_cabb$Species)

# Extract vOTU metadata
Classification_cabb <- MAG_annotations_cabb["Species"]

# Creat dataframe `row_ann` for the row annotation
row_ann_cabb <- data.frame(Classification_cabb)
col_ann_cabb <- sample_annotation_cabb["sampling_day"]
col_ann_cabb$sampling_day <- as.character(col_ann_cabb$sampling_day)
col_ann_cabb$sampling_day <- factor (col_ann_cabb$sampling_day, levels = c("0","2","7","14","21","28"))

custom_colnames <- 
  if ("Figure_label" %in% colnames(sample_annotation_cabb)) {
  custom_colnames_MAGS_cabb <- sample_annotation_cabb$Figure_label
} else {
  stop(" 'NewNames' column doesn't exist in samples data")
}

names(col_ann_cabb)[names(col_ann_cabb) == "sampling_day"] <- "Sampling day"

ann_color_cabb = list("Sampling day"=c("0"="#1C3333FF" , 
                            "2"="#226060FF", 
                            "7"="#639CA4FF",
                            "14"="#D2AD7CFF",
                            "21"="#BE7245FF",
                            "28"="#8C311CFF"),
                 "Species"=c("Leuconostoc mesenteroides"= "#a6cee3",
                             "Leuconostoc carnosum"="#1f78b4",
                             "Pantoea agglomerans" = "#b2df8a",
                             "Pediococcus parvulus" = "#33a02c",
                             "Levilactobacillus parabrevis" = "#fb9a99",
                             "Levilactobacillus yonginensis" = "#e31a1c",
                             "Secundilactobacillus malefermentans" = "#fdbf6f",
                              "Secundilactobacillus silagei"= "#ff7f00",
                               'Lactiplantibacillus paraplantarum' = '#cab2d6',
                                'Weissella soli' = '#ffff99'))

ph_MAGs_cabb <- pheatmap(MAGs_cabbage_abundance,
               show_rownames = TRUE,
               cluster_cols = FALSE,
               annotation_row = row_ann_cabb,
               annotation_col = col_ann_cabb,
               annotation_colors = ann_color_cabb,
               fontsize_col = 7,
               fontsize = 6,
               angle_col = 45,
               labels_col = custom_colnames_MAGS_cabb)
               #legend_breaks = c(-5, -3, -1, max(MAGs_carrot_abundance_norm)),
               #legend_labels = c("-5", "-3", "-1", "log10"))

ph_MAGs_cabb

t.MAGs_cabbage_abundance.scale <- scale(t(MAGs_cabbage_abundance)) 
MAGs_cabbage_abundance.scale <- t(t.MAGs_cabbage_abundance.scale) 
breaks <- seq(-3, 3, length.out = 101)

color_palette <- colorRampPalette(c("#2c7bb6","white","#d7191c"))(length(breaks))

ph_MAGs_cabb_Zs <- pheatmap(MAGs_cabbage_abundance.scale,
               show_rownames = TRUE,
               cluster_cols = FALSE,
               annotation_row = row_ann_cabb,
               annotation_col = col_ann_cabb,
               annotation_colors = ann_color_cabb,
               fontsize_col = 7,
               fontsize = 6,
               angle_col = 45,
               labels_col = custom_colnames,
               legend_breaks = c(-3, 0, 2.5, 3),
               legend_labels = c("-3", "0", "2.5", "Z-score"),
               color = color_palette,  # Appliquer la palette de couleurs
               breaks = breaks) 
ph_MAGs_cabb_Zs



## A graph showing the relative abundance of MAGs during fermentation. (idem carrot part)

otu_table_df <- as.data.frame(t(otu_table(obj_metag_cabbage_TPM)))
otu_table_df$SampleID <- rownames(otu_table_df)

metadata_df <- as.data.frame(sample_data(obj_metag_cabbage_TPM))
metadata_df$SampleID <- rownames(metadata_df)

tax_df <- as.data.frame(tax_table(obj_metag_cabbage_TPM))
tax_df$MAG_ID <- rownames(tax_df)

tax_df <- tax_df %>%
  select(MAG_ID, Species)

otu_long <- otu_table_df %>%
  pivot_longer(cols = -SampleID, names_to = "MAG_ID", values_to = "Abundance") %>%
  left_join(metadata_df, by = "SampleID") %>%
  left_join(tax_df, by = "MAG_ID")

otu_long <- otu_long %>%
  group_by(SampleID) %>%
  mutate(Rel_Abundance = (Abundance / sum(Abundance, na.rm = TRUE)*100))

otu_summary <- otu_long %>%
  group_by(sampling_day, Species,MAG_ID) %>%
  summarise(mean_abund = mean(Rel_Abundance, na.rm = TRUE),
            sd_abund = sd(Rel_Abundance, na.rm = TRUE), .groups = "drop")

otu_summary$MAG_species <- paste(otu_summary$MAG_ID, " (", otu_summary$Species, ")", sep = "")

table (otu_summary$MAG_species)


species_color_cabb_intense <- list("Species" = c(
  "SA_2C_6 (Leuconostoc mesenteroides)" = "#42A5F5",  
  "SA_0A_9 (Leuconostoc carnosum)" = "#004D99",  
  "SA_0B_4 (Pantoea agglomerans)" = "#33CC33", 
  "SA_28A_1 (Pediococcus parvulus)" = "#006600",  
  "SA_7A_15 (Levilactobacillus parabrevis)" = "#F44336",  
  "SA_14A_9 (Levilactobacillus yonginensis)" = "indianred4",  
  "SA_21C_10 (Secundilactobacillus malefermentans)" = "#FF9800",  
  "SA_28B_7 (Secundilactobacillus silagei)" = "darkorange3", 
  "SA_14B_5 (Lactiplantibacillus paraplantarum)" = "#8E24AA",  
  "SA_0C_7 (Weissella soli)" = "#FFD700"  
))

g_MAGs_cabb_species <- ggplot(otu_summary, aes(x = sampling_day, y = mean_abund, color = MAG_species, group = MAG_species)) +
  geom_line() +  
  geom_point() +  
  geom_ribbon(aes(ymin = mean_abund - sd_abund, ymax = mean_abund + sd_abund, fill = MAG_species), 
              alpha = 0.2, color = NA) +  
  theme_minimal() +
  scale_color_manual(values = species_color_cabb_intense$Species) +  
  scale_fill_manual(values = species_color_cabb_intense$Species) +
  labs(x = "Fermentation time (Day)", y = "Relative abundance (%)", color = "MAG (Species)", fill = "MAG (Species)") +
  scale_x_continuous(breaks = unique(otu_summary$sampling_day)) +
  theme(legend.position = "right")
g_MAGs_cabb_species


```


## Correlation virome bacteriome
You need to run the Metavirome_analyse.Rmd script to do this part.
```{r Mantel correlation, message=FALSE, warning=FALSE}
# give the same sample names to both virome and bacteriome datasets
sample_names(obj_metag_cabbage_TPM) <- sample_names(obj_cabbage_TPM)
# PCoA phage
sample_sums(obj_cabbage_TPM)
dist_bc_phage <- phyloseq::distance(obj_cabbage_TPM, method = "bray") 
# PCoA bacteria
sample_sums(obj_metag_cabbage_TPM)
#count_to_prop <- function(x) {return( x / sum(x) )}
#Bact_dyn_fs_tss <- transform_sample_counts(obj_metag_cabbage_TPM, count_to_prop)
sample_sums(obj_metag_cabbage_TPM)
dist_bc_bact <- phyloseq::distance(obj_metag_cabbage_TPM, method = "bray") 
# mantel test
mantel(dist_bc_phage, dist_bc_bact, method="pearson", permutations = 999)


```


